---
title: "Hands-on Exercise 4c"
subtitle: "Visualising Uncertainty — Error Bars, ggdist & HOPs"
author: "Thi Thanh Van Tran"
date: "May 3, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    number-depth: 3
    code-fold: true
    code-summary: "Show code"
    code-tools: false
    code-copy: true
    css: styles.css
    theme: cosmo
    page-layout: article
execute:
  eval: true
  echo: true
  warning: false
  message: false
  freeze: false
knitr:
  opts_chunk:
    out.width: "100%"
    fig.width: 10
    fig.align: "center"
    dpi: 120
---

```{=html}
<style>
  .quarto-figure, figure {
    max-width: 100% !important; width: 100% !important;
    margin: 0.5rem auto !important; overflow: visible !important;
  }
  img.figure-img, .cell > img, .cell-output > img, .cell-output-display img {
    width: 100% !important; height: auto !important;
    object-fit: contain !important; display: block; margin: 0.3rem auto;
  }
  .cell-output-display {
    margin: 0.4rem 0 !important; padding: 0 !important; overflow: visible !important;
  }
  .callout { margin-top: 0.75rem; }
  h2 { margin-top: 2rem; }
</style>
```

---

::: welcome-section
## Visualising Uncertainty — Error Bars, ggdist & HOPs

Welcome! Today, let's learn visualising uncertainty is a modern and essential skill in statistical graphics. Applied to **DKI Jakarta COVID-19 sub-district data**, this exercise covers:

- Plotting **standard error bars** and **confidence interval bars** with `ggplot2`
- Building **interactive error bars** via `ggplot2` + `plotly` + `crosstalk` + `DT`
- Creating advanced visualisations with **`ggdist`** — point intervals and gradient intervals
- Generating animated **Hypothetical Outcome Plots (HOPs)** with `ungeviz` + `gganimate`
:::

::: callout-important
**Key distinction:** The *uncertainty of a point estimate* (how reliable is the mean?) is not the same as *variation in the sample* (how spread are individual values?). All techniques here address the former.
:::


---

## Getting Started

### Installing and Loading Packages

Packages used in this exercise:

- `tidyverse` — data wrangling and `ggplot2`
- `plotly` — interactive plots
- `gganimate` — animation
- `DT` — interactive HTML tables
- `crosstalk` — cross-widget linked brushing and filtering
- `ggdist` — distribution and uncertainty visualisation

```{r}
#| code-fold: false
pacman::p_load(
  tidyverse,
  plotly,
  crosstalk,
  DT,
  ggdist,
  ggridges,
  colorspace,
  gganimate,
  ungeviz,
  janitor,
  scales,
  knitr,
  kableExtra
)

# ungeviz must be installed once via:
# devtools::install_github("wilkelab/ungeviz")
```

### Importing and Preparing Data

The dataset `COVID-19_DKI_Jakarta.csv` contains sub-district-level COVID-19 case counts across six cities in DKI Jakarta. **Case Fatality Rate (CFR)** — the proportion of confirmed cases resulting in death — is the focal variable throughout this exercise.

```{r}
#| code-fold: false
covid_raw <- read_csv("data/COVID-19_DKI_Jakarta.csv")

covid <- covid_raw %>%
  janitor::clean_names() %>%
  rename(id = sub_district_id, subdistrict = sub_district) %>%
  mutate(
    cfr        = if_else(positive > 0, death / positive * 100, NA_real_),
    city_label = case_when(
      str_detect(toupper(city), "SELATAN") ~ "Jkt. Selatan",
      str_detect(toupper(city), "UTARA")   ~ "Jkt. Utara",
      str_detect(toupper(city), "BARAT")   ~ "Jkt. Barat",
      str_detect(toupper(city), "PUSAT")   ~ "Jkt. Pusat",
      str_detect(toupper(city), "TIMUR")   ~ "Jkt. Timur",
      str_detect(toupper(city), "SERIBU")  ~ "Kep. Seribu",
      TRUE ~ city
    )
  )

# Retain only sub-districts with >= 10 positive cases for reliable CFR
covid_cfr <- covid %>% filter(!is.na(cfr), positive >= 10)
```

---

## Data Exploration

Before applying any uncertainty technique, we need to understand two properties of the data: **how many sub-districts each city has** and **the shape of CFR distributions**. These directly determine how wide every interval will be and whether the mean is even a suitable summary.

### Summary Statistics by City

`group_by()`, `summarise()`, and `mutate()` derive the summary table — the same pipeline powering all `ggplot2` error bar plots that follow.

- **n** — sub-district count (governs interval width)
- **mean**, **sd** — point estimate and spread of CFR
- **se** = `sd / sqrt(n - 1)` — precision of the mean

```{r}
#| code-fold: false
cfr_sum <- covid_cfr %>%
  group_by(city_label) %>%
  summarise(
    n    = n(),
    mean = mean(cfr, na.rm = TRUE),
    sd   = sd(cfr, na.rm = TRUE)
  ) %>%
  mutate(se = sd / sqrt(n - 1))

knitr::kable(cfr_sum, format = "html", digits = 2,
             caption = "CFR summary statistics by city (sub-districts with ≥ 10 positive cases)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

### Distribution of CFR by City

```{r}
#| fig-height: 4.5
ggplot(covid_cfr, aes(x = cfr, fill = city_label)) +
  geom_histogram(bins = 25, alpha = 0.75, position = "identity") +
  facet_wrap(~city_label, scales = "free_y", nrow = 2) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title    = "Distribution of Case Fatality Rate (CFR) by City",
    subtitle = "Sub-districts with ≥ 10 positive cases",
    x = "CFR (%)", y = "Count"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none")
```

::: callout-note
## What I Learned

- **Unequal sample sizes** are immediately clear: Jakarta Timur has many sub-districts; Kepulauan Seribu has very few. This single fact will explain why Kepulauan Seribu produces wider intervals in every technique that follows.
- **All distributions are right-skewed**: the mean is pulled up by a handful of high-CFR sub-districts, making it a poor summary. This motivates switching to medians in the `ggdist` sections.
- This EDA is the opening act of a continuous story — every interval we plot next is a direct answer to what we see here.
:::

---

## Visualising Uncertainty with `ggplot2`

A **point estimate** is a single number such as a mean. Uncertainty is expressed as standard error, confidence interval, or credible interval. The `cfr_sum` table above feeds all three plots in this section.

### Standard Error Bars

Error bars use the formula **mean ± SE**. For `geom_point()`, `stat = "identity"` is required because we are plotting pre-computed values, not raw observations.

```{r}
#| fig-height: 5
ggplot(cfr_sum) +
  geom_errorbar(
    aes(x = city_label, ymin = mean - se, ymax = mean + se),
    width = 0.2, colour = "black", alpha = 0.9, linewidth = 0.5
  ) +
  geom_point(
    aes(x = city_label, y = mean),
    stat = "identity", color = "red", size = 1.5, alpha = 1
  ) +
  labs(
    title    = "Standard error of mean CFR (%) by city",
    subtitle = "Error bars = mean ± 1 SE",
    x = NULL, y = "Mean CFR (%)"
  ) +
  theme_minimal(base_size = 12)
```

::: callout-note
## What I Learned

- Jakarta Timur has the narrowest SE bars (largest n); Kepulauan Seribu the widest — exactly as the EDA predicted.
- SE bars express *precision of the mean estimate*, not the plausible range of the true mean. That requires a confidence interval.
:::

### Confidence Interval of Point Estimates

The **95% CI** uses **mean ± 1.96 × SE**. Cities are sorted from highest to lowest mean CFR using `reorder()`. `labs()` is used to update the axis label.

```{r}
#| fig-height: 5
ggplot(cfr_sum) +
  geom_errorbar(
    aes(
      x    = reorder(city_label, -mean),
      ymin = mean - 1.96 * se,
      ymax = mean + 1.96 * se
    ),
    width = 0.2, colour = "black", alpha = 0.9, linewidth = 0.5
  ) +
  geom_point(
    aes(x = city_label, y = mean),
    stat = "identity", color = "red", size = 1.5, alpha = 1
  ) +
  labs(
    x        = "City",
    title    = "95% confidence interval of mean CFR (%) by city",
    subtitle = "Error bars = mean ± 1.96 × SE; sorted by mean CFR"
  ) +
  theme_minimal(base_size = 12)
```

::: callout-note
## What I Learned

- Overlapping CIs between Jakarta Timur, Selatan and Barat show their means are **not statistically distinguishable** — ranking cities without this view is misleading.
- Kepulauan Seribu's CI spans nearly the full observed CFR range. A single-point mean here is analytically meaningless without the interval.
:::

### Interactive Error Bars — 99% Confidence Interval

The **99% CI** uses **mean ± 2.58 × SE**. `SharedData$new()` from `crosstalk` creates a shared data object linking the DT table and the plotly chart — clicking a table row highlights the corresponding city in the plot.

```{r}
#| fig-height: 5
shared_df <- SharedData$new(cfr_sum)

bscols(
  widths = c(4, 8),

  DT::datatable(
    shared_df,
    rownames = FALSE,
    class    = "compact",
    width    = "100%",
    options  = list(pageLength = 10, scrollX = TRUE),
    colnames = c("City", "No. of Sub-districts",
                 "Avg CFR (%)", "Std Dev", "Std Error")
  ) %>%
    formatRound(columns = c("mean", "sd", "se"), digits = 2),

  ggplotly(
    ggplot(shared_df) +
      geom_errorbar(
        aes(
          x    = reorder(city_label, -mean),
          ymin = mean - 2.58 * se,
          ymax = mean + 2.58 * se
        ),
        width = 0.2, colour = "black", alpha = 0.9, size = 0.5
      ) +
      geom_point(
        aes(
          x    = city_label,
          y    = mean,
          text = paste(
            "City:", city_label,
            "<br>N:", n,
            "<br>Avg CFR:", round(mean, 2), "%",
            "<br>99% CI: [",
            round(mean - 2.58 * se, 2), ",",
            round(mean + 2.58 * se, 2), "]"
          )
        ),
        stat = "identity", color = "red", size = 1.5, alpha = 1
      ) +
      xlab("City") + ylab("Mean CFR (%)") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +
      ggtitle("99% Confidence interval of mean CFR (%) by city"),
    tooltip = "text"
  )
)
```

::: callout-note
## What I Learned

- `SharedData$new()` + `bscols()` creates **linked interactivity**: table row selections highlight plotly points; hover tooltips built with `paste()` + `<br>` show CI bounds and sample size in one view.
- Custom column names via `colnames = c(...)` in `DT::datatable()` improve readability without touching the underlying data.
- The 99% CI (multiplier 2.58) is wider than the 95% (1.96) — the gap is most visible for Kepulauan Seribu, instantly cross-checkable against its small n in the linked table.
:::

---

## Visualising Uncertainty with `ggdist`

[`ggdist`](https://mjskay.github.io/ggdist/index.html) provides `ggplot2` geoms and stats purpose-built for distributions and uncertainty. It supports both **frequentist** (bootstrap / confidence distributions) and **Bayesian** (probability distributions via `tidybayes`) frameworks. Unlike `geom_errorbar()`, `ggdist` functions work directly on raw data — no pre-computation required.

### `stat_pointinterval()` — Default

`stat_pointinterval()` draws a point estimate with nested intervals directly from raw data. The default shows **66% and 95% quantile-based intervals**.

```{r}
#| fig-height: 5
covid_cfr %>%
  ggplot(aes(x = city_label, y = cfr)) +
  stat_pointinterval() +
  labs(
    title    = "Visualising confidence intervals of mean CFR",
    subtitle = "Mean Point + Multiple-interval plot",
    x = NULL, y = "CFR (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

::: callout-note
## What I Learned

- No `summarise()` needed — `stat_pointinterval()` computes everything from raw data using **quantile-based** intervals, not the normal approximation assumed by `geom_errorbar()`.
- The long upper tails for Jakarta Timur are structurally visible in the interval — right-skew is now built into the uncertainty representation itself.
:::

### `stat_pointinterval()` — Median + Quantile Interval

Using `.width = 0.95`, `.point = median`, `.interval = qi`:

```{r}
#| fig-height: 5
covid_cfr %>%
  ggplot(aes(x = city_label, y = cfr)) +
  stat_pointinterval(
    .width    = 0.95,
    .point    = median,
    .interval = qi
  ) +
  labs(
    title    = "Visualising confidence intervals of median CFR",
    subtitle = "Median Point + Quantile-interval plot (.width = 0.95)",
    x = NULL, y = "CFR (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

::: callout-note
## What I Learned

- Switching to **median** lowers the point estimate for every city — confirming the EDA finding that a few high-CFR sub-districts were inflating the mean.
- `.point`, `.interval`, `.width` are the three key arguments: they control *what to estimate*, *how to build the interval*, and *how wide* — all without changing the plot structure.
:::

### `stat_pointinterval()` — 95% and 99% Intervals (*Your Turn*)

Passing `.width = c(0.95, 0.99)` overlays both intervals simultaneously.

```{r}
#| fig-height: 5
covid_cfr %>%
  ggplot(aes(x = city_label, y = cfr)) +
  stat_pointinterval(
    .width      = c(0.95, 0.99),
    show.legend = FALSE
  ) +
  labs(
    title    = "Visualising 95% and 99% confidence intervals of mean CFR",
    subtitle = "Mean Point + Multiple-interval plot (.width = c(0.95, 0.99))",
    x = NULL, y = "CFR (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

::: callout-note
## What I Learned

- A vector in `.width` draws nested intervals at different thicknesses — the gap between 95% and 99% is compact for large-n cities and wide for Kepulauan Seribu.
- `show.legend = FALSE` keeps the plot clean; the subtitle carries the annotation.
:::

### `stat_gradientinterval()` — Gradient Interval

```{r}
#| fig-height: 5
covid_cfr %>%
  ggplot(aes(x = city_label, y = cfr)) +
  stat_gradientinterval(
    fill        = "skyblue",
    show.legend = TRUE
  ) +
  labs(
    title    = "Visualising confidence intervals of mean CFR",
    subtitle = "Gradient + interval plot — colour intensity reflects density",
    x = NULL, y = "CFR (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

::: callout-note
## What I Learned

- `stat_gradientinterval()` fills the interval with a **density-proportional colour gradient** — darker regions = more sub-districts clustered there. This shows *where within the interval the data actually concentrate*, which flat error bars cannot.
- For Jakarta Timur, the dark band sits at the lower end, confirming that most sub-districts have modest CFR even though the mean is pulled up by outliers — the most information-dense static view in this exercise.
:::

---

## Visualising Uncertainty with Hypothetical Outcome Plots (HOPs)

### Installing the `ungeviz` Package

```{r}
#| eval: false
#| code-fold: false
devtools::install_github("wilkelab/ungeviz")
```

> You only need to perform this step once.

### Loading the Package

```{r}
#| code-fold: false
library(ungeviz)
```

### Building the HOPs

HOPs animate a series of **bootstrap draws** from each group. Each frame shows one draw; the animation lets viewers *watch* the estimate move — making sampling variability tangible without any statistical training required.

The plot combines:

- **Blue jittered points** — all raw CFR values per city  
- **Orange horizontal line (`geom_hpline`)** — the mean CFR of one bootstrap sample, redrawn each frame via `sampler(25, group = city_label)`  
- **`transition_states(.draw, 1, 3)`** — cycles through 25 draws; pause is 3× longer than the transition

```{r}
#| fig-height: 7
#| cache: true
ggplot(data = covid_cfr,
       aes(x = factor(city_label), y = cfr)) +
  geom_point(
    position = position_jitter(height = 0.3, width = 0.05),
    size     = 0.4,
    color    = "#0072B2",
    alpha    = 1 / 2
  ) +
  geom_hpline(
    data   = sampler(25, group = city_label),
    height = 0.6,
    color  = "#D55E00"
  ) +
  labs(
    title    = "Hypothetical Outcome Plots (HOPs) — CFR by City",
    subtitle = "Orange line = bootstrap sample mean; each frame = one draw",
    x = NULL, y = "CFR (%)"
  ) +
  theme_bw(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  transition_states(.draw, 1, 3)
```

::: callout-note
## What I Learned

- `sampler(25, group = city_label)` from `ungeviz` generates 25 bootstrap resamples per city, adding a `.draw` column that `transition_states()` animates over. `geom_hpline()` draws one short horizontal line per group per frame — purpose-built for HOPs.
- **The animation unifies the whole story**: Kepulauan Seribu's orange line jumps wildly (small n, high uncertainty); Jakarta Timur's barely moves (large n, stable estimate). Everything the static plots pointed toward becomes unmissable in motion.
- For non-technical audiences, watching the line move is far more intuitive than reading "95% CI = [x, y]". Uncertainty becomes experiential rather than abstract.
:::

---

## Summary

Every technique in this exercise was a direct answer to what the EDA revealed: right-skewed CFR and vastly unequal sample sizes across Jakarta's cities.

| Technique | Key function / argument | Core insight |
|---|---|---|
| Standard error bars | `geom_errorbar(ymin = mean - se, ymax = mean + se)` | Kep. Seribu most imprecise; Jakarta Timur most reliable |
| 95% CI bars | `mean ± 1.96 × se`, `reorder()` | Several cities statistically indistinguishable |
| Interactive 99% CI | `SharedData` + `bscols()` + `ggplotly()` + `DT` | Linked table+plot cross-references CI width against n |
| Point interval (default) | `stat_pointinterval()` | Quantile intervals expose right-skew; no normality assumed |
| Median + qi | `.point = median`, `.interval = qi`, `.width = 0.95` | Median lower than mean in all cities — confirms skew |
| 95% + 99% intervals | `.width = c(0.95, 0.99)`, `show.legend = FALSE` | Interval gap largest for small-n cities |
| Gradient interval | `stat_gradientinterval(fill = "skyblue")` | Density gradient reveals where data truly cluster |
| HOPs | `sampler(25, group = city_label)` + `geom_hpline()` + `transition_states(.draw, 1, 3)` | Animation makes uncertainty impossible to ignore |

A single CFR mean per city looks clean — but every layer added here erodes that false confidence. Visualising uncertainty is not decorative; it is the difference between a reliable public health decision and a misleading one.


