---
title: "Hands-on Exercise 4a"
subtitle: "Visualising Distribution — Ridgeline & Raincloud Plots"
author: "Thi Thanh Van Tran"
date: "May 3, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    number-depth: 3
    code-fold: true
    code-summary: "Show code"
    code-tools: false
    code-copy: true
    css: styles.css
    theme: cosmo
    page-layout: article
execute:
  eval: true
  echo: true
  warning: false
  message: false
  freeze: false
knitr:
  opts_chunk:
    out.width: "100%"
    fig.width: 10
    fig.align: "center"
    dpi: 120
---

```{=html}
<style>
  .quarto-figure, figure {
    max-width: 100% !important; width: 100% !important;
    margin: 0.5rem auto !important; overflow: visible !important;
  }
  img.figure-img, .cell > img, .cell-output > img, .cell-output-display img {
    width: 100% !important; height: auto !important;
    object-fit: contain !important; display: block; margin: 0.3rem auto;
  }
  .cell-output-display {
    margin: 0.4rem 0 !important; padding: 0 !important; overflow: visible !important;
  }
  .callout { margin-top: 0.75rem; }
  h2 { margin-top: 2rem; }
</style>
```

---

::: welcome-section
# Visualising Distribution — Ridgeline & Raincloud Plots

Welcome! This exercise introduces two advanced distribution techniques applied to **COVID-19 sub-district data for DKI Jakarta**:

- **Ridgeline plot** — compares distribution shapes across many groups on a shared axis using overlapping density curves. Best for 5+ groups.
- **Raincloud plot** — layers a half-density **cloud** (above the line), a compact **boxplot** (on the line), and individual **rain dots** (below the line). Nothing is hidden about the distribution.
:::

---

## Getting Started

### Installing and Loading Packages

```{r}
#| code-fold: false
pacman::p_load(
  ggdist,      # stat_halfeye(), stat_dots()
  ggridges,    # geom_density_ridges(), stat_density_ridges()
  ggthemes,    # theme_economist(), theme_ridges()
  colorspace,  # lighten()
  tidyverse,   # ggplot2 + data wrangling
  scales,      # comma_format()
  plotly,      # interactive charts
  patchwork,   # side-by-side layouts
  janitor,     # clean_names()
  kableExtra   # styled tables
)
```

### Importing and Preparing Data

```{r}
#| code-fold: false
covid_raw <- read_csv("data/COVID-19_DKI_Jakarta.csv")

covid <- covid_raw %>%
  janitor::clean_names() %>%
  rename(id = sub_district_id, subdistrict = sub_district) %>%
  mutate(
    cfr           = if_else(positive > 0, death / positive * 100,     NA_real_),
    recovery_rate = if_else(positive > 0, recovered / positive * 100, NA_real_),
    active        = positive - recovered - death,
    # Short two-line city labels — prevent x-axis overlap in un-flipped charts
    city_label = case_when(
      str_detect(toupper(city), "SELATAN") ~ "Jkt.\nSelatan",
      str_detect(toupper(city), "UTARA")   ~ "Jkt.\nUtara",
      str_detect(toupper(city), "BARAT")   ~ "Jkt.\nBarat",
      str_detect(toupper(city), "PUSAT")   ~ "Jkt.\nPusat",
      str_detect(toupper(city), "TIMUR")   ~ "Jkt.\nTimur",
      str_detect(toupper(city), "SERIBU")  ~ "Kep.\nSeribu",
      TRUE ~ city
    )
  )

# Sub-districts with reliable CFR (>= 10 positive cases)
covid_cfr <- covid %>% filter(!is.na(cfr), positive >= 10)
```

---

## Data Exploration

::: {.panel-tabset}

### Structure & Coverage

```{r}
#| fig-height: 3.5

p_count <- covid %>%
  count(city, name = "n") %>%
  ggplot(aes(x = reorder(city, n), y = n, fill = city)) +
  geom_col(width = 0.65, show.legend = FALSE, alpha = 0.85) +
  geom_text(aes(label = n), hjust = -0.2, fontface = "bold", size = 3.8) +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(title = "Sub-districts per City",
       subtitle = "Unequal coverage affects ridgeline curve smoothness",
       x = NULL, y = "Sub-districts") +
  theme_minimal(base_size = 11)

p_missing <- tibble(
  Variable = c("positive","recovered","death","cfr","recovery_rate"),
  Missing  = c(sum(is.na(covid$positive)), sum(is.na(covid$recovered)),
               sum(is.na(covid$death)),    sum(is.na(covid$cfr)),
               sum(is.na(covid$recovery_rate)))
) %>%
  ggplot(aes(x = reorder(Variable, Missing), y = Missing)) +
  geom_col(fill = "#e74c3c", width = 0.55, alpha = 0.85) +
  geom_text(aes(label = Missing), hjust = -0.2, fontface = "bold", size = 3.8) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.3))) +
  labs(title = "Missing Values per Column",
       subtitle = "cfr / recovery_rate NA where positive = 0",
       x = NULL, y = "Count") +
  theme_minimal(base_size = 11)

p_count | p_missing
```

::: callout-note
## What I Learned

The dataset is cross-sectional — cumulative totals per sub-district with no time dimension. Jakarta Timur has the most sub-districts (~65), producing smoother ridgeline density curves. `cfr` and `recovery_rate` are `NA` where `positive = 0` to prevent division by zero.
:::

### Case Distributions

```{r}
#| fig-height: 3.5

p1 <- ggplot(covid, aes(x = positive)) +
  geom_histogram(bins = 35, fill = "#e74c3c", colour = "white", alpha = 0.85) +
  scale_x_continuous(labels = comma_format()) +
  labs(title = "Positive Cases", x = "Cases / sub-district", y = "Count") +
  theme_minimal(base_size = 10)

p2 <- ggplot(covid, aes(x = recovered)) +
  geom_histogram(bins = 35, fill = "#27ae60", colour = "white", alpha = 0.85) +
  scale_x_continuous(labels = comma_format()) +
  labs(title = "Recovered", x = "Recovered / sub-district", y = "Count") +
  theme_minimal(base_size = 10)

p3 <- ggplot(covid, aes(x = death)) +
  geom_histogram(bins = 35, fill = "#7f8c8d", colour = "white", alpha = 0.85) +
  scale_x_continuous(labels = comma_format()) +
  labs(title = "Deaths", x = "Deaths / sub-district", y = "Count") +
  theme_minimal(base_size = 10)

p1 | p2 | p3
```

::: callout-note
## What I Learned

All three metrics are heavily right-skewed — most sub-districts record modest counts but a few hotspot sub-districts hold extreme values. A boxplot compresses this right tail to near-invisibility. A ridgeline plot exposes the full density shape per city; a raincloud makes each individual hotspot pinpointable as a named dot.
:::

### City Comparison — Boxplot Baseline

```{r}
#| fig-height: 4.5

p_b1 <- ggplot(covid,
       aes(x = reorder(city, positive, FUN = median), y = positive, fill = city)) +
  geom_boxplot(alpha = 0.75, outlier.size = 0.8,
               outlier.alpha = 0.5, outlier.colour = "#c0392b") +
  stat_summary(fun = mean, geom = "point", shape = 23,
               size = 3, fill = "white", colour = "black") +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = comma_format()) +
  labs(title = "Positive Cases by City",
       subtitle = "Ordered by median. Red dots = hotspot sub-districts",
       x = NULL, y = "Cases / sub-district") +
  theme_minimal(base_size = 10) + theme(legend.position = "none")

p_b2 <- ggplot(covid_cfr,
       aes(x = reorder(city, cfr, FUN = median), y = cfr, fill = city)) +
  geom_boxplot(alpha = 0.75, outlier.size = 0.8,
               outlier.alpha = 0.5, outlier.colour = "#c0392b") +
  stat_summary(fun = mean, geom = "point", shape = 23,
               size = 3, fill = "white", colour = "black") +
  coord_flip() +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  labs(title = "CFR by City (>= 10 positive)",
       subtitle = "Motivates raincloud to reveal individual sub-districts",
       x = NULL, y = "CFR (%)") +
  theme_minimal(base_size = 10) + theme(legend.position = "none")

p_b1 | p_b2
```

::: callout-note
## What I Learned

The boxplots confirm inter-city differences but hide distribution shape. The long upper whiskers and red outlier dots show that hotspot sub-districts exist — but the boxplot cannot reveal whether those outliers are isolated extreme cases or part of a broader secondary cluster. This directly motivates the ridgeline and raincloud plots below.
:::

:::

---

## Visualising Distribution with Ridgeline Plot

A ridgeline plot reveals distribution **shape** across many groups simultaneously. Most useful when comparing 5+ groups where separate panels waste space.

::: callout-note
**ggridges** provides two main geoms:

- `geom_density_ridges()` — estimates KDE density then draws ridgelines.
- `geom_density_ridges_gradient()` — same but fill colour varies along x.

`bandwidth` controls KDE smoothing and must be **scaled to the data range** — the reference uses `bandwidth = 3.4` for exam scores (range ~60 units); for case counts (range ~6,000) we use `bandwidth = 300`.
:::

### Basic Ridgeline Plot

```{r}
#| fig-height: 6

ggplot(covid,
       aes(x = positive, y = reorder(city, positive, FUN = median))) +
  geom_density_ridges(scale = 3, rel_min_height = 0.01, bandwidth = 300,
                      fill = lighten("#7097BB", .3), colour = "white") +
  scale_x_continuous(name = "Positive cases per sub-district",
                     labels = comma_format(), expand = c(0, 0)) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  labs(title    = "Distribution of Sub-district Positive Cases by City",
       subtitle = "Cities ordered by median. Heavy right-skew visible in all cities") +
  theme_ridges(font_size = 11)
```

::: callout-note
## What I Learned

`bandwidth = 300` scales KDE smoothing to match the data range. `scale = 3` allows moderate ridge overlap keeping all city labels readable. `rel_min_height = 0.01` trims tails below 1% of peak height, preventing invisible ridge "legs" extending along the x-axis. All cities show right-skewed shapes with long hotspot tails.
:::

### Varying Fill Colours Along the X Axis

```{r}
#| fig-height: 6

ggplot(covid,
       aes(x = positive, y = reorder(city, positive, FUN = median),
           fill = after_stat(x))) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01) +
  scale_x_continuous(name = "Positive cases per sub-district",
                     labels = comma_format(), expand = c(0, 0)) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  scale_fill_viridis_c(name = "Cases", option = "C", labels = comma_format()) +
  labs(title    = "Sub-district Cases — Gradient Fill",
       subtitle = "Yellow/orange = hotspot sub-districts. Purple = low-case areas") +
  theme_ridges(font_size = 11, grid = TRUE) +
  theme(legend.position = "right")
```

::: callout-important
`geom_density_ridges_gradient()` does **not** support alpha transparency — gradient fill and `alpha` are mutually exclusive.
:::

::: callout-note
## What I Learned

`after_stat(x)` maps case count to fill colour, adding a second encoding channel that reinforces the x-axis position. Cities with yellow-orange extending further right contain more extreme hotspot sub-districts relative to their city population.
:::

### Mapping ECDF Probabilities onto Colour

```{r}
#| fig-height: 6

ggplot(covid,
       aes(x = positive, y = reorder(city, positive, FUN = median),
           fill = 0.5 - abs(0.5 - after_stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_x_continuous(name = "Positive cases per sub-district",
                     labels = comma_format(), expand = c(0, 0)) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  scale_fill_viridis_c(name = "Tail\nprobability", direction = -1, option = "D") +
  labs(title    = "Sub-district Cases — ECDF Probability Mapped to Colour",
       subtitle = "Bright = typical sub-districts near median. Dark = rare extreme hotspot areas") +
  theme_ridges(font_size = 11, grid = TRUE)
```

::: callout-important
`calc_ecdf = TRUE` is mandatory in `stat_density_ridges()` when using `after_stat(ecdf)`.
:::

::: callout-note
## What I Learned

The ECDF approach answers: "where does the bulk of probability mass sit relative to the median?" Bright = typical sub-district case counts; dark = rare extreme hotspots. The width of the bright band reveals whether sub-district burden is evenly spread or concentrated in a few extreme areas.
:::

### Ridgeline with Quantile Lines

```{r}
#| fig-height: 6

ggplot(covid,
       aes(x = positive, y = reorder(city, positive, FUN = median),
           fill = factor(after_stat(quantile)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE,
                      quantiles = 4, quantile_lines = TRUE) +
  scale_x_continuous(name = "Positive cases per sub-district",
                     labels = comma_format(), expand = c(0, 0)) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  scale_fill_viridis_d(name = "Quartile", alpha = 0.7) +
  labs(title    = "Sub-district Cases — Quartile Colouring",
       subtitle = "Vertical lines = Q1 / Median / Q3. Wide Q4 = concentrated extreme hotspots") +
  theme_ridges(font_size = 11, grid = TRUE)
```

```{r}
#| fig-height: 6

ggplot(covid,
       aes(x = positive, y = reorder(city, positive, FUN = median),
           fill = factor(after_stat(quantile)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE,
                      quantiles = c(0.025, 0.975)) +
  scale_x_continuous(name = "Positive cases per sub-district",
                     labels = comma_format(), expand = c(0, 0)) +
  scale_y_discrete(name = NULL, expand = expansion(add = c(0.2, 2.6))) +
  scale_fill_manual(
    name   = "Probability\nregion",
    values = c("#FF0000A0", "#A0A0A0A0", "#0000FFA0"),
    labels = c("Bottom 2.5%", "Middle 95%", "Top 2.5%")
  ) +
  labs(title    = "Sub-district Cases — 2.5% and 97.5% Tail Highlighting",
       subtitle = "Red = near-zero sub-districts. Blue = extreme top-2.5% hotspot sub-districts") +
  theme_ridges(font_size = 11, grid = TRUE)
```

::: callout-note
## What I Learned

`quantiles = 4` passes the number of equal sections; `quantiles = c(0.025, 0.975)` passes specific probability cut points — syntactically different, producing different outputs. The blue top-2.5% region identifies the most extreme hotspot sub-districts for priority resource allocation.
:::

---

## Visualising Distribution with Raincloud Plot

A raincloud plot is built in four incremental steps following [r4va.netlify.app/chap09](https://r4va.netlify.app/chap09).  
We use **CFR (Case Fatality Rate)** as the y‑variable and compare it across cities.

**Steps 1–3** show the chart in its natural orientation (cities on x‑axis, CFR on y‑axis).  
`coord_flip()` and `theme_economist()` are applied only at the final step to produce the horizontal raincloud silhouette.

### Step 1 — Half Eye: `stat_halfeye()`

```{r}
#| fig-height: 5.5

ggplot(covid_cfr,
       aes(x = reorder(city_label, cfr, FUN = median), y = cfr)) +
  stat_halfeye(
    adjust        = 0.5,
    justification = -0.2,
    .width        = 0,
    point_colour  = NA
  ) +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  labs(title    = "Sub-district CFR by City — Step 1: Half Eye",
       subtitle = ".width = 0 and point_colour = NA suppress the slab interval",
       x = NULL, y = "Case Fatality Rate (%)") +
  theme_bw(base_size = 12)
```

::: callout-note
## What I Learned

`stat_halfeye()` draws a one‑sided density — flat on one side, curved on the other — leaving the opposite side free for the boxplot and dot layers. `adjust = 0.5` uses a narrower KDE bandwidth, appropriate for CFR data (range 0–4%). `.width = 0` and `point_colour = NA` suppress ggdist’s default slab interval and median dot.
:::

### Step 2 — Adding the Boxplot: `geom_boxplot()`

```{r}
#| fig-height: 5.5

ggplot(covid_cfr,
       aes(x = reorder(city_label, cfr, FUN = median), y = cfr)) +
  stat_halfeye(
    adjust        = 0.5,
    justification = -0.2,
    .width        = 0,
    point_colour  = NA
  ) +
  geom_boxplot(width = .20, outlier.shape = NA) +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  labs(title    = "Sub-district CFR by City — Step 2: + Boxplot",
       subtitle = "width = .20. outlier.shape = NA prevents duplicate dots in Step 3",
       x = NULL, y = "Case Fatality Rate (%)") +
  theme_bw(base_size = 12)
```

::: callout-note
## What I Learned

`width = .20` keeps the boxplot compact. `outlier.shape = NA` is essential – without it, boxplot outlier dots AND the dots added by `stat_dots()` in Step 3 would both appear for the same sub‑districts.
:::

### Step 3 — Adding Dot Plot: `stat_dots()`

```{r}
#| fig-height: 5.5
#| fig-width: 14

city_order <- covid_cfr %>%
  group_by(city_label) %>%
  summarise(med_cfr = median(cfr, na.rm = TRUE)) %>%
  arrange(med_cfr) %>%
  pull(city_label)

covid_cfr <- covid_cfr %>%
  mutate(
    city_ordered = factor(city_label, levels = city_order),
    x_pos        = as.numeric(city_ordered) * 20   # ← was 10, now 20
  )

ggplot(covid_cfr, aes(x = x_pos, y = cfr, group = city_ordered)) +
  stat_halfeye(
    adjust        = 0.6,
    justification = -0.08,
    .width        = 0,
    point_colour  = NA,
    width         = 8   # ← caps density at 0.45 × 20 = 9 units rightward
  ) +
  geom_boxplot(
    width         = 1.5,
    outlier.shape = NA
  ) +
  stat_dots(
    side          = "left",
    justification = 3.5,
    binwidth      = 0.2,
    dotsize       = 0.4,
    width         = 0.45   # ← match dots width to halfeye width
  ) +
  scale_x_continuous(
    breaks = unique(covid_cfr$x_pos),
    labels = city_order,
    name   = NULL,
    expand = expansion(mult = c(0.05, 0.05))
  ) +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  labs(
    title = "Sub-district CFR by City — Step 3: + Dots",
    x = NULL, y = "Case Fatality Rate (%)"
  ) +
  theme_bw(base_size = 10)
```

::: callout-note
## What I Learned

`stat_dots()` creates a stacked dotplot (like a histogram that preserves every point). `binwidth = 0.1` groups CFR values into bins of 0.1%. `dotsize = 0.3` keeps dots small and readable. `side = "left"` places the dots on the left side of the boxplot – after flipping, they will hang below (the “rain”).
:::

### Final Raincloud — `coord_flip()` + `theme_economist()`

```{r}
#| fig-height: 6
#| fig-width: 7

city_order <- covid_cfr %>%
  group_by(city_label) %>%
  summarise(med_cfr = median(cfr, na.rm = TRUE)) %>%
  arrange(med_cfr) %>%
  pull(city_label)

covid_cfr <- covid_cfr %>%
  mutate(
    city_ordered = factor(city_label, levels = city_order),
    x_pos        = as.numeric(city_ordered) * 20
  )

ggplot(covid_cfr, aes(x = x_pos, y = cfr, group = city_ordered)) +
  stat_halfeye(
    adjust        = 0.6,
    justification = -0.08,
    .width        = 0,
    point_colour  = NA,
    width         = 7
  ) +
  geom_boxplot(
    width         = 1.2,
    outlier.shape = NA
  ) +
  stat_dots(
    side          = "left",
    justification = 4,
    binwidth      = 0.2,
    dotsize       = 0.2,
    width         = 0.2
  ) +
  coord_flip() +                                    # ← flips to horizontal raincloud
  scale_x_continuous(                               # after flip: this becomes the horizontal CFR axis
    breaks = unique(covid_cfr$x_pos),
    labels = city_order,
    name   = NULL,
    expand = expansion(mult = c(0.05, 0.05))
  ) +
  scale_y_continuous(                               # after flip: this becomes the vertical CFR axis
    labels = function(x) paste0(round(x, 1), "%"),
    name = "Case Fatality Rate (%)"
  ) +
  labs(
    title = "Sub-district CFR by City — Raincloud Plot",
    x = NULL, y = NULL                              # city names are on y‑axis after flip
  ) +
  theme_economist(base_size = 11) +                 # ← professional publishing theme
  theme(
    axis.text.y = element_text(hjust = 1, face = "bold"),
    panel.grid.major.y = element_line(color = "white", linewidth = 0.5),
    panel.grid.major.x = element_blank()
  )
```

::: callout-note
## What I Learned

- **`coord_flip()`** is applied ***only at the final step*** – exactly as in the reference site.  
- The density becomes the “cloud” above the line, the boxplot sits on the line, and the dots become the “rain” below.  
- `theme_economist()` gives a professional, publication‑ready look (light blue background, clean grid lines).  
- The plot reveals that **Kepulauan Seribu** has very different CFR characteristics: a few isolated, extreme dots far from the main cluster. Jakarta Timur shows the widest spread, while Jakarta Pusat is tightly clustered at low CFR values.
:::

---

## Interactive Exploration

The raincloud plots reveal distribution patterns; the interactive charts below identify *which specific sub-district* each dot represents.

### Interactive CFR by Sub-district

```{r}
#| fig-height: 5.5

hover_cfr <- covid_cfr %>%
  mutate(tip = paste0(
    "<b>", subdistrict, "</b><br>",
    "District: ",  district, "<br>",
    "City: ",      city, "<br>",
    "Positive: ",  comma(positive), "<br>",
    "Deaths: ",    death, "<br>",
    "CFR: ",       round(cfr, 2), "%"
  ))

plot_ly() %>%
  add_trace(
    data = hover_cfr, x = ~cfr,
    y    = ~reorder(city, cfr, FUN = median),
    type = "box", hoverinfo = "none",
    fillcolor = "rgba(200,200,200,0.3)",
    line = list(color = "grey50"), showlegend = FALSE
  ) %>%
  add_trace(
    data   = hover_cfr, x = ~cfr,
    y      = ~reorder(city, cfr, FUN = median),
    type   = "scatter", mode = "markers",
    marker = list(size = 5, opacity = 0.65, color = "#e74c3c",
                  line = list(color = "white", width = 0.5)),
    text = ~tip, hoverinfo = "text", showlegend = FALSE
  ) %>%
  layout(
    title = list(
      text = "<b>Sub-district CFR by City — Interactive</b><br><sup>Hover dots to identify each sub-district: name, district, cases, deaths, CFR</sup>",
      x = 0.5
    ),
    xaxis = list(title = "Case Fatality Rate (%)", ticksuffix = "%"),
    yaxis = list(title = ""),
    plot_bgcolor  = "rgba(0,0,0,0)",
    paper_bgcolor = "rgba(0,0,0,0)"
  )
```

### Interactive Positive Cases by Sub-district

```{r}
#| fig-height: 5.5

hover_pos <- covid %>%
  mutate(tip = paste0(
    "<b>", subdistrict, "</b><br>",
    "District: ",  district, "<br>",
    "City: ",      city, "<br>",
    "Positive: ",  comma(positive), "<br>",
    "Recovered: ", comma(recovered), "<br>",
    "Deaths: ",    death, "<br>",
    "Active: ",    comma(active)
  ))

plot_ly() %>%
  add_trace(
    data = hover_pos, x = ~positive,
    y    = ~reorder(city, positive, FUN = median),
    type = "box", hoverinfo = "none",
    fillcolor = "rgba(200,200,200,0.3)",
    line = list(color = "grey50"), showlegend = FALSE
  ) %>%
  add_trace(
    data   = hover_pos, x = ~positive,
    y      = ~reorder(city, positive, FUN = median),
    type   = "scatter", mode = "markers",
    marker = list(size = 5, opacity = 0.6, color = "#3498db",
                  line = list(color = "white", width = 0.5)),
    text = ~tip, hoverinfo = "text", showlegend = FALSE
  ) %>%
  layout(
    title = list(
      text = "<b>Sub-district Positive Cases by City — Interactive</b><br><sup>Hover to identify hotspot sub-districts: name, cases, recoveries, deaths, active</sup>",
      x = 0.5
    ),
    xaxis = list(title = "Positive cases per sub-district", tickformat = ","),
    yaxis = list(title = ""),
    plot_bgcolor  = "rgba(0,0,0,0)",
    paper_bgcolor = "rgba(0,0,0,0)"
  )
```

::: callout-note
## What I Learned

The static raincloud shows *where* extreme dots are; the interactive chart names *which* sub‑districts they belong to. Hovering reveals the sub‑district name, district, city, and all case metrics – bridging distributional understanding with operational intelligence.
:::

---

## Summary of Techniques

```{r}
#| echo: false
tibble(
  `#` = 1:9,
  Technique = c(
    "Basic ridgeline", "Gradient fill ridgeline", "ECDF ridgeline",
    "Quartile ridgeline", "Tail cut-point ridgeline",
    "Step 1-3: Building raincloud (un-flipped)",
    "Final raincloud — CFR (flipped + theme_economist)",
    "Interactive CFR", "Interactive Positive Cases"
  ),
  `Key function / argument` = c(
    "geom_density_ridges(bandwidth = 300)",
    "geom_density_ridges_gradient() + after_stat(x) + viridis",
    "stat_density_ridges(calc_ecdf = TRUE) + after_stat(ecdf)",
    "stat_density_ridges(quantiles = 4, quantile_lines = TRUE)",
    "stat_density_ridges(quantiles = c(0.025, 0.975))",
    "stat_halfeye(.width=0) + geom_boxplot(outlier.shape=NA) + stat_dots(side='left')",
    "coord_flip() + theme_economist(); adjust=0.5, binwidth=0.1, dotsize=0.3",
    "plot_ly() box + scatter with hover text",
    "plot_ly() box + scatter with hover text"
  ),
  `Key insight for Jakarta COVID data` = c(
    "All cities right-skewed; hotspot tails differ in extent across cities",
    "Warm gradient regions confirm cities with most extreme hotspot sub-districts",
    "Narrow bright = typical; dark tail = rare extremes needing resources",
    "Wide Q4 confirms top 25% of sub-districts carry disproportionate burden",
    "Blue top-2.5% identifies the most extreme hotspot sub-districts per city",
    "Each layer added separately shows function and role of each raincloud element",
    "Kepulauan Seribu CFR is distinct — extreme isolated outlier dots visible",
    "Sub-district name, district, and all case metrics visible on hover",
    "Identifies which sub-districts drive the positive-case distribution"
  )
) %>%
  kbl(caption = "Visualising Distribution — Techniques and Insights Summary",
      align = c("c","l","l","l")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  row_spec(1:5, background = "#eaf4fb") %>%
  row_spec(6:9, background = "#fef9e7")
```

---

## Final Remarks

This exercise demonstrated the full range of modern distribution visualisation: from ridgeline plots (with gradient, ECDF, and quantile variants) to raincloud plots built step by step exactly as in the reference chapter. By applying these techniques to the DKI Jakarta COVID‑19 data, we uncovered that all cities have right‑skewed positive‑case distributions, that CFR varies substantially across cities, and that Kepulauan Seribu behaves very differently from the mainland. Interactive plots then allowed us to identify the specific sub‑districts driving those patterns. Together, these tools provide a complete toolkit for exploring and communicating distributions in public health data.


## References

- Wilke, C. O. *R for Visual Analytics*, [Chapter 9: Visualising Distribution](https://r4va.netlify.app/chap09).
- Wilke, C. O. (2019). [*Fundamentals of Data Visualization*](https://clauswilke.com/dataviz/).
- Kay, M. (2023). [ggdist: Visualizations of Distributions and Uncertainty](https://mjskay.github.io/ggdist/).
- Allen, M. et al. (2021). [Raincloud plots](https://wellcomeopenresearch.org/articles/4-63).
