---
title: "Hands-on Exercise 4d"
subtitle: "Visualising Models — EDA, Diagnostics, Parameters & Coefficients"
author: "Thi Thanh Van Tran"
date: "May 3, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    number-depth: 3
    code-fold: true
    code-summary: "Show code"
    code-tools: false
    code-copy: true
    css: styles.css
    theme: cosmo
    page-layout: article
execute:
  eval: true
  echo: true
  warning: false
  message: false
  freeze: false
knitr:
  opts_chunk:
    out.width: "100%"
    fig.width: 10
    fig.align: "center"
    dpi: 120
---

```{=html}
<style>
  .quarto-figure, figure {
    max-width: 100% !important; width: 100% !important;
    margin: 0.5rem auto !important; overflow: visible !important;
  }
  img.figure-img, .cell > img, .cell-output > img, .cell-output-display img {
    width: 100% !important; height: auto !important;
    object-fit: contain !important; display: block; margin: 0.3rem auto;
  }
  .cell-output-display {
    margin: 0.4rem 0 !important; padding: 0 !important; overflow: visible !important;
  }
  .callout { margin-top: 0.75rem; }
  h2 { margin-top: 2rem; }
</style>
```

---

## Introduction

This exercise follows the workflow in [Chapter 12 of R for Visual Analytics](https://r4va.netlify.app/chap12), adapted to the **DKI Jakarta COVID-19 sub-district data**. The goal is to build a multiple regression model explaining **Case Fatality Rate (CFR)** across sub-districts, then use the `easystats` and `ggstatsplot` toolchain to interrogate and communicate that model visually.

The workflow is deliberately **sequential and diagnostic**: each step either surfaces a problem that must be fixed (collinearity, violated assumptions) or confirms something that allows us to move forward. Nothing is skipped, and nothing is plotted for its own sake.

---

## Visual Analytics for Building Better Explanatory Models

### The Case

The Jakarta COVID-19 dataset contains sub-district-level counts of positive cases, recoveries, deaths, and active cases across five cities (Kepulauan Seribu is excluded due to insufficient observations). The target variable is:

$$\text{CFR} = \frac{\text{death}}{\text{positive}} \times 100$$

The question we want to answer visually: **which sub-district characteristics — case volume, recovery dynamics, or geography — drive CFR variation, and how confidently can we say so?**

This is an *explanatory* modelling exercise, not predictive. The goal is to understand the structure, not forecast future CFR.

---

## Getting Started

### Installing and Loading Packages

```{r}
#| code-fold: false
pacman::p_load(
  readxl,       # reading Excel files (as in chapter)
  SmartEDA,     # ExpCatStat() for variable screening
  tidyverse,    # data wrangling + ggplot2
  ggstatsplot,  # ggcoefstats() for annotated coefficient plots
  easystats,    # performance, see, parameters
  tidymodels,   # modern modelling framework
  broom,        # tidy() / glance() for model summaries as data frames
  janitor,      # clean_names()
  scales,
  knitr,
  kableExtra
)

# Important: tidymodels loads dials, which exports its own parameters() function.
# Always call parameters::parameters() explicitly to avoid a silent namespace conflict.
```

### Importing and Preparing Data

```{r}
#| code-fold: false
covid_raw <- read_csv("data/COVID-19_DKI_Jakarta.csv")

covid <- covid_raw %>%
  janitor::clean_names() %>%
  rename(id = sub_district_id, subdistrict = sub_district) %>%
  mutate(
    cfr           = if_else(positive > 0, death / positive * 100, NA_real_),
    recovery_rate = if_else(positive > 0, recovered / positive * 100, NA_real_),
    active        = positive - recovered - death,
    city_label = case_when(
      str_detect(toupper(city), "SELATAN") ~ "Jkt. Selatan",
      str_detect(toupper(city), "UTARA")   ~ "Jkt. Utara",
      str_detect(toupper(city), "BARAT")   ~ "Jkt. Barat",
      str_detect(toupper(city), "PUSAT")   ~ "Jkt. Pusat",
      str_detect(toupper(city), "TIMUR")   ~ "Jkt. Timur",
      str_detect(toupper(city), "SERIBU")  ~ "Kep. Seribu",
      TRUE ~ city
    )
  )

# Sub-districts with reliable CFR; exclude Kepulauan Seribu (n = 6, too few)
covid_model <- covid %>%
  filter(!is.na(cfr), positive >= 10,
         city != "KAB. ADM. KEP. SERIBU")
```

---

## Visualising Modelling Variables with `SmartEDA`

Before fitting any model, `ExpCatStat()` from **SmartEDA** screens all candidate variables for association with the target using **Cramér's V** and a chi-squared test. Because `ExpCatStat()` requires a categorical target, we bin `cfr` into three groups: **Low** (< 1 %), **Medium** (1–2 %), **High** (> 2 %).

```{r}
#| code-fold: false
covid_eda <- covid_model %>%
  mutate(
    cfr_bin       = cut(cfr,       breaks = c(-Inf, 1, 2, Inf),
                        labels = c("Low","Medium","High")),
    positive_bin  = cut(positive,  breaks = quantile(positive,  c(0,.33,.67,1), na.rm=TRUE),
                        labels = c("Low","Medium","High"), include.lowest = TRUE),
    recovered_bin = cut(recovered, breaks = quantile(recovered, c(0,.33,.67,1), na.rm=TRUE),
                        labels = c("Low","Medium","High"), include.lowest = TRUE),
    death_bin     = cut(death,     breaks = quantile(death,     c(0,.33,.67,1), na.rm=TRUE),
                        labels = c("Low","Medium","High"), include.lowest = TRUE),
    active_bin    = cut(active,    breaks = quantile(active,    c(0,.33,.67,1), na.rm=TRUE),
                        labels = c("Low","Medium","High"), include.lowest = TRUE)
  ) %>%
  select(cfr_bin, positive_bin, recovered_bin, death_bin, active_bin, city_label)

ExpCatStat(covid_eda,
           Target  = "cfr_bin",
           result  = "Stat",
           clim    = 10,
           nlim    = 5,
           bins    = 10,
           Pclass  = "Medium") %>%
  arrange(desc(`Cramers V`)) %>%
  select(Variable, `Chi-squared`, `p-value`, `Cramers V`,
         `Degree of Association`, `Predictive Power`) %>%
  kbl(digits  = 3,
      caption = "Variable association with CFR group — sorted by Cramér's V (descending)",
      align   = c("l","r","r","r","l","l")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  row_spec(1, background = "#fde8e8", bold = TRUE)
```

::: callout-note
## What I Learned

- `death_bin` produces the highest Cramér's V — expected, since CFR is arithmetically *constructed* from death. This is a red flag: including `death` as a raw predictor would create a circular model, not an explanatory one.
- `city_label` shows moderate-to-strong association, confirming that **geography adds information beyond case counts** — cities differ systematically in CFR even among sub-districts with similar caseloads.
- The screening output is a pre-model sanity check: it tells us both *what to include* and *what to watch out for*. `death_bin`'s near-perfect association signals collinearity before any `lm()` is ever called.
:::

---

## Multiple Regression Model using `lm()`

### Initial Model — Including the Collinear Predictor

Following the chapter, we first build a deliberately "full" model that includes `death` — the variable flagged as circular by the EDA. This is intentional: we want to see what the diagnostics reveal when we break model assumptions.

```{r}
#| code-fold: false
model_full <- lm(cfr ~ positive + recovered + death + active + city_label,
                 data = covid_model)

# Coefficient table
broom::tidy(model_full) %>%
  mutate(across(where(is.numeric), ~round(.x, 4)),
         significance = case_when(
           p.value < 0.001 ~ "***",
           p.value < 0.01  ~ "**",
           p.value < 0.05  ~ "*",
           p.value < 0.1   ~ ".",
           TRUE            ~ ""
         )) %>%
  kbl(caption = "Initial model coefficients (includes `death` — collinear)",
      col.names = c("Term","Estimate","Std. Error","t statistic","p-value",""),
      align = c("l","r","r","r","r","c")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white")

# Model fit summary
broom::glance(model_full) %>%
  select(r.squared, adj.r.squared, sigma, statistic, p.value, df, nobs) %>%
  mutate(across(where(is.numeric), ~round(.x, 4))) %>%
  kbl(caption = "Model fit statistics",
      col.names = c("R²","Adj. R²","Residual SE","F statistic","p-value","df","n"),
      align = "c") %>%
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white")
```

::: callout-warning
## What I Learned

- The model summary shows high R² and many significant coefficients — which looks good on the surface but is **misleading**. When a predictor is arithmetically entangled with the outcome, the model fits numerically but estimates are unstable and uninterpretable.
- This mirrors the Toyota Corolla chapter exactly: `Age_08_04` and `Mfg_Year` produced an apparent fit while hiding severe collinearity. The lesson is that R² alone cannot flag a broken model — only diagnostics can.
:::

---

## Model Diagnostics

### Checking Multicollinearity

`check_collinearity()` from **performance** computes the **Variance Inflation Factor (VIF)** for every predictor. A VIF > 10 signals problematic collinearity; VIF > 5 warrants investigation.

```{r}
#| code-fold: false
check_collinearity(model_full) %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~round(.x, 2))) %>%
  arrange(desc(VIF)) %>%
  kbl(caption = "VIF table — initial model (includes `death`)",
      align   = c("l","r","r","r","r","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  row_spec(which(as.data.frame(check_collinearity(model_full)) %>%
                   arrange(desc(VIF)) %>% pull(VIF) > 10),
           background = "#fde8e8", bold = TRUE)
```

```{r}
#| fig-height: 5
check_c_full <- check_collinearity(model_full)
plot(check_c_full)
```

::: callout-important
## What I Learned

- `death` has VIF well above 10 — confirming the circularity flagged in EDA. Its standard error is massively inflated, making its coefficient estimate meaningless.
- The other predictors (`positive`, `recovered`, `active`) also show elevated VIFs because they are all inflated *by* `death`'s presence — removing one collinear term often rescues the rest.
- **Decision rule from this plot**: any bar in the red zone → remove that predictor and refit before trusting any coefficient.
:::

---

### Reduced Model — Removing the Collinear Predictor

Following the same logic as the chapter (dropping `Mfg_Year` to resolve `Age_08_04` collinearity), we drop `death` and refit.

```{r}
#| code-fold: false
model1 <- lm(cfr ~ positive + recovered + active + city_label,
             data = covid_model)

# Coefficient table
broom::tidy(model1) %>%
  mutate(across(where(is.numeric), ~round(.x, 4)),
         significance = case_when(
           p.value < 0.001 ~ "***",
           p.value < 0.01  ~ "**",
           p.value < 0.05  ~ "*",
           p.value < 0.1   ~ ".",
           TRUE            ~ ""
         )) %>%
  kbl(caption = "Reduced model coefficients (`death` removed)",
      col.names = c("Term","Estimate","Std. Error","t statistic","p-value",""),
      align = c("l","r","r","r","r","c")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  row_spec(which(broom::tidy(model1)$p.value < 0.05) , background = "#eafbea")

# Model fit summary
broom::glance(model1) %>%
  select(r.squared, adj.r.squared, sigma, statistic, p.value, df, nobs) %>%
  mutate(across(where(is.numeric), ~round(.x, 4))) %>%
  kbl(caption = "Model fit statistics — reduced model",
      col.names = c("R²","Adj. R²","Residual SE","F statistic","p-value","df","n"),
      align = "c") %>%
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white")
```

```{r}
#| fig-height: 5
check_c <- check_collinearity(model1)

# Table
check_c %>%
  as.data.frame() %>%
  mutate(across(where(is.numeric), ~round(.x, 2))) %>%
  arrange(desc(VIF)) %>%
  kbl(caption = "VIF table — reduced model (all predictors green)",
      align   = c("l","r","r","r","r","r","r")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white")
```

```{r}
#| fig-height: 5
# Visual
plot(check_c)
```

::: callout-note
## What I Learned

- All VIFs fall below 5 after removing `death`. The bars are now uniformly green — the model is structurally sound.
- Crucially, the R² barely changes from `model_full` to `model1`. This confirms that `death` was not adding *new* information — it was merely duplicating what `positive` and `recovered` already encode.
:::

---

### Checking Normality of Residuals

`check_normality()` overlays the empirical residual distribution against the theoretical normal curve, and produces a Q-Q plot for tail behaviour.

```{r}
#| fig-height: 5
check_n <- check_normality(model1)
plot(check_n)
```

::: callout-note
## What I Learned

- Residuals are approximately normal in the centre but show mild heavy tails — common for rate outcomes bounded at zero (CFR cannot go negative). This level of deviation is acceptable at our sample size (n ≈ 260) via the Central Limit Theorem.
- If the tails were severely non-normal, the next step would be a Box-Cox transformation of CFR — but that is not needed here.
:::

---

### Checking Homogeneity of Variances

`check_heteroscedasticity()` plots √|standardised residuals| against fitted values. A flat, horizontal band indicates constant variance.

```{r}
#| fig-height: 5
check_h <- check_heteroscedasticity(model1)
plot(check_h)
```

::: callout-note
## What I Learned

- The reference line is broadly flat — no strong funnel pattern, so homoscedasticity holds reasonably well.
- Slight spread at the extremes corresponds to a handful of sub-districts with very high CFR. These are real-world anomalies, not data errors — retaining them is appropriate.
:::

---

### Complete Diagnostic Check — `check_model()`

`check_model()` consolidates all individual diagnostics into a single six-panel dashboard. This is the most efficient way to audit all linear regression assumptions simultaneously.

```{r}
#| fig-height: 10
check_model(model1)
```

::: callout-note
## What I Learned

The six panels address six distinct questions — each one either clears a concern or flags it:

| Panel | Question | Finding |
|---|---|---|
| Posterior predictive check | Does the model reproduce the observed distribution? | Good overlap — model captures the bulk of CFR values |
| Linearity (Residuals vs Fitted) | Is the outcome-predictor relationship truly linear? | Reference line flat → linearity holds |
| Homoscedasticity | Is residual spread constant across fitted values? | Broadly yes; minor spread at extremes |
| Influential observations (Cook's D) | Do any sub-districts distort the whole model? | A few points exceed threshold — but represent real variation |
| Collinearity (VIF) | Are any predictors redundant? | All green after removing `death` |
| Normality of residuals (Q-Q) | Are residuals Gaussian? | Acceptable; slight heavy tails |

The **influential observations** panel is the one most often overlooked. High Cook's Distance means a single sub-district, if removed, would shift coefficients noticeably. Here those points are genuine high-CFR cases — not data errors — so we retain them and note their influence in interpretation.
:::

---

## Visualising Regression Parameters

With all diagnostics cleared, we can now interpret the model coefficients with confidence. Two complementary methods are demonstrated.

### Method 1: `plot(parameters::parameters())` — `see` Package

`parameters()` from the **easystats** ecosystem extracts model parameters; `plot()` from **see** renders them as a dot-and-whisker coefficient plot. The explicit namespace `parameters::parameters()` is mandatory here because `tidymodels` loads `dials`, which exports its own `parameters()` for hyperparameter tuning.

```{r}
#| fig-height: 5
plot(parameters::parameters(model1), show_intercept = FALSE) +
  labs(
    title    = "Regression Coefficients — CFR Model (DKI Jakarta)",
    subtitle = "Point estimates with 95% confidence intervals"
  ) +
  theme_minimal(base_size = 12)
```

::: callout-note
## What I Learned

- **`recovered` has the largest negative coefficient** — sub-districts where more cases resolve into recoveries have systematically lower CFR. This is the single most important driver in the model and is analytically coherent: recovery dynamics reflect healthcare capacity.
- **City effects are statistically significant and substantial** — even after conditioning on case volume and recovery, geography explains residual CFR differences. This points to unmeasured city-level factors: healthcare infrastructure, testing rates, demographic composition.
- **`positive` sits near zero** — once recovery and active case load are in the model, total confirmed cases carry minimal marginal signal. Volume alone does not predict fatality.
- A coefficient whose CI **does not cross zero** is statistically distinguishable from no effect. The `see` plot makes this visible at a glance without reading a table.
:::

---

### Method 2: `ggcoefstats()` — `ggstatsplot` Package

`ggcoefstats()` embeds coefficient estimates, t-statistics, and p-value annotations directly on the plot — making it self-contained for reports and presentations.

```{r}
#| fig-height: 10
ggcoefstats(
  x            = model1,
  output       = "plot",
  title        = "Regression Coefficients — CFR Model (DKI Jakarta)",
  subtitle     = "Unstandardised estimates with 95% CIs and p-value annotations",
  point.shape  = 15,
  point.color  = "white",
  point.fill   = "#e74c3c",
  stats.labels = TRUE
) +
  theme(
    axis.text.y = element_text(margin = margin(t = 4, b = 4)),
    plot.margin = margin(t = 10, b = 10)
  )
```

::: callout-note
## What I Learned

- `ggcoefstats()` shows **unstandardised (raw) coefficients** in the original unit (% CFR), while `plot(parameters::parameters())` shows standardised ones. Both are useful — standardised for comparing relative importance across predictors, unstandardised for direct real-world interpretation.
- The embedded p-value labels and estimate annotations make the plot stand alone without a companion table — practical for policy-facing reports where readers cannot look up a regression table.
- Together, both methods confirm the same story: **recovery rate and city are the dominant structural drivers of CFR**. Raw case counts, once composition is controlled for, fade into statistical irrelevance.
:::

---

## Summary

The full workflow — from variable screening to parameter communication — follows a single logical thread:

```{r}
#| echo: false
tibble(
  `#` = 1:9,
  Step = c(
    "Variable screening",
    "Initial model",
    "Multicollinearity check",
    "Reduced model",
    "Normality of residuals",
    "Homoscedasticity check",
    "Complete diagnostic",
    "Coefficient plot (see)",
    "Coefficient plot (ggcoefstats)"
  ),
  `Key function` = c(
    "SmartEDA::ExpCatStat()",
    "lm(cfr ~ positive + recovered + death + active + city_label)",
    "performance::check_collinearity() · plot()",
    "lm(cfr ~ positive + recovered + active + city_label)",
    "performance::check_normality() · plot()",
    "performance::check_heteroscedasticity() · plot()",
    "performance::check_model()",
    "plot(parameters::parameters(model1))",
    "ggstatsplot::ggcoefstats(model1, output = 'plot')"
  ),
  `What it revealed` = c(
    "death_bin has the highest Cramér's V — circularity risk before any model is fit.",
    "R² looks fine; estimates are not. Surface fit masks structural instability.",
    "death VIF >> 10. Removal required before any coefficient can be trusted.",
    "All VIF < 5. Model is structurally sound. R² barely changes — death added no real information.",
    "Residuals approximately normal; mild heavy tails acceptable at n ≈ 260.",
    "Variance broadly constant; minor spread at extremes from genuine high-CFR outliers.",
    "All six panels pass. A few influential sub-districts noted but retained as real variation.",
    "recovered has the largest significant negative effect; city effects persist after case adjustment.",
    "Same story with p-values and raw estimates embedded — ready for non-technical audiences."
  )
) %>%
  kbl(caption = "Visualising Models — Step-by-Step Summary",
      align = c("c","l","l","l")) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  row_spec(seq(1, 9, 2), background = "#eaf4fb")
```

**The story in one paragraph:** The `ExpCatStat()` EDA signalled before any model was fit that `death` would create circularity — CFR is constructed from it. The initial model confirmed this: `check_collinearity()` showed extreme VIFs. Dropping `death` and refitting gave a structurally clean model that passed all three diagnostic checks. Visualising the parameters then revealed the substantive finding: across Jakarta sub-districts, **where a patient is located (city) and whether they recover (recovered) matter far more for fatality than how many people were infected (positive)**. Visualising models is not just validation — it is the mechanism by which data tells its story, and by which a decision-maker can trust the answer.
