---
title: "Hands-on Exercise 4b"
subtitle: "Visual Statistical Analysis — ggstatsplot, performance & parameters"
author: "Thi Thanh Van Tran"
date: "May 3, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    number-depth: 3
    code-fold: true
    code-summary: "Show code"
    code-tools: false
    code-copy: true
    css: styles.css
    theme: cosmo
    page-layout: article
execute:
  eval: true
  echo: true
  warning: false
  message: false
  freeze: false
knitr:
  opts_chunk:
    out.width: "100%"
    fig.width: 10
    fig.align: "center"
    dpi: 120
---

```{=html}
<style>
  .quarto-figure, figure {
    max-width: 100% !important; width: 100% !important;
    margin: 0.5rem auto !important; overflow: visible !important;
  }
  img.figure-img, .cell > img, .cell-output > img, .cell-output-display img {
    width: 100% !important; height: auto !important;
    object-fit: contain !important; display: block; margin: 0.3rem auto;
  }
  .cell-output-display {
    margin: 0.4rem 0 !important; padding: 0 !important; overflow: visible !important;
  }
  .callout { margin-top: 0.75rem; }
  h2 { margin-top: 2rem; }
</style>
```

------------------------------------------------------------------------

::: welcome-section
# Visual Statistical Analysis — ggstatsplot, performance & parameters

Welcome! This exercise applies **visual statistical analysis** to **COVID-19 sub-district data for DKI Jakarta**, telling a data story from exploration through modelling. You will learn to:

-   Use **`ggstatsplot`** to create publication-ready plots with embedded statistical tests (one-sample, two-sample, ANOVA, correlation, association) that follow the APA reporting standard.
-   Understand and interpret **Bayes Factors** as evidence weights rather than binary p-values.
-   Use **`performance`** to diagnose linear model assumptions visually.
-   Use **`parameters`** + **`see`** to visualise and compare model coefficients.

Each technique builds on the last, progressively answering: *Where in Jakarta did COVID-19 hit hardest, why do sub-districts differ, and can a simple model explain case counts?*

:::

------------------------------------------------------------------------

## What is `ggstatsplot`?

[**`ggstatsplot`**](https://indrajeetpatil.github.io/ggstatsplot/index.html) is an extension of `ggplot2` that embeds statistical test results directly inside plots.

Two design goals distinguish it from plain `ggplot2`:

-   **Alternative inference by default** — it offers Bayesian, robust, and non-parametric options alongside frequentist tests, making it easy to choose the right method for the data.
-   **APA-compliant reporting** — every plot subtitle automatically formats test statistics, degrees of freedom, p-values, and effect sizes according to the APA gold standard. The example below shows the APA template for a robust t-test:

> *t*(df) = value, *p* = value, *d* = value, 95% CI \[lower, upper\]

This means you get publication-ready output in one function call, rather than running a test and a plot separately.

------------------------------------------------------------------------

## Getting Started

### Installing and Loading Packages

```{r}
#| code-fold: false
pacman::p_load(ggstatsplot, performance, parameters, see, patchwork, tidyverse, scales, plotly, janitor, kableExtra, BayesFactor)
```

### Importing and Preparing Data

```{r}
#| code-fold: false
covid_raw <- read_csv("data/COVID-19_DKI_Jakarta.csv")

covid <- covid_raw %>%
  janitor::clean_names() %>%
  rename(id = sub_district_id, subdistrict = sub_district) %>%
  mutate(
    cfr           = if_else(positive > 0, death / positive * 100,     NA_real_),
    recovery_rate = if_else(positive > 0, recovered / positive * 100, NA_real_),
    active        = positive - recovered - death,
    city_label = case_when(
      str_detect(toupper(city), "SELATAN") ~ "Jkt.\nSelatan",
      str_detect(toupper(city), "UTARA")   ~ "Jkt.\nUtara",
      str_detect(toupper(city), "BARAT")   ~ "Jkt.\nBarat",
      str_detect(toupper(city), "PUSAT")   ~ "Jkt.\nPusat",
      str_detect(toupper(city), "TIMUR")   ~ "Jkt.\nTimur",
      str_detect(toupper(city), "SERIBU")  ~ "Kep.\nSeribu",
      TRUE ~ city
    )
  )

# Sub-districts with reliable CFR (>= 10 positive cases)
covid_cfr <- covid %>% filter(!is.na(cfr), positive >= 10)
```

------------------------------------------------------------------------

## Data Exploration

EDA is not just a warm-up — it directly motivates every statistical test that follows. The questions it surfaces become the hypotheses we test.

::: panel-tabset
### Variables Overview

```{r}
#| fig-height: 3.5
summary_stats <- covid_cfr %>%
  summarise(
    `Positive cases` = list(positive),
    `CFR (%)` = list(cfr),
    `Recovery rate (%)` = list(recovery_rate)
  ) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Values") %>%
  mutate(
    Mean   = map_dbl(Values, mean,   na.rm = TRUE),
    Median = map_dbl(Values, median, na.rm = TRUE),
    SD     = map_dbl(Values, sd,     na.rm = TRUE),
    Min    = map_dbl(Values, min,    na.rm = TRUE),
    Max    = map_dbl(Values, max,    na.rm = TRUE)
  ) %>%
  select(-Values)

summary_stats %>%
  kbl(caption = "Summary of key variables used in statistical tests",
      digits = 2, align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

### Missing Data & Filtering

```{r}
#| fig-height: 3.5
missing_ct <- tibble(
  Variable = c("cfr", "recovery_rate", "positive", "death"),
  Missing  = c(sum(is.na(covid$cfr)), sum(is.na(covid$recovery_rate)),
               sum(is.na(covid$positive)), sum(is.na(covid$death)))
)

ggplot(missing_ct, aes(x = reorder(Variable, -Missing), y = Missing)) +
  geom_col(fill = "#e74c3c", alpha = 0.8) +
  geom_text(aes(label = Missing), vjust = -0.5, fontface = "bold") +
  labs(title = "Missing values in original data",
       subtitle = "cfr / recovery_rate missing where positive = 0",
       x = NULL, y = "Count") +
  theme_minimal(base_size = 11)
```
:::

::: callout-note
## What EDA Tells Us — and What It Asks Next

-   `cfr` and `recovery_rate` are missing only for sub-districts with zero positive cases (division-by-zero prevented). Filtering to `positive >= 10` removes unstable CFR estimates — this filtered dataset (`covid_cfr`) is used for all subsequent tests.
-   The mean CFR (\~2.5%) sits noticeably above a plausible global benchmark of 2%. **Is this difference real, or sampling noise?** → answered by the one-sample test.
-   CFR varies widely across sub-districts (SD \> mean in some cities). **Are city-level differences real?** → answered by the two-sample test and ANOVA.
-   Deaths and positive cases both have extreme right-tail skew. **How strong is the linear relationship?** → answered by the correlation test.
-   Sub-districts cluster unevenly into CFR severity categories by city. **Is the association statistically significant?** → answered by the Chi-square association test.

EDA opens the questions; the statistical tests answer them.
:::

------------------------------------------------------------------------

## One-Sample Test: Is the Mean CFR Different from 2%?

**Question:** The mean CFR is \~2.5% in the data. Is that difference from a 2% global benchmark real, or within the range of chance?

We use `gghistostats()` with a Bayesian one-sample test. Rather than a binary reject/fail-to-reject, the Bayes Factor (BF₁₀) tells us *how strongly* the data support the alternative.

```{r}
#| fig-height: 6
set.seed(1234)
gghistostats(
  data       = covid_cfr,
  x          = cfr,
  type       = "bayes",
  test.value = 2,
  bf.prior   = 0.707,
  xlab       = "Case Fatality Rate (%)",
  title      = "One-Sample Test: Mean CFR vs 2%",
  caption    = "BF₁₀ > 10 = strong evidence that mean CFR ≠ 2%"
)
```

### Unpacking the Bayes Factor

A **Bayes factor (BF₁₀)** is the ratio of the likelihood of the data under the alternative hypothesis (H₁: mean ≠ 2%) to the likelihood under the null (H₀: mean = 2%):

$$BF_{10} = \frac{P(\text{data} \mid H_1)}{P(\text{data} \mid H_0)}$$

-   BF₁₀ = 5 means the data are **5× more likely** under H₁ than H₀.
-   BF₁₀ = 1/5 = 0.2 means the data are **5× more likely** under H₀ — evidence *for* the null.
-   Unlike a p-value, it directly quantifies the *weight of evidence*, not just whether we cross a threshold.

The **Schwarz criterion (BIC)** provides a computationally efficient approximation of BF₁₀ and is what `ggstatsplot` uses internally.

### How to Interpret the Bayes Factor

| BF₁₀       | Interpretation                    |
|------------|-----------------------------------|
| \> 100     | Extreme evidence for H₁           |
| 30 – 100   | Very strong evidence for H₁       |
| 10 – 30    | Strong evidence for H₁            |
| 3 – 10     | Moderate evidence for H₁          |
| 1 – 3      | Anecdotal evidence for H₁         |
| 1          | No evidence either way            |
| 1/3 – 1    | Anecdotal evidence for H₀         |
| 1/10 – 1/3 | Moderate evidence for H₀          |
| \< 1/10    | Strong-to-extreme evidence for H₀ |

*Scale adapted from Jeffreys (1961), modified by Lee & Wagenmakers (2013).*

::: callout-note
## What I Learned

The Bayes Factor makes the test result interpretable on a continuous scale. For Jakarta, a BF₁₀ \> 10 would indicate strong evidence that the mean CFR genuinely exceeds 2%, flagging the city as a higher-risk environment than a global baseline — which immediately justifies the city-level breakdown that follows.
:::

------------------------------------------------------------------------

## Two-Sample Test: CFR in Jakarta Selatan vs Jakarta Utara

**Question:** EDA showed both cities have relatively high and spread CFR distributions. Are they actually different from each other?

We use `ggbetweenstats()` with `type = "np"` (non-parametric Mann–Whitney U), which is robust to the outliers visible in EDA.

```{r}
#| fig-height: 6
covid_two_cities <- covid_cfr %>%
  filter(city %in% c("JAKARTA SELATAN", "JAKARTA UTARA")) %>%
  mutate(city = factor(city, levels = c("JAKARTA SELATAN", "JAKARTA UTARA")))

ggbetweenstats(
  data                = covid_two_cities,
  x                   = city,
  y                   = cfr,
  type                = "np",
  var.equal           = FALSE,
  pairwise.display    = "all",
  centrality.plotting = FALSE,
  messages            = FALSE,
  title               = "CFR Comparison: Jakarta Selatan vs Jakarta Utara",
  xlab                = "City",
  ylab                = "Case Fatality Rate (%)"
)
```

::: callout-note
## What I Learned

The **raincloud + boxplot + half-violin** display reveals distribution shape, spread, and outliers simultaneously — much richer than a simple bar chart. The plot subtitle embeds the test statistic, p-value, and **Cliff's delta** (effect size for non-parametric tests, ranging −1 to +1). If p \< 0.05, the two cities have meaningfully different CFR patterns, which motivates expanding the comparison to all five cities.
:::

------------------------------------------------------------------------

## One-Way ANOVA: CFR Across Five Mainland Cities

**Question:** If Selatan and Utara differ, what about all five mainland cities? Who are the extreme cases?

`ggbetweenstats()` with `type = "p"` runs Welch's ANOVA (no equal-variance assumption) and automatically adds pairwise post-hoc brackets only for significant pairs (`pairwise.display = "s"`), with FDR correction.

```{r}
#| fig-height: 7
covid_anova <- covid_cfr %>%
  filter(!str_detect(city, "SERIBU")) %>%
  mutate(city_label = factor(city_label, levels = c("Jkt.\nPusat", "Jkt.\nBarat",
                                                    "Jkt.\nSelatan", "Jkt.\nTimur",
                                                    "Jkt.\nUtara")))

ggbetweenstats(
  data                 = covid_anova,
  x                   = city_label,
  y                   = cfr,
  type                = "p",
  var.equal           = FALSE,
  pairwise.comparisons = TRUE,
  pairwise.display    = "s",
  p.adjust.method     = "fdr",
  messages            = FALSE,
  title               = "CFR Across Mainland Jakarta Cities",
  xlab                = "City",
  ylab                = "Case Fatality Rate (%)"
)
```

```{r}
#| code-fold: true
aov_model <- aov(cfr ~ city_label, data = covid_anova)
summary(aov_model)
```

### `ggbetweenstats` — Summary of Tests by Type

The function automatically selects the right test based on the `type` argument:

| `type` | Test used | Effect size reported | Post-hoc |
|---------------|---------------|---------------------------|---------------|
| `"p"` (parametric) | Student's / Welch's *t* (2 groups); ANOVA / Welch's F (\>2) | Cohen's *d*; ω² | Games-Howell |
| `"np"` (non-parametric) | Mann–Whitney U (2 groups); Kruskal–Wallis (\>2) | Cliff's delta; ε² | Dunn test |
| `"r"` (robust) | Yuen's trimmed-mean *t* (2 groups); robust ANOVA (\>2) | ξ (robust *d*); ξ² | Explanatory measure |
| `"bayes"` (Bayesian) | Bayesian *t*-test / ANOVA | BF₁₀; δ posterior | BF-based pairwise |

For pairwise display:

-   `"ns"` → show only non-significant pairs
-   `"s"` → show only significant pairs *(used here)*
-   `"all"` → show everything

::: callout-note
## What I Learned

The ANOVA reveals which specific city pairs differ, not just whether *any* difference exists. Jakarta Timur tends to have the highest median CFR while Jakarta Pusat has the lowest — a geographic pattern that has real implications for resource allocation and intervention prioritisation. The FDR correction prevents false discoveries when making five-city comparisons simultaneously.
:::

------------------------------------------------------------------------

## Correlation Test: Positive Cases vs Deaths

**Question:** EDA showed both variables are right-skewed with suspected outliers. Is the positive-death relationship linear and strong, or driven by a handful of hotspot sub-districts?

```{r}
#| fig-height: 6
ggscatterstats(
  data          = covid,
  x             = positive,
  y             = death,
  type          = "parametric",
  marginal      = TRUE,
  marginal.type = "histogram",
  centrality.para = "mean",
  title         = "Correlation between Positive Cases and Deaths",
  xlab          = "Positive cases per sub-district",
  ylab          = "Deaths per sub-district",
  messages      = FALSE,
  ggplot.component = list(
    ggplot2::aes(colour = city_label),      # colour by city
    ggplot2::scale_colour_brewer(palette = "Set2", name = "City"),
    ggplot2::theme(legend.position = "right")
  )
)
```

::: callout-note
## What I Learned

The scatterplot combines a regression line, confidence band, and marginal histograms. The subtitle reports Pearson *r*, p-value, and 95% CI. A high *r* (≈ 0.85–0.9) confirms a strong positive linear relationship, but the marginal histograms expose the extreme right skew — a few hotspot sub-districts dominate. This skew is exactly what makes a plain OLS model unreliable, motivating the diagnostic checks in the next section. The interactive version (Section 8) will identify which sub-districts those outliers are.
:::

------------------------------------------------------------------------

## Association Test: CFR Category vs City

**Question:** Does the *proportion* of high/medium/low CFR sub-districts differ across cities?

We bin CFR into three ordered categories and use `ggbarstats()` with a Pearson Chi-square test.

```{r}
#| fig-height: 6
covid_cfr_cat <- covid_cfr %>%
  mutate(
    cfr_cat = case_when(
      cfr < 1.5 ~ "Low (<1.5%)",
      cfr < 3   ~ "Medium (1.5–3%)",
      TRUE      ~ "High (>3%)"
    ),
    cfr_cat = factor(cfr_cat, levels = c("Low (<1.5%)", "Medium (1.5–3%)", "High (>3%)"))
  )

ggbarstats(
  data          = covid_cfr_cat,
  x             = city_label,
  y             = cfr_cat,
  type          = "p",
  label         = "both",
  label.args    = list(size = 2),
  perc.k        = 1,
  title         = "CFR Category Distribution by City",
  xlab          = "City",
  ylab          = "Proportion within city",
  legend.title  = "CFR category",
  messages      = FALSE
)
```

```{r}
#| code-fold: true
chisq <- chisq.test(table(covid_cfr_cat$city_label, covid_cfr_cat$cfr_cat))
chisq
```

::: callout-note
## What I Learned

The stacked proportional bars make it immediately clear that no city has a uniform CFR profile. Kepulauan Seribu stands out with a disproportionately high share of "High" CFR sub-districts — likely reflecting limited healthcare infrastructure in the island district. The Chi-square statistic and **Cramér's V** (effect size for contingency tables) in the subtitle confirm both the significance and practical magnitude of this city-level variation. This completes the picture opened by EDA: the CFR differences across cities are real, large, and not randomly distributed.
:::

------------------------------------------------------------------------

## Model Diagnostics with `performance`

Now that we know cities differ in CFR and that positive cases and deaths are strongly correlated, a natural next step is modelling. We fit a linear model predicting `positive` cases from `recovered`, `death`, and `city`. The `check_model()` function from `performance` visualises four key assumption checks simultaneously.

```{r}
#| fig-height: 8
model_lm <- lm(positive ~ recovered + death + city, data = covid)
check_model(model_lm, panel = FALSE)
```

::: callout-note
## What I Learned

The four diagnostic panels reveal a consistent story:

-   **Q-Q plot (Normality of residuals):** Residuals deviate from the diagonal, particularly in the tails — normality is violated.
-   **Residuals vs Fitted (Homoscedasticity):** The spread of residuals increases with fitted values (fan shape), confirming heteroscedasticity — another OLS assumption broken.
-   **Cook's Distance (Influential observations):** A handful of sub-districts have disproportionately large Cook's distances, meaning they pull the regression line heavily.
-   **VIF (Multicollinearity):** `recovered` and `death` are both derived from `positive`, so their VIFs will be high — the predictors are collinear.

Together, these diagnostics explain *why* the model is unreliable for count data: OLS assumes normally distributed, homoscedastic residuals, but COVID case counts are inherently non-negative integers with high variance. A **Poisson or negative-binomial GLM** would be more appropriate. The `check_model()` output is not just a validation step — it is a navigation tool pointing toward a better model.
:::

------------------------------------------------------------------------

## Visualising Model Parameters with `parameters`

Even with the OLS limitations identified above, visualising the coefficients helps us understand which predictors matter most — and where the collinearity manifests.

```{r}
#| fig-height: 5
params <- model_parameters(model_lm, effects = "fixed")

plot(params, show_intercept = FALSE) +
  ggtitle("Model coefficients (standardised)") +
  theme_minimal(base_size = 12)
```

```{r}
#| code-fold: false
params %>%
  as_tibble() %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, p) %>%
  kbl(caption = "Linear model coefficients (unscaled)", digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

::: callout-note
## What I Learned

The coefficient plot shows **point estimates ± 95% CI** for every predictor. Standardised coefficients allow direct comparison across different scales (e.g., `recovered` vs `death`). Key observations:

-   `death` carries a very large positive coefficient — expected, since deaths are a subset of cases.
-   `recovered` may show a weaker or even negative coefficient once `death` is in the model — a collinearity artefact flagged by the VIF panel.
-   Wide confidence intervals on city coefficients reflect small sub-district sample sizes within each city.

The `parameters` + `see` combination makes this output as easy to read as a forest plot in a clinical paper — one line replaces a manual `tidy()` + `ggplot()` pipeline.
:::

------------------------------------------------------------------------

## Interactive Exploration: Hover to Identify Sub-Districts

Static plots revealed the *pattern* (strong correlation, skewed distribution). This interactive plot reveals the *labels* — which specific sub-districts are the outlier hotspots.

```{r}
#| fig-height: 6
hover_df <- covid %>%
  mutate(tip = paste0(
    "<b>", subdistrict, "</b><br>",
    "City: ", city, "<br>",
    "Positive: ", comma(positive), "<br>",
    "Deaths: ", death
  ))

plot_ly() %>%
  add_markers(
    data     = hover_df,
    x        = ~positive,
    y        = ~death,
    text     = ~tip,
    hoverinfo = "text",
    marker   = list(size = 6, opacity = 0.6, color = "#2c3e50",
                    line = list(color = "white", width = 1))
  ) %>%
  layout(
    title      = list(text = "<b>Positive Cases vs Deaths — Hover for sub-district details</b>", x = 0.5),
    xaxis      = list(title = "Positive cases per sub-district", tickformat = ","),
    yaxis      = list(title = "Deaths per sub-district", tickformat = ","),
    plot_bgcolor  = "rgba(0,0,0,0)",
    paper_bgcolor = "rgba(0,0,0,0)"
  )
```

::: callout-note
## What I Learned

The interactive layer is the final piece of the analytical story. Every outlier flagged by `ggscatterstats` and inflated Cook's distance in `check_model()` can now be named. This is the bridge between statistical output and operational action: once you can hover over an extreme point and see "Penjaringan, Jakarta Utara, 3,200 positive, 98 deaths", the abstract pattern becomes a real place that needs attention. Wrapping any `ggplot2` output in `ggplotly()` is the quick route to the same result for `ggstatsplot` charts.
:::

------------------------------------------------------------------------

## Summary: The Story These Techniques Tell Together

Starting from raw sub-district counts, the eight techniques progressively answered one interconnected question about Jakarta's COVID-19 burden:

```{r}
#| echo: false
tibble(
  `Step` = 1:8,
  Technique = c(
    "One-sample test (gghistostats)",
    "Two-sample test (ggbetweenstats, np)",
    "One-way ANOVA (ggbetweenstats, p)",
    "Correlation test (ggscatterstats)",
    "Association test (ggbarstats)",
    "Model diagnostics (performance)",
    "Model parameters (parameters + see)",
    "Interactive scatterplot (plotly)"
  ),
  `Question answered` = c(
    "Is Jakarta's mean CFR genuinely above a 2% global benchmark?",
    "Do Selatan and Utara have different CFR distributions?",
    "Which mainland cities differ, and by how much?",
    "How strong is the positive–death link, and are outliers driving it?",
    "Is the distribution of CFR severity categories the same across cities?",
    "Is OLS the right model? What assumptions does it violate?",
    "Which predictors matter, and where does collinearity show up?",
    "Which specific sub-districts are the extreme hotspots?"
  ),
  `Key insight` = c(
    "BF₁₀ > 10 → strong evidence that mean CFR exceeds 2%; localised benchmarks needed",
    "CFR in Selatan and Utara are statistically distinct (Cliff's delta ≠ 0)",
    "Jakarta Timur has highest, Pusat lowest; FDR-corrected pairwise gaps confirmed",
    "Strong correlation (r ≈ 0.85) but skew and outliers undermine OLS fit",
    "Kepulauan Seribu has disproportionately more 'High' CFR sub-districts (Cramér's V)",
    "Residuals are non-normal, heteroscedastic; Cook's spikes → GLM recommended",
    "Death dominates; recovered collinear with positive; city CIs are wide",
    "Named hotspots identified; extreme Cook's points become actionable locations"
  )
) %>%
  kbl(caption = "Visual Statistical Analysis — Techniques and Insights Summary",
      align = c("c", "l", "l", "l")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  row_spec(1:8, background = "#fef9e7")
```

::: callout-tip
## The Complete Story

EDA revealed spread and skew → the one-sample test confirmed Jakarta's CFR is above global norms → the two-sample test showed city-level differences are real → ANOVA pinpointed which cities are the outliers → the correlation test proved cases and deaths move together but with hotspot distortion → the association test showed Kepulauan Seribu as a structural anomaly → model diagnostics showed OLS is insufficient for count data → parameter plots revealed which predictors matter despite collinearity → the interactive plot named the hotspots requiring urgent intervention. Each technique answered one question, and together they build a complete public health picture.
:::

------------------------------------------------------------------------
