---
title: "In-class Exercise"
subtitle: "A Layered Grammar of Graphics: ggplot2 methods"
author: "Thi Thanh Van Tran"
date: "April 18, 2026"
date-modified: "April 18, 2026"
format:
  html:
    toc: true
    number-sections: true
    number-depth: 3
    code-fold: false
    code-tools: true
    code-copy: true
execute:
  eval: true
  echo: true
  warning: false
  freeze: false
---

# Getting started

## Installing and loading the required libraries

::: callout-note
**Note:** Ensure that the [pacman](https://cran.r-project.org/web/packages/pacman/) package has already been installed.
:::

The code chunk below uses `p_load()` of pacman package to check if readxl packages are installed in the computer. If they are, then they will be launched into R. Otherwise,readxl will be installed and launched into R.

```{r}
pacman::p_load(readxl)
```

## Importing data

-   The code chunk below imports *Superstore.csv* into R environment by using [*read_excel()*](https://readr.tidyverse.org/reference/read_delim.html) function of [**readr**](https://readr.tidyverse.org/) package.
-   **readr** is one of the tidyverse package.

```{r}
Superstore <- read_excel("data/Superstore.csv")
```

**exam_data contains:**
---
