---
title: "Hands-on Exercise 3"
subtitle: "Interactive Data Visualisation — ggiraph & plotly"
author: "Thi Thanh Van Tran"
date: "April 22, 2026"
date-modified: "April 22, 2026"
format:
  html:
    toc: true
    number-sections: true
    number-depth: 4
    code-fold: false
    code-summary: "Show code"
    code-tools: false
    code-copy: true
    css: styles.css
    include-after-body: toc-tracker.html
    theme: cosmo
    page-layout: article
execute:
  eval: true
  echo: true
  warning: false
  freeze: false
knitr:
  opts_chunk:
    out.width: "80%"
    fig.align: "center"
    fig.width: 12
    fig.height: 7
    dpi: 120
---

::: welcome-section
# Interactive Data Visualisation: Mastering ggiraph, plotly & crosstalk

Welcome! This hands-on exercise covers **three complementary tools** for interactive visualisation in R:

1. **ggiraph** — extends ggplot2 with hover tooltips, data\_id linking, and onclick actions
2. **plotly** — full-featured interactive charts via `plot_ly()` and `ggplotly()`
3. **crosstalk** — cross-widget interactions linking plots to interactive tables
:::

---

# Getting Started

## Installing Packages

The code chunk below uses `pacman::p_load()` to check, install and launch all required packages in one step.

```{r}
#| code-fold: false
pacman::p_load(ggiraph, plotly, DT, tidyverse, readxl,
               knitr, patchwork, viridis, scales, crosstalk)
```

Packages used:

- **ggiraph** — makes ggplot2 graphics interactive via tooltip, data\_id, and onclick aesthetics
- **plotly** — creates interactive web graphics from ggplot2 or via its own `plot_ly()` interface
- **DT** — wraps the JavaScript DataTables library for interactive HTML tables
- **tidyverse** — modern R packages for data science (includes ggplot2, dplyr, readr)
- **patchwork** — combines multiple ggplot2 graphs; used inside `girafe()` for coordinated views
- **crosstalk** — enables cross-widget interactions (linked brushing and filtering) between htmlwidgets
- **readxl**, **viridis**, **scales** — supporting utilities for data import and visual styling

## Importing Data

```{r}
#| code-fold: false
pop_data <- read_xls("data/GlobalPopulation.xls", sheet = "Data") %>%
  mutate(across(where(is.character), as.factor))
```

The dataset `GlobalPopulation.xls` contains country-level demographic records with fields: `Country`, `Continent`, `Year`, `Population`, `Young` (young-age population), and `Old` (elderly population). It is the dataset we will use throughout this exercise to demonstrate interactive visualisation techniques.

---

# Visual Data Exploration

::: {.panel-tabset}

## Data Overview

```{r}
#| code-fold: true
#| fig.height: 10

n_records  <- nrow(pop_data)
n_complete <- sum(complete.cases(pop_data))

var_types <- data.frame(Type = sapply(pop_data, function(x) class(x)[1])) %>%
  count(Type)

p_sample <- ggplot(data.frame(x = 1, y = 1), aes(x, y)) +
  geom_tile(fill = "#fce7f3", alpha = 0.5, width = 0.9, height = 0.9) +
  geom_text(label = paste0("n = ", format(n_records, big.mark = ","), "\nrecords"),
            size = 8, fontface = "bold", color = "#831843") +
  labs(title = "Sample Size") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

p_types <- ggplot(var_types, aes(x = reorder(Type, -n), y = n, fill = Type)) +
  geom_col(alpha = 0.8, width = 0.6) +
  geom_text(aes(label = n), vjust = -0.5, fontface = "bold", size = 5) +
  scale_fill_manual(values = c("character" = "#fbcfe8", "factor" = "#e9d5ff",
                                "numeric" = "#bfdbfe")) +
  labs(title = "Variable Types", x = "Type", y = "Count") +
  ylim(0, max(var_types$n) * 1.2) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        legend.position = "none")

missing_df <- reshape2::melt(is.na(pop_data))

p_missing <- ggplot(missing_df, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white", linewidth = 0.05) +
  scale_fill_manual(values = c("FALSE" = "#10b981", "TRUE" = "#ef4444"),
                    labels = c("Complete", "Missing"), name = "Status") +
  labs(title = "Missing Value Heatmap", x = "Variables", y = "Observations") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x  = element_text(angle = 45, hjust = 1),
        panel.grid   = element_blank())

(p_sample | p_types) / p_missing +
  plot_annotation(title = "Dataset Overview",
                  theme = theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))) +
  plot_layout(heights = c(1, 1.2))
```

::: {.callout-note}
## Analysis
The dataset has `r format(n_records, big.mark=",")` records with `r sum(var_types$n[var_types$Type %in% c("character","factor")])` categorical and `r sum(var_types$n[var_types$Type=="numeric"])` numeric variables. The missing value heatmap reveals 100% complete no missing value
:::

## Data Insights

```{r}
#| code-fold: true
#| fig.height: 10

cat_vars      <- names(pop_data)[sapply(pop_data, function(x) is.character(x) || is.factor(x))]
pastel_colors <- c("#fbcfe8","#bfdbfe","#bbf7d0","#e9d5ff","#fde68a","#fecaca","#a7f3d0","#ddd6fe")

plot_list <- lapply(cat_vars, function(var) {
  plot_data <- pop_data %>%
    group_by(.data[[var]]) %>%
    summarise(n = n(), .groups = "drop") %>%
    mutate(pct = n / sum(n) * 100) %>%
    arrange(desc(n))

  n_cats      <- nrow(plot_data)
  fill_colors <- rep_len(pastel_colors, n_cats)
  title_suffix <- if (n_cats > 20) { plot_data <- slice_head(plot_data, n = 20); " (Top 20)" } else ""

  ggplot(plot_data, aes(x = reorder(.data[[var]], -n), y = n, fill = .data[[var]])) +
    geom_col(alpha = 0.9, width = 0.7, color = "white") +
    geom_text(aes(label = paste0(n, "\n(", round(pct,1), "%)")),
              vjust = -0.2, size = 2.5, fontface = "bold", color = "#831843") +
    scale_fill_manual(values = fill_colors) +
    labs(title = paste(var, "Distribution", title_suffix), y = "Count", x = "") +
    ylim(0, max(plot_data$n) * 1.15) +
    theme_minimal(base_size = 10) +
    theme(plot.title    = element_text(hjust = 0.5, face = "bold", size = 11),
          legend.position = "none",
          axis.text.x   = element_text(angle = 45, hjust = 1, size = 7))
})

wrap_plots(plot_list, ncol = 2) +
  plot_annotation(title = "Categorical Variable Distributions",
                  theme = theme(plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))
```

::: {.callout-note}
## Analysis
Always `group_by` before counting to get true frequencies per category. Africa has the most country-year records while Asia dominates by total population — a distinction only visible after proper aggregation. These distributions guide which grouping variables are most useful for colour-encoding in interactive charts.
:::

## Understanding Data Relationships

```{r}
#| code-fold: true
#| fig.height: 10

latest_year     <- max(pop_data$Year, na.rm = TRUE)
example_country <- pop_data %>% filter(Country == "China") %>% arrange(Year)

v1 <- pop_data %>% filter(Year == latest_year) %>%
  ggplot(aes(x = Young, y = Population, size = Population, color = Continent)) +
  geom_point(alpha = 0.6) +
  scale_size_continuous(range = c(2, 15)) +
  scale_color_viridis_d(option = "plasma") +
  labs(title = "Q1: Young vs Population", x = "Young Population", y = "Total Population") +
  theme_minimal() + theme(plot.title = element_text(face = "bold"))

v2 <- ggplot(example_country, aes(x = Year, y = Old)) +
  geom_line(color = "#ec4899", linewidth = 1.2) + geom_point(color = "#831843", size = 3) +
  labs(title = "Q2: China — Aging Trend", x = "Year", y = "Old Population") +
  theme_minimal() + theme(plot.title = element_text(face = "bold"))

v3 <- ggplot(example_country, aes(x = Year, y = Young)) +
  geom_line(color = "#3b82f6", linewidth = 1.2) + geom_point(color = "#1e40af", size = 3) +
  labs(title = "Q3: China — Youth Trend", x = "Year", y = "Young Population") +
  theme_minimal() + theme(plot.title = element_text(face = "bold"))

v4 <- pop_data %>% filter(Year == latest_year) %>%
  ggplot(aes(x = Young, y = Old, color = Continent)) +
  geom_point(alpha = 0.7, size = 3) +
  scale_color_viridis_d(option = "turbo") +
  labs(title = "Q4: Young vs Old by Continent", x = "Young Population", y = "Old Population") +
  theme_minimal() + theme(legend.position = "right", plot.title = element_text(face = "bold"))

(v1 | v2) / (v3 | v4) +
  plot_annotation(title = "Understanding Data Relationships",
                  theme = theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5)))
```

::: {.callout-important}
## Analysis
Each chart type matches the analytical question: bubble charts encode three variables at once; line charts trace temporal change; scatter plots expose bivariate structure. China's Q2 and Q3 panels together tell the demographic squeeze story — a rising elderly share alongside a plateauing youth share — that no single chart alone reveals.
:::

## Data Summary

```{r}
#| code-fold: true
glimpse(pop_data)
```

```{r}
#| code-fold: true
summary(pop_data %>% select(where(is.numeric)))
```

::: {.callout-note}
## Analysis
`glimpse()` confirms variable types and detects encoding issues in one step; `summary()` exposes range and quartile information, flagging extreme population values before interactive visualisation begins.
:::

:::

---

# ggiraph Methods

::: {.callout-tip}
## Core Concept
**ggiraph** is an htmlwidget and ggplot2 extension that makes ggplot graphics interactive via three aesthetics: `tooltip` (text shown on hover), `data_id` (key for hover highlighting and cross-plot linking), and `onclick` (JavaScript executed on click). Workflow: swap `geom_*()` → `geom_*_interactive()`, then wrap in `girafe()`.
:::

## Tooltip Effect with `tooltip` Aesthetic {#sec-tooltip-basic}

The code chunk below plots a basic interactive chart using `geom_point_interactive()`. Two steps are involved: first build a ggplot object with an interactive geom; then pass it to `girafe()` to generate an SVG displayed on the HTML page.

```{r}
#| code-fold: true

p <- pop_data %>%
  filter(Year == max(Year, na.rm = TRUE)) %>%
  ggplot(aes(x = Young, y = Population)) +
  geom_point_interactive(
    aes(tooltip = Country),
    size = 3, alpha = 0.7, color = "#831843"
  ) +
  labs(title = "Young vs Population — Basic Tooltip (hover for country name)",
       x = "Young Population", y = "Total Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 13))

girafe(
  ggobj      = p,
  width_svg  = 10,
  height_svg = 10 * 0.618
)
```

Interactivity: hovering the mouse over any data point displays the country name. Notice that the only change from a standard ggplot is replacing `geom_point()` with `geom_point_interactive()` and adding `tooltip = Country` inside `aes()`, then wrapping the whole object in `girafe()`.

::: {.callout-note}
## What I Learned
The two-step pattern — interactive geom → `girafe()` — is the mental model for all ggiraph work. The SVG output embeds directly in HTML reports with no extra server required. This is the cleanest entry point: minimal code change, immediate hover feedback.
:::

## Displaying Multiple Information on Tooltip {#sec-tooltip-multi}

Tooltip content can be customised by constructing a character column that combines multiple fields:

```{r}
#| code-fold: true

latest <- pop_data %>%
  filter(Year == max(Year, na.rm = TRUE)) %>%
  mutate(tooltip_text = paste0(
    "Country = ", Country,
    "\nContinent = ", Continent
  ))

p <- ggplot(data = latest, aes(x = Young, y = Population)) +
  geom_point_interactive(
    aes(tooltip = tooltip_text),
    size = 3, alpha = 0.7, color = "#1e40af"
  ) +
  labs(title = "Young vs Population — Multi-field Tooltip",
       x = "Young Population", y = "Total Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 13))

girafe(
  ggobj      = p,
  width_svg  = 10,
  height_svg = 10 * 0.618
)
```

Interactivity: hovering shows both Country and Continent. The `paste0()` call builds a multi-line string by inserting `\n` between fields; this string is pre-computed as a new column and then referenced in `aes(tooltip = ...)`. The key insight is that the tooltip column must exist in the data frame before being passed to the aesthetic.

::: {.callout-note}
## What I Learned
Pre-computing a tooltip column with `mutate()` before the ggplot call keeps the `aes()` clean and makes complex tooltips easy to debug. Calling `paste0()` with explicit field labels (e.g., `"Country = "`) makes tooltips self-explanatory for end users.
:::

## Customising Tooltip Style {#sec-tooltip-style}

`opts_tooltip()` inside `girafe()` applies CSS to the tooltip container:

```{r}
#| code-fold: true

tooltip_css <- "background-color:#1e293b; color:#f8fafc; font-weight:bold;
                padding:8px 12px; border-radius:6px; font-size:13px;"

latest <- pop_data %>%
  filter(Year == max(Year, na.rm = TRUE)) %>%
  mutate(tooltip_text = paste0(
    "Country = ", Country,
    "\nContinent = ", Continent
  ))

p <- ggplot(data = latest, aes(x = Young, y = Population)) +
  geom_point_interactive(
    aes(tooltip = tooltip_text),
    size = 3, alpha = 0.7, color = "#059669"
  ) +
  labs(title = "Young vs Population — Styled Tooltip (dark background)",
       x = "Young Population", y = "Total Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 13))

girafe(
  ggobj      = p,
  width_svg  = 10,
  height_svg = 10 * 0.618,
  options    = list(
    opts_tooltip(css = tooltip_css)
  )
)
```

The `tooltip_css` string is a standard CSS declaration. `opts_tooltip(css = ...)` passes it directly to the SVG tooltip element. The same CSS properties available in web development — `background-color`, `color`, `font-weight`, `border-radius` — work here. Refer to [Customizing girafe objects](https://www.ardata.fr/ggiraph-book/customize.html) for a full list of options.

::: {.callout-note}
## What I Learned
Tooltip CSS is applied at the `girafe()` level, not inside `aes()`, which keeps styling concerns cleanly separated from data mapping. Dark backgrounds with light text (`#1e293b` / `#f8fafc`) immediately distinguish the tooltip from the chart background.
:::

## Displaying Statistics on Tooltip {#sec-tooltip-stats}

A function can compute derived statistics and embed them in the tooltip using `after_stat()`:

```{r}
#| code-fold: true

tooltip_fn <- function(y, ymax, accuracy = 0.01) {
  mean_val <- scales::number(y, accuracy = accuracy)
  sem_val  <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean Young Population:", mean_val, "+/-", sem_val)
}

gg_bar <- ggplot(data = pop_data %>% filter(Year == max(Year, na.rm = TRUE)),
                 aes(x = Continent)) +
  stat_summary(
    aes(y = Young,
        tooltip = after_stat(tooltip_fn(y, ymax))),
    fun.data = "mean_se",
    geom     = GeomInteractiveCol,
    fill     = "lightblue"
  ) +
  stat_summary(
    aes(y = Young),
    fun.data = mean_se,
    geom     = "errorbar",
    width    = 0.2, linewidth = 0.4
  ) +
  labs(title = "Mean Young Population by Continent — Statistics on Tooltip",
       x = "Continent", y = "Mean Young Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 13),
        axis.text.x = element_text(angle = 30, hjust = 1))

girafe(ggobj = gg_bar, width_svg = 10, height_svg = 10 * 0.618)
```

`after_stat()` defers evaluation of `tooltip_fn(y, ymax)` until `stat_summary` has computed the summary statistics, meaning `y` and `ymax` refer to the computed mean and upper confidence bound — not raw data columns. `GeomInteractiveCol` is the interactive version of `GeomCol`, enabling the tooltip on bar charts.

::: {.callout-note}
## What I Learned
`after_stat()` is the bridge between statistical transformations and interactive aesthetics. Without it, computed values like `y` (mean) and `ymax` (mean + SE) would not be accessible inside `aes()`. This pattern is essential whenever you want to show computed statistics — not raw values — on hover.
:::

## Hover Effect with `data_id` Aesthetic {#sec-data-id}

`data_id` is the second interactive feature of ggiraph. Elements sharing the same `data_id` value are highlighted together on hover:

```{r}
#| code-fold: true

latest <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

p <- ggplot(data = latest, aes(x = Young, y = Population, color = Continent)) +
  geom_point_interactive(
    aes(data_id = Continent),
    size = 3, alpha = 0.7
  ) +
  scale_color_viridis_d(option = "plasma") +
  labs(title = "Hover to highlight all countries in the same continent",
       x = "Young Population", y = "Total Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 13))

girafe(
  ggobj      = p,
  width_svg  = 10,
  height_svg = 10 * 0.618
)
```

Interactivity: hovering over any point highlights all points sharing the same `Continent` value. The default hover CSS is `fill:orange`. Note that `data_id` and `tooltip` are independent: you can use either or both.

::: {.callout-note}
## What I Learned
`data_id` is the mechanism that makes ggiraph's linking magic work. Assigning the same key to multiple elements (all points from Asia, for example) causes them to respond as a group. This is the foundation for coordinated views later.
:::

## Styling the Hover Effect {#sec-hover-style}

CSS codes inside `opts_hover()` and `opts_hover_inv()` change the highlighting colour and dim non-hovered elements:

```{r}
#| code-fold: true

latest <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

p <- ggplot(data = latest, aes(x = Young, y = Population, color = Continent)) +
  geom_point_interactive(
    aes(data_id = Continent),
    size = 3, alpha = 0.7
  ) +
  scale_color_viridis_d(option = "plasma") +
  labs(title = "Styled hover: highlighted in dark, others dimmed",
       x = "Young Population", y = "Total Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 13))

girafe(
  ggobj      = p,
  width_svg  = 10,
  height_svg = 10 * 0.618,
  options    = list(
    opts_hover(css     = "fill:#202020; stroke:#202020;"),
    opts_hover_inv(css = "opacity:0.15;")
  )
)
```

`opts_hover()` defines the CSS applied to the *hovered* element; `opts_hover_inv()` applies to all *other* elements. Setting `opacity:0.15` on non-hovered elements creates a focus effect that naturally draws attention to the active group without removing context.

::: {.callout-note}
## What I Learned
The `opts_hover_inv()` pattern — dim everything else — is far more effective than just highlighting the active element. The contrast between `opacity:0.15` (background) and the bold dark fill (hovered) creates immediate visual hierarchy with only two lines of CSS.
:::

## Combining Tooltip and Hover Effect {#sec-tooltip-hover}

Tooltip and `data_id` can be combined in the same `aes()` call:

```{r}
#| code-fold: true

latest <- pop_data %>%
  filter(Year == max(Year, na.rm = TRUE)) %>%
  mutate(tooltip_text = paste0(
    "Country = ", Country,
    "\nContinent = ", Continent,
    "\nYoung = ", format(Young, big.mark = ","),
    "\nPopulation = ", format(Population, big.mark = ",")
  ))

p <- ggplot(data = latest, aes(x = Young, y = Population, color = Continent)) +
  geom_point_interactive(
    aes(tooltip = tooltip_text,
        data_id = Continent),
    size = 3, alpha = 0.7
  ) +
  scale_color_viridis_d(option = "turbo") +
  labs(title = "Combined: tooltip shows detail, hover highlights continent group",
       x = "Young Population", y = "Total Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 13))

girafe(
  ggobj      = p,
  width_svg  = 10,
  height_svg = 10 * 0.618,
  options    = list(
    opts_hover(css     = "fill:#202020; stroke:#202020;"),
    opts_hover_inv(css = "opacity:0.15;")
  )
)
```

Interactivity: hovering over a point highlights the entire continent group **and** shows the individual country's detail in the tooltip. The two aesthetics work simultaneously — `tooltip` gives individual context while `data_id` gives group context.

::: {.callout-note}
## What I Learned
Combining tooltip and data_id delivers two levels of information at once: the tooltip tells you *what* a specific point is, while the hover highlight tells you *which group it belongs to*. This layering is a powerful technique for hierarchical datasets like country-within-continent.
:::

## Click Effect with `onclick` {#sec-onclick}

The `onclick` aesthetic accepts any valid JavaScript string and executes it when a point is clicked:

```{r}
#| code-fold: true

latest_onclick <- pop_data %>%
  filter(Year == max(Year, na.rm = TRUE)) %>%
  mutate(onclick_js = paste0(
    "window.open('https://www.google.com/search?q=",
    URLencode(paste(as.character(Country), "population statistics")),
    "')"
  ),
  tooltip_text = paste0(Country, "\n(click to search Google)"))

p <- ggplot(data = latest_onclick, aes(x = Young, y = Population, color = Continent)) +
  geom_point_interactive(
    aes(tooltip = tooltip_text,
        onclick  = onclick_js),
    size = 3.5, alpha = 0.7
  ) +
  scale_color_viridis_d(option = "plasma") +
  labs(title = "Click any point to open a Google search for that country",
       x = "Young Population", y = "Total Population") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 13))

girafe(
  ggobj      = p,
  width_svg  = 10,
  height_svg = 10 * 0.618,
  options    = list(
    opts_hover(css = "fill:#FF3333; cursor:pointer;")
  )
)
```

::: {.callout-warning}
Note that click actions must be a **string column** in the dataset containing valid JavaScript instructions. The `onclick` value is evaluated as raw JavaScript in the browser — errors in the JS string will silently fail.
:::

Interactivity: clicking a country point opens a new browser tab with a Google search for that country's population statistics. `sprintf()` or `paste0()` with `URLencode()` are the standard approaches for building JavaScript `window.open()` strings safely.

::: {.callout-note}
## What I Learned
The `onclick` aesthetic turns every data point into a hyperlink — bridging the visualisation to external information systems. `cursor:pointer` in `opts_hover()` signals clickability to users without any additional UI elements. Always pre-compute the JS string as a data column; never build it inside `aes()`.
:::

## Coordinated Multiple Views with ggiraph {#sec-ggiraph-coordinated}

When multiple ggiraph plots share the same `data_id` values and are combined with `patchwork` inside `girafe()`, hovering or clicking any element highlights its counterpart across all panels simultaneously:

```{r}
#| code-fold: true

latest <- pop_data %>%
  filter(Year == max(Year, na.rm = TRUE)) %>%
  mutate(tooltip_text = paste0("<b>", Country, "</b><br>Continent: ", Continent))

p1 <- ggplot(data = latest, aes(x = Young, y = Population, color = Continent)) +
  geom_point_interactive(
    aes(data_id = Country,
        tooltip = paste0(tooltip_text,
                         "<br>Young: ", format(Young, big.mark = ","),
                         "<br>Population: ", format(Population, big.mark = ","))),
    size = 3, alpha = 0.7
  ) +
  scale_color_viridis_d(option = "plasma") +
  coord_cartesian(xlim = c(0, max(latest$Young, na.rm = TRUE) * 1.05)) +
  labs(title = "Young vs Total Population", x = "Young", y = "Total Population") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold", size = 11))

p2 <- ggplot(data = latest, aes(x = Young, y = Old, color = Continent)) +
  geom_point_interactive(
    aes(data_id = Country,
        tooltip = paste0(tooltip_text,
                         "<br>Young: ", format(Young, big.mark = ","),
                         "<br>Old: ", format(Old, big.mark = ","))),
    size = 3, alpha = 0.7
  ) +
  scale_color_viridis_d(option = "turbo") +
  coord_cartesian(xlim = c(0, max(latest$Young, na.rm = TRUE) * 1.05)) +
  labs(title = "Young vs Old Population", x = "Young", y = "Old Population") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold", size = 11))

girafe(
  code       = print(p1 + p2),
  width_svg  = 12,
  height_svg = 4.5,
  options    = list(
    opts_hover(css     = "fill:#202020; stroke:#202020;"),
    opts_hover_inv(css = "opacity:0.2;")
  )
)
```

Notice that when a data point in the left panel is hovered, the corresponding country point in the right panel is highlighted too. The `data_id` aesthetic is critical to link observations between plots — here `Country` is the shared key. The tooltip aesthetic is optional but provides useful per-point detail.

The programming strategy:

1. Appropriate interactive geoms with matching `data_id` values are used to create each panel.
2. `patchwork` (`+`) is used *inside* `girafe(code = print(...))` — this is the key: passing combined plots through the `code` argument activates cross-panel linking.

::: {.callout-important}
## What I Learned
Coordinated views require two things to work: (1) identical `data_id` keys across all panels, and (2) the patchwork combination placed *inside* `girafe()` via the `code` argument rather than `ggobj`. When both conditions are met, ggiraph handles all the JavaScript linking automatically — no manual event handling required.
:::

---

# plotly Methods

::: {.callout-tip}
## Core Concept
Plotly's R library creates interactive web graphics via two approaches: `plot_ly()` (native plotly syntax using `~` column references) and `ggplotly()` (converts existing ggplot2 objects). Both produce charts with built-in zoom, pan, hover tooltips, and download buttons.
:::

## Creating an Interactive Scatter Plot: `plot_ly()` Method {#sec-plotly-basic}

`plot_ly()` is plotly's native interface. Column references use the tilde (`~`) notation:

::: {.panel-tabset}

### The Plot

```{r}
#| code-fold: false
#| echo: false

latest_plotly <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

plot_ly(data = latest_plotly,
        x    = ~Young,
        y    = ~Old)
```

### The Code Chunk

```{r}
#| eval: false

latest_plotly <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

plot_ly(data = latest_plotly,
        x    = ~Young,
        y    = ~Old)
```

:::

The default output is already interactive: hover to see coordinates, drag to zoom, double-click to reset. The `~` before column names tells plotly to look up those names in the `data` argument. No additional configuration is required for a basic scatter.

::: {.callout-note}
## What I Learned
`plot_ly()` produces a fully interactive chart with a single call — no `ggplot()` + `geom_*()` chain required. The tilde syntax is distinctive to plotly and signals that the value is a column reference, not a literal. The default toolbar (zoom, pan, select, download) is added automatically.
:::

## Working with Visual Variables: `plot_ly()` Method {#sec-plotly-color}

The `color` argument maps a qualitative variable to colour, and clicking legend symbols toggles trace visibility:

::: {.panel-tabset}

### The Plot

```{r}
#| code-fold: false
#| echo: false

latest_plotly <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

plot_ly(data   = latest_plotly,
        x      = ~Young,
        y      = ~Old,
        color  = ~Continent,
        colors = viridis::viridis(length(unique(latest_plotly$Continent))))
```

### The Code Chunk

```{r}
#| eval: false

latest_plotly <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

plot_ly(data   = latest_plotly,
        x      = ~Young,
        y      = ~Old,
        color  = ~Continent,
        colors = viridis::viridis(length(unique(latest_plotly$Continent))))
```

:::

Interactivity: click on a continent's colour symbol in the legend to show/hide that group. Double-click a legend item to isolate it. The `colors` argument accepts a named vector or a palette function — here `viridis::viridis()` generates the correct number of distinct colours automatically.

::: {.callout-note}
## What I Learned
Mapping `color` to a factor in `plot_ly()` automatically creates a legend with toggle interactivity — a feature that would require custom JavaScript in a static chart. Clicking continent symbols to isolate groups is far more intuitive for exploratory analysis than filtering data and re-rendering.
:::

## Creating an Interactive Scatter Plot: `ggplotly()` Method {#sec-ggplotly}

`ggplotly()` converts any existing ggplot2 object into a plotly chart. The only extra line is the `ggplotly()` wrapper:

::: {.panel-tabset}

### The Plot

```{r}
#| code-fold: false
#| echo: false

latest_gg <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

p <- ggplot(data = latest_gg, aes(x = Young, y = Old, color = Continent)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_viridis_d(option = "plasma") +
  coord_cartesian(xlim = c(0, max(latest_gg$Young, na.rm = TRUE) * 1.05),
                  ylim = c(0, max(latest_gg$Old,   na.rm = TRUE) * 1.05)) +
  labs(title = "Young vs Old Population", x = "Young Population", y = "Old Population") +
  theme_minimal()

ggplotly(p)
```

### The Code Chunk

```{r}
#| eval: false

latest_gg <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

p <- ggplot(data = latest_gg, aes(x = Young, y = Old, color = Continent)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_viridis_d(option = "plasma") +
  coord_cartesian(xlim = c(0, max(latest_gg$Young, na.rm = TRUE) * 1.05),
                  ylim = c(0, max(latest_gg$Old,   na.rm = TRUE) * 1.05)) +
  labs(title = "Young vs Old Population", x = "Young Population", y = "Old Population") +
  theme_minimal()

ggplotly(p)
```

:::

Notice that the only additional line needed is `ggplotly(p)`. The zoom toolbar, hover tooltips, and legend toggles are added automatically. This is the fastest path to interactivity when you already have a ggplot2 pipeline.

::: {.callout-note}
## What I Learned
`ggplotly()` is the bridge between the ggplot2 ecosystem and plotly interactivity. When you already have carefully crafted ggplot code (with proper themes, colour scales, labels), `ggplotly()` preserves all of it while adding the full plotly toolbar. The trade-off is that some advanced ggplot2 features (like `geom_sf`) may not translate perfectly.
:::

## Coordinated Multiple Views with plotly {#sec-plotly-coordinated}

Creating coordinated linked plots with plotly involves three steps:

1. `highlight_key()` wraps the shared data object
2. Two scatter plots are built using ggplot2 functions on that shared data
3. `subplot()` places them side by side; clicking a point in one selects the corresponding point in the other

::: {.panel-tabset}

### The Plot

```{r}
#| code-fold: false
#| echo: false

latest_coord <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

d <- highlight_key(latest_coord)

p1 <- ggplot(data = d, aes(x = Young, y = Old)) +
  geom_point(size = 2, alpha = 0.7, color = "#831843") +
  coord_cartesian(xlim = c(0, max(latest_coord$Young, na.rm = TRUE) * 1.05),
                  ylim = c(0, max(latest_coord$Old,   na.rm = TRUE) * 1.05)) +
  labs(title = "Young vs Old", x = "Young Population", y = "Old Population") +
  theme_minimal()

p2 <- ggplot(data = d, aes(x = Young, y = Population)) +
  geom_point(size = 2, alpha = 0.7, color = "#1e40af") +
  coord_cartesian(xlim = c(0, max(latest_coord$Young,      na.rm = TRUE) * 1.05),
                  ylim = c(0, max(latest_coord$Population,  na.rm = TRUE) * 1.05)) +
  labs(title = "Young vs Total Population", x = "Young Population", y = "Total Population") +
  theme_minimal()

subplot(ggplotly(p1), ggplotly(p2))
```

### The Code Chunk

```{r}
#| eval: false

latest_coord <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

d <- highlight_key(latest_coord)

p1 <- ggplot(data = d, aes(x = Young, y = Old)) +
  geom_point(size = 2, alpha = 0.7, color = "#831843") +
  coord_cartesian(xlim = c(0, max(latest_coord$Young, na.rm = TRUE) * 1.05),
                  ylim = c(0, max(latest_coord$Old,   na.rm = TRUE) * 1.05)) +
  labs(title = "Young vs Old", x = "Young Population", y = "Old Population") +
  theme_minimal()

p2 <- ggplot(data = d, aes(x = Young, y = Population)) +
  geom_point(size = 2, alpha = 0.7, color = "#1e40af") +
  coord_cartesian(xlim = c(0, max(latest_coord$Young,      na.rm = TRUE) * 1.05),
                  ylim = c(0, max(latest_coord$Population,  na.rm = TRUE) * 1.05)) +
  labs(title = "Young vs Total Population", x = "Young Population", y = "Total Population") +
  theme_minimal()

subplot(ggplotly(p1), ggplotly(p2))
```

:::

Click on a data point in one scatter plot and see how the corresponding point in the other scatter plot is selected.

Key things to learn:

- `highlight_key()` creates an object of class `crosstalk::SharedData` — both plots reference the same shared data object, which enables the selection to propagate across panels.
- `subplot()` places plotly charts side by side and maintains the shared data linkage.
- Visit the [crosstalk documentation](https://rstudio.github.io/crosstalk/) to learn more.

::: {.callout-note}
## What I Learned
`highlight_key()` is the glue between the two plots — it wraps a data frame with a crosstalk SharedData object so plotly knows which rows correspond across charts. Without it, `subplot()` would show two independent charts. The pattern is: shared data → two ggplots → two ggplotly calls → one `subplot()`.
:::

---

# crosstalk Methods

::: {.callout-tip}
## Core Concept
[crosstalk](https://rstudio.github.io/crosstalk/index.html) extends htmlwidgets with cross-widget interactions — specifically **linked brushing** (selection propagates across widgets) and **filtering** (one widget filters another). It currently works with plotly, DT, and leaflet.
:::

## Interactive Data Table: DT Package {#sec-dt}

The **DT** package wraps the JavaScript DataTables library. R data frames rendered with `DT::datatable()` become searchable, sortable, and paginated HTML tables:

```{r}
DT::datatable(pop_data, class = "compact")
```

The resulting table supports column sorting (click headers), global search (top-right box), and pagination. `class = "compact"` reduces row padding for denser display. No JavaScript knowledge is required.

::: {.callout-note}
## What I Learned
`DT::datatable()` is the fastest way to make any R data frame explorable in an HTML report. The search box alone — filtering 322+ rows in real time — replaces multiple manual `filter()` calls. Tables and charts are **complementary**: charts show patterns, tables expose individual values.
:::

## Linked Brushing: crosstalk Method {#sec-crosstalk-brushing}

Linked brushing connects a scatter plot and a data table so that selecting points in the plot filters the table rows, and vice versa:

::: {.panel-tabset}

### The Plot

```{r}
#| code-fold: false
#| echo: false

latest_brush <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

d <- highlight_key(latest_brush)

p <- ggplot(d, aes(Young, Old, color = Continent)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_viridis_d(option = "plasma") +
  coord_cartesian(xlim = c(0, max(latest_brush$Young, na.rm = TRUE) * 1.05),
                  ylim = c(0, max(latest_brush$Old,   na.rm = TRUE) * 1.05)) +
  labs(x = "Young Population", y = "Old Population") +
  theme_minimal()

gg <- highlight(ggplotly(p), "plotly_selected")

crosstalk::bscols(gg,
                  DT::datatable(d),
                  widths = 5)
```

### The Code Chunk

```{r}
#| eval: false

latest_brush <- pop_data %>% filter(Year == max(Year, na.rm = TRUE))

d <- highlight_key(latest_brush)

p <- ggplot(d, aes(Young, Old, color = Continent)) +
  geom_point(size = 2, alpha = 0.7) +
  scale_color_viridis_d(option = "plasma") +
  coord_cartesian(xlim = c(0, max(latest_brush$Young, na.rm = TRUE) * 1.05),
                  ylim = c(0, max(latest_brush$Old,   na.rm = TRUE) * 1.05)) +
  labs(x = "Young Population", y = "Old Population") +
  theme_minimal()

gg <- highlight(ggplotly(p), "plotly_selected")

crosstalk::bscols(gg,
                  DT::datatable(d),
                  widths = 5)
```

:::

Click and drag to select points in the scatter plot — the table on the right will filter to show only those rows. Things to learn from the code:

- `highlight()` is a **plotly** function that sets brushing options for linked selection. `"plotly_selected"` specifies that selection is triggered by a drag-select (lasso or box) in plotly.
- `bscols()` is a **crosstalk** helper function that places HTML elements side by side using Bootstrap columns. It is designed for R Markdown and Quarto documents. Note: this brings in Bootstrap CSS.
- The `widths = 5` argument allocates 5/12 Bootstrap columns to each widget (plotly + DT), leaving 2 columns as margin.

::: {.callout-important}
## What I Learned
Linked brushing with crosstalk combines two mental models: plotly handles the selection event (`plotly_selected`), `highlight_key()` propagates the selection to all widgets sharing the same SharedData object, and DT reacts by filtering its rows. The result is a powerful explore-and-drill-down workflow: spot a cluster in the chart, drag to select it, read the exact country names and values in the table — all without writing any JavaScript.
:::

---

# References

## ggiraph

- [ggiraph package documentation](https://davidgohel.github.io/ggiraph/index.html) — online reference guide and articles
- [Customizing girafe objects](https://www.ardata.fr/ggiraph-book/customize.html) — CSS options for tooltip, hover, and selection
- [ggiraph Book — Selections](https://www.ardata.fr/ggiraph-book/selections.html) — data_id, Shiny integration

## plotly for R

- [Getting Started with Plotly in R](https://plotly.com/r/getting-started/)
- Carson Sievert (2020) *Interactive web-based data visualization with R, plotly, and shiny*, Chapman and Hall/CRC — [online version](https://plotly-r.com/)
- [Plotly R Figure Reference](https://plotly.com/r/reference/index/) — comprehensive visual reference
- [Plotly R Library Fundamentals](https://plotly.com/r/plotly-fundamentals/)

## crosstalk & DT

- [crosstalk documentation](https://rstudio.github.io/crosstalk/)
- [DT package guide](https://rstudio.github.io/DT/)

## Source

- [R for Visual Analytics Chapter 3](https://r4va.netlify.app/chap03)