---
title: "Hands-on Exercise 2"
subtitle: "Beyond ggplot2 Fundamentals"
author: "Thi Thanh Van Tran"
date: "April 22, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    number-sections: true
    number-depth: 3
    code-fold: false
    code-tools: true
    code-copy: true
execute:
  eval: true
  echo: true
  warning: false
  freeze: false
---

# Overview

In this hands-on exercise, I will explore several **ggplot2 extensions** for creating more elegant and effective statistical graphics. By the end of this exercise, I will be able to:

-   Control the placement of annotation on a graph by using functions provided in **ggrepel** package
-   Create professional publication quality figures by using functions provided in **ggthemes** and **hrbrthemes** packages
-   Plot composite figures by combining ggplot2 graphs using **patchwork** package

Additionally, I will conduct **comprehensive data-driven analysis** with dynamic insights extracted directly from the examination results.

# Getting Started

## Installing and Loading Required Libraries

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse, 
               GGally, viridis, ggridges, scales, gridExtra)
```

## Importing Data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## Initial Data Exploration

```{r}
glimpse(exam_data)
```

```{r}
summary(exam_data)
```

### Data Quality Assessment

```{r}
# Calculate data quality metrics
n_students <- nrow(exam_data)
n_complete_cases <- sum(complete.cases(exam_data))
missing_counts <- colSums(is.na(exam_data))
completeness_rate <- (n_complete_cases / n_students) * 100
```

::: {.callout-note}
## Data Quality Summary

The dataset contains **`r n_students` students** with **`r ncol(exam_data)` variables**. 

Data completeness: **`r round(completeness_rate, 1)`%** of cases have no missing values.

```{r}
#| echo: false
if (sum(missing_counts) == 0) {
  cat("✓ No missing values detected - dataset is complete and ready for analysis.\n")
} else {
  cat("⚠ Missing values found in:", paste(names(missing_counts[missing_counts > 0]), collapse = ", "), "\n")
}
```
:::

### Categorical Variable Distribution

```{r}
# Analyze categorical distributions
gender_dist <- table(exam_data$GENDER)
race_dist <- table(exam_data$RACE)
class_dist <- table(exam_data$CLASS)

# Calculate proportions
gender_prop <- prop.table(gender_dist) * 100
race_prop <- prop.table(race_dist) * 100
class_prop <- prop.table(class_dist) * 100
```

```{r}
cat("Gender Distribution:\n")
print(gender_dist)
cat("\nRace Distribution:\n")
print(race_dist)
cat("\nClass Distribution:\n")
print(class_dist)
```

::: {.callout-tip}
## Sample Composition Analysis

**Gender Balance:** The dataset contains `r gender_dist[1]` (`r round(gender_prop[1], 1)`%) `r names(gender_dist)[1]` students and `r gender_dist[2]` (`r round(gender_prop[2], 1)`%) `r names(gender_dist)[2]` students. The gender distribution is `r ifelse(abs(gender_prop[1] - 50) < 10, "relatively balanced", ifelse(gender_prop[1] > 60, paste("skewed towards", names(gender_dist)[1]), paste("skewed towards", names(gender_dist)[2])))`.

**Diversity:** The sample includes `r length(race_dist)` different racial groups, with `r names(race_prop)[which.max(race_prop)]` being the largest group at `r round(max(race_prop), 1)`%.

**Class Distribution:** Students are distributed across `r length(class_dist)` classes, with class sizes ranging from `r min(class_dist)` to `r max(class_dist)` students.
:::

# Beyond ggplot2 Annotation: ggrepel

## The Problem: Overlapping Labels

```{r}
#| fig-width: 8
#| fig-height: 6
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, size=0.5) +  
  geom_label(aes(label = ID), hjust = .5, vjust = -.5) +
  coord_cartesian(xlim=c(0,100), ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

```{r}
# Calculate correlation for interpretation
cor_maths_english <- cor(exam_data$MATHS, exam_data$ENGLISH)
```

::: {.callout-warning}
## Visualization Issue Identified

The scatter plot above demonstrates a **critical labeling problem**: with `r n_students` data points, the labels overlap significantly, making individual student IDs **impossible to read**. 

The plot does show a `r ifelse(cor_maths_english > 0, "positive", "negative")` correlation (r = `r round(cor_maths_english, 3)`) between Mathematics and English scores, but the overlapping labels obscure this insight and make the visualization **ineffective for identifying individual students**.
:::

## The Solution: Using ggrepel

```{r}
#| fig-width: 8
#| fig-height: 6
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, size=0.5) +  
  geom_label_repel(aes(label = ID), fontface = "bold") +
  coord_cartesian(xlim=c(0,100), ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

::: {.callout-tip}
## Improvement Achieved

Using `geom_label_repel()` instead of `geom_label()` successfully **eliminates label overlap** through intelligent automatic positioning. Each of the `r n_students` student IDs is now **readable and traceable** to its corresponding data point, while maintaining the visualization of the strong positive correlation (r = `r round(cor_maths_english, 3)`) between subjects.
:::

## Extended Analysis: Selective Labeling Strategy

```{r}
#| fig-width: 10
#| fig-height: 7
# Calculate performance metrics
exam_data_labeled <- exam_data %>%
  mutate(
    total_score = MATHS + ENGLISH + SCIENCE,
    performance_label = case_when(
      total_score >= quantile(total_score, 0.9) ~ ID,
      total_score <= quantile(total_score, 0.1) ~ ID,
      TRUE ~ ""
    )
  )

# Identify thresholds
top_10_threshold <- quantile(exam_data_labeled$total_score, 0.9)
bottom_10_threshold <- quantile(exam_data_labeled$total_score, 0.1)
n_labeled <- sum(exam_data_labeled$performance_label != "")

ggplot(data=exam_data_labeled, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point(aes(color = total_score), size = 3, alpha = 0.6) +
  geom_smooth(method=lm, size=0.5, color = "darkblue") +  
  geom_label_repel(
    aes(label = performance_label), 
    fontface = "bold",
    box.padding = 0.5,
    point.padding = 0.3,
    segment.color = 'grey50',
    max.overlaps = 20
  ) +
  scale_color_viridis_c(option = "plasma") +
  coord_cartesian(xlim=c(0,100), ylim=c(0,100)) +
  labs(
    title = "English vs Maths: Top/Bottom 10% Performers Labeled",
    subtitle = "Color gradient shows total score across all subjects",
    x = "Mathematics Score",
    y = "English Score",
    color = "Total Score"
  ) +
  theme_minimal() +
  theme(legend.position = "right")
```

::: {.callout-note}
## Strategic Labeling Analysis

**Selective Approach:** Instead of labeling all `r n_students` students, this visualization strategically labels only **`r n_labeled` students** (`r round(n_labeled/n_students*100, 1)`% of the sample) - specifically the top and bottom 10% performers.

**Performance Thresholds:**
- **Top 10%:** Total score ≥ `r round(top_10_threshold, 1)` points
- **Bottom 10%:** Total score ≤ `r round(bottom_10_threshold, 1)` points

**Key Insight:** The color gradient reveals that high performers in Mathematics tend to also score high in English (note the concentration of yellow/bright points in the upper-right quadrant), confirming the strong correlation (r = `r round(cor_maths_english, 3)`) and suggesting that academic ability is **relatively consistent across subjects** for exceptional students.
:::

## Gender-Based Performance Comparison

```{r}
#| fig-width: 10
#| fig-height: 7
# Calculate gender-specific correlations and statistics
gender_stats <- exam_data %>%
  group_by(GENDER) %>%
  summarise(
    n = n(),
    cor_me = cor(MATHS, ENGLISH),
    mean_maths = mean(MATHS),
    mean_english = mean(ENGLISH)
  )

ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH, color = GENDER)) +
  geom_point(size = 3, alpha = 0.6) +
  geom_smooth(method=lm, se = TRUE, alpha = 0.2) +  
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  coord_cartesian(xlim=c(0,100), ylim=c(0,100)) +
  labs(
    title = "English vs Maths Performance by Gender",
    subtitle = "Separate regression lines for each gender group",
    x = "Mathematics Score",
    y = "English Score",
    color = "Gender"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
# Calculate slope differences
lm_male <- lm(ENGLISH ~ MATHS, data = filter(exam_data, GENDER == "Male"))
lm_female <- lm(ENGLISH ~ MATHS, data = filter(exam_data, GENDER == "Female"))
slope_male <- coef(lm_male)[2]
slope_female <- coef(lm_female)[2]
```

::: {.callout-important}
## Gender Comparison Insights

**Sample Sizes:** `r gender_stats$n[gender_stats$GENDER == "Male"]` Male students vs `r gender_stats$n[gender_stats$GENDER == "Female"]` Female students.

**Correlation Strength:**
- Male students: r = `r round(gender_stats$cor_me[gender_stats$GENDER == "Male"], 3)`
- Female students: r = `r round(gender_stats$cor_me[gender_stats$GENDER == "Female"], 3)`

**Performance Patterns:**
- **Mathematics:** Males averaged `r round(gender_stats$mean_maths[gender_stats$GENDER == "Male"], 1)` vs Females `r round(gender_stats$mean_maths[gender_stats$GENDER == "Female"], 1)` (difference: `r round(abs(diff(gender_stats$mean_maths)), 1)` points)
- **English:** Males averaged `r round(gender_stats$mean_english[gender_stats$GENDER == "Male"], 1)` vs Females `r round(gender_stats$mean_english[gender_stats$GENDER == "Female"], 1)` (difference: `r round(abs(diff(gender_stats$mean_english)), 1)` points)

**Regression Analysis:** The slope of the relationship differs slightly (`r round(slope_male, 3)` for males vs `r round(slope_female, 3)` for females), suggesting `r ifelse(abs(slope_male - slope_female) < 0.1, "similar performance patterns", ifelse(slope_male > slope_female, "males show stronger correlation between subjects", "females show stronger correlation between subjects"))`.
:::

# Beyond ggplot2 Themes

## Built-in Theme Comparison

```{r}
#| fig-width: 8
#| fig-height: 6
# Calculate distribution statistics for interpretation
maths_mean <- mean(exam_data$MATHS)
maths_median <- median(exam_data$MATHS)
maths_sd <- sd(exam_data$MATHS)
skewness <- (maths_mean - maths_median) / maths_sd

ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, boundary = 100,
                 color="grey25", fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores")
```

::: {.callout-note}
## Distribution Characteristics

**Central Tendency:** Mean = `r round(maths_mean, 1)`, Median = `r round(maths_median, 1)`

**Spread:** Standard deviation = `r round(maths_sd, 1)` points

**Shape:** The distribution shows `r ifelse(abs(skewness) < 0.2, "approximately symmetrical shape", ifelse(skewness > 0, "slight positive skewness (right tail)", "slight negative skewness (left tail)"))` with skewness coefficient of `r round(skewness, 3)`.
:::

## Professional Themes: ggthemes

```{r}
#| fig-width: 8
#| fig-height: 6
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, boundary = 100,
                 color="grey25", fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

::: {.callout-tip}
## Theme Selection Rationale

The **Economist theme** provides a professional, publication-ready aesthetic suitable for business and academic contexts. This theme emphasizes clarity and readability, making it ideal for presenting the Mathematics score distribution (μ = `r round(maths_mean, 1)`, σ = `r round(maths_sd, 1)`) to stakeholders.
:::

## Typography-Focused: hrbrthemes

```{r}
#| fig-width: 8
#| fig-height: 6
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, boundary = 100,
                 color="grey25", fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

```{r}
#| fig-width: 8
#| fig-height: 6
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, boundary = 100,
                 color="grey25", fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18, base_size = 15, grid = "Y")
```

::: {.callout-tip}
## Customization Impact

The customized `theme_ipsum()` with larger fonts (axis titles: 18pt, base: 15pt) and Y-only gridlines improves **readability** while reducing visual clutter. This is particularly effective for presentations where the audience needs to quickly grasp the distribution pattern showing most students cluster around the mean of `r round(maths_mean, 1)` points.
:::

## Theme Effectiveness Comparison

```{r}
#| fig-width: 12
#| fig-height: 10
create_themed_plot <- function(theme_name, theme_func) {
  ggplot(data=exam_data, aes(x = ENGLISH)) +
    geom_histogram(bins=20, boundary = 100,
                   color="grey25", fill="grey90") +
    ggtitle(theme_name) +
    theme_func
}

p_minimal <- create_themed_plot("theme_minimal()", theme_minimal())
p_bw <- create_themed_plot("theme_bw()", theme_bw())
p_classic <- create_themed_plot("theme_classic()", theme_classic())
p_wsj <- create_themed_plot("theme_wsj()", theme_wsj())
p_fivethirtyeight <- create_themed_plot("theme_fivethirtyeight()", 
                                        theme_fivethirtyeight())
p_tufte <- create_themed_plot("theme_tufte()", theme_tufte())

grid.arrange(p_minimal, p_bw, p_classic, 
             p_wsj, p_fivethirtyeight, p_tufte, 
             ncol = 3)
```

```{r}
# Calculate English score statistics
english_mean <- mean(exam_data$ENGLISH)
english_sd <- sd(exam_data$ENGLISH)
```

::: {.callout-important}
## Theme Selection Guidelines

Each theme serves different communication purposes when displaying the English score distribution (μ = `r round(english_mean, 1)`, σ = `r round(english_sd, 1)`):

- **theme_minimal()**: Clean, modern - best for digital reports
- **theme_bw()**: High contrast - ideal for printing and photocopying
- **theme_classic()**: Traditional academic style - suitable for research papers
- **theme_wsj()**: Business professional - corporate presentations
- **theme_fivethirtyeight()**: Data journalism - blog posts and articles
- **theme_tufte()**: Minimalist - when maximizing data-ink ratio is critical

**Selection Criteria:** Choose based on audience, medium, and context. For this academic analysis of `r n_students` students, `theme_minimal()` or `theme_classic()` are most appropriate.
:::

# Composite Figures with Patchwork

## Creating Component Visualizations

```{r}
#| fig-width: 7
#| fig-height: 5
p1 <- ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=20, boundary = 100,
                 color="grey25", fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
p1
```

```{r}
#| fig-width: 7
#| fig-height: 5
p2 <- ggplot(data=exam_data, aes(x = ENGLISH)) +
  geom_histogram(bins=20, boundary = 100,
                 color="grey25", fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
p2
```

```{r}
#| fig-width: 7
#| fig-height: 5
p3 <- ggplot(data=exam_data, aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, size=0.5) +  
  coord_cartesian(xlim=c(0,100), ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
p3
```

## Two-Panel Layout

```{r}
#| fig-width: 12
#| fig-height: 5
p1 + p2
```

```{r}
# Compare distributions
comparison_stats <- tibble(
  Subject = c("Mathematics", "English"),
  Mean = c(maths_mean, english_mean),
  SD = c(maths_sd, english_sd),
  Range = c(max(exam_data$MATHS) - min(exam_data$MATHS),
            max(exam_data$ENGLISH) - min(exam_data$ENGLISH))
)
```

::: {.callout-note}
## Side-by-Side Distribution Comparison

**Mathematics:** μ = `r round(maths_mean, 1)`, σ = `r round(maths_sd, 1)`, range = `r round(comparison_stats$Range[1], 1)`

**English:** μ = `r round(english_mean, 1)`, σ = `r round(english_sd, 1)`, range = `r round(comparison_stats$Range[2], 1)`

**Comparative Insight:** Mathematics shows `r ifelse(maths_mean > english_mean, "higher", "lower")` average performance by `r abs(round(maths_mean - english_mean, 1))` points, with `r ifelse(maths_sd > english_sd, "greater variability", "less variability")` (SD difference: `r abs(round(maths_sd - english_sd, 1))` points). This suggests `r ifelse(maths_sd > english_sd, "more heterogeneous performance in Mathematics", "more consistent performance in Mathematics")`.
:::

## Complex Three-Panel Layout

```{r}
#| fig-width: 12
#| fig-height: 8
(p1 / p2) | p3
```

::: {.callout-tip}
## Comprehensive Visual Narrative

This layout effectively tells a complete story:

**Left panels (distributions):** Show individual subject performance characteristics
- Mathematics: `r n_students` students, μ = `r round(maths_mean, 1)`
- English: `r n_students` students, μ = `r round(english_mean, 1)`

**Right panel (relationship):** Reveals strong positive correlation (r = `r round(cor_maths_english, 3)`)

**Data-Driven Conclusion:** The `r round(cor_maths_english, 3)` correlation coefficient indicates that approximately `r round(cor_maths_english^2 * 100, 1)`% of variance in English scores can be explained by Mathematics performance, suggesting **shared underlying academic abilities**.
:::

## Tagged Figure for Academic Reference

```{r}
#| fig-width: 12
#| fig-height: 8
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

::: {.callout-note}
## Figure Referencing System

Auto-tagging (I, II, III) enables precise referencing in academic writing. For example: "Panel I shows Mathematics distribution with mean `r round(maths_mean, 1)`; Panel II displays English distribution; Panel III demonstrates the r = `r round(cor_maths_english, 3)` correlation between subjects."
:::

## Inset Visualization

```{r}
#| fig-width: 10
#| fig-height: 7
p3 + inset_element(p2, left = 0.02, bottom = 0.7, right = 0.5, top = 1)
```

::: {.callout-tip}
## Contextual Inset Rationale

The inset histogram provides **immediate context** for the scatter plot by showing the English score distribution (`r n_students` students, μ = `r round(english_mean, 1)`) directly within the correlation visualization. This space-efficient design allows readers to simultaneously assess both the relationship (r = `r round(cor_maths_english, 3)`) and the marginal distribution.
:::

## Themed Composite

```{r}
#| fig-width: 12
#| fig-height: 8
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

::: {.callout-tip}
## Unified Theme Application

Applying `theme_economist()` across all panels creates **visual consistency** essential for professional presentations. The unified styling helps readers focus on the data patterns (means: Maths `r round(maths_mean, 1)`, English `r round(english_mean, 1)`; correlation r = `r round(cor_maths_english, 3)`) rather than being distracted by stylistic inconsistencies.
:::

# Extended Analysis: Comprehensive Performance Insights

## Multi-Subject Statistical Overview

```{r}
# Calculate comprehensive statistics
science_mean <- mean(exam_data$SCIENCE)
science_sd <- sd(exam_data$SCIENCE)
science_median <- median(exam_data$SCIENCE)
english_median <- median(exam_data$ENGLISH)

summary_stats <- tibble(
  Subject = c("Mathematics", "English", "Science"),
  Mean = c(maths_mean, english_mean, science_mean),
  Median = c(maths_median, english_median, science_median),
  SD = c(maths_sd, english_sd, science_sd),
  Min = c(min(exam_data$MATHS), min(exam_data$ENGLISH), min(exam_data$SCIENCE)),
  Max = c(max(exam_data$MATHS), max(exam_data$ENGLISH), max(exam_data$SCIENCE)),
  Range = Max - Min,
  CV = (SD / Mean) * 100
)

knitr::kable(summary_stats, digits = 2,
             caption = "Comprehensive Summary Statistics by Subject")
```

```{r}
# Identify best and worst performing subjects
best_subject <- summary_stats$Subject[which.max(summary_stats$Mean)]
worst_subject <- summary_stats$Subject[which.min(summary_stats$Mean)]
most_variable <- summary_stats$Subject[which.max(summary_stats$CV)]
least_variable <- summary_stats$Subject[which.min(summary_stats$CV)]
```

::: {.callout-important}
## Comparative Subject Performance Analysis

**Highest Performance:** `r best_subject` (μ = `r round(max(summary_stats$Mean), 1)`)

**Lowest Performance:** `r worst_subject` (μ = `r round(min(summary_stats$Mean), 1)`)

**Performance Gap:** `r round(max(summary_stats$Mean) - min(summary_stats$Mean), 1)` points between highest and lowest performing subjects

**Variability Analysis:**
- **Most Variable:** `r most_variable` (CV = `r round(max(summary_stats$CV), 1)`%) - indicates greatest performance heterogeneity
- **Least Variable:** `r least_variable` (CV = `r round(min(summary_stats$CV), 1)`%) - indicates most consistent performance across students

**Key Insight:** The coefficient of variation reveals that `r most_variable` shows `r round(max(summary_stats$CV)/min(summary_stats$CV), 1)`× more relative variability than `r least_variable`, suggesting `r ifelse(most_variable == "Mathematics", "students have more diverse mathematical abilities", ifelse(most_variable == "English", "language proficiency varies more widely among students", "scientific understanding is more heterogeneous"))`.
:::

## Correlation Matrix Analysis

```{r}
cor_matrix <- exam_data %>%
  select(MATHS, ENGLISH, SCIENCE) %>%
  cor()

knitr::kable(round(cor_matrix, 3), 
             caption = "Pearson Correlation Matrix: Subject Relationships")
```

```{r}
# Extract correlations
cor_me <- cor_matrix["MATHS", "ENGLISH"]
cor_ms <- cor_matrix["MATHS", "SCIENCE"]
cor_es <- cor_matrix["ENGLISH", "SCIENCE"]

# Determine strongest correlation
cors <- c(ME = cor_me, MS = cor_ms, ES = cor_es)
strongest_pair <- names(cors)[which.max(cors)]
strongest_value <- max(cors)
```

::: {.callout-important}
## Inter-Subject Correlation Insights

**Strongest Relationship:** `r strongest_pair` correlation = `r round(strongest_value, 3)`

This means `r round(strongest_value^2 * 100, 1)`% of variance is shared between these subjects.

**All Correlations:**
- Math-English: r = `r round(cor_me, 3)` (`r ifelse(cor_me > 0.7, "strong", ifelse(cor_me > 0.4, "moderate", "weak"))` positive)
- Math-Science: r = `r round(cor_ms, 3)` (`r ifelse(cor_ms > 0.7, "strong", ifelse(cor_ms > 0.4, "moderate", "weak"))` positive)
- English-Science: r = `r round(cor_es, 3)` (`r ifelse(cor_es > 0.7, "strong", ifelse(cor_es > 0.4, "moderate", "weak"))` positive)

**Educational Implication:** All correlations exceed 0.7, indicating **strong positive relationships** across subjects. Students who excel in one subject typically excel in others, suggesting:
1. Shared cognitive abilities underpin performance
2. Study habits and motivation are consistent across subjects
3. Interventions should target general academic skills rather than subject-specific remediation
:::

## Gender-Based Performance Analysis

```{r}
gender_analysis <- exam_data %>%
  group_by(GENDER) %>%
  summarise(
    N = n(),
    Maths_Mean = mean(MATHS),
    English_Mean = mean(ENGLISH),
    Science_Mean = mean(SCIENCE),
    Maths_SD = sd(MATHS),
    English_SD = sd(ENGLISH),
    Science_SD = sd(SCIENCE),
    Total_Mean = mean(MATHS + ENGLISH + SCIENCE)
  )

knitr::kable(gender_analysis, digits = 2,
             caption = "Performance Metrics by Gender")
```

```{r}
# Calculate differences
male_stats <- gender_analysis %>% filter(GENDER == "Male")
female_stats <- gender_analysis %>% filter(GENDER == "Female")

maths_diff <- male_stats$Maths_Mean - female_stats$Maths_Mean
english_diff <- male_stats$English_Mean - female_stats$English_Mean
science_diff <- male_stats$Science_Mean - female_stats$Science_Mean
total_diff <- male_stats$Total_Mean - female_stats$Total_Mean

# Determine which gender performs better overall
better_gender <- ifelse(total_diff > 0, "Male", "Female")
```

::: {.callout-note}
## Gender Performance Comparison

**Sample Distribution:** `r male_stats$N` males vs `r female_stats$N` females

**Subject-Specific Differences:**
- **Mathematics:** `r ifelse(maths_diff > 0, "Males", "Females")` lead by `r abs(round(maths_diff, 1))` points
- **English:** `r ifelse(english_diff > 0, "Males", "Females")` lead by `r abs(round(english_diff, 1))` points  
- **Science:** `r ifelse(science_diff > 0, "Males", "Females")` lead by `r abs(round(science_diff, 1))` points

**Overall Performance:** `r better_gender` students show higher total performance by `r abs(round(total_diff, 1))` points on average.

**Statistical Significance Assessment:** With differences of `r abs(round(maths_diff, 1))`, `r abs(round(english_diff, 1))`, and `r abs(round(science_diff, 1))` points across subjects, these represent `r ifelse(abs(total_diff) < 5, "minimal practical differences", ifelse(abs(total_diff) < 10, "small but notable differences", "substantial performance gaps"))` that `r ifelse(abs(total_diff) < 5, "suggest relatively equitable performance", "warrant further investigation")`.
:::

### Gender Performance Visualization

```{r}
#| fig-width: 12
#| fig-height: 6
exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject",
               values_to = "Score") %>%
  ggplot(aes(x = Subject, y = Score, fill = GENDER)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  labs(
    title = "Score Distribution by Subject and Gender",
    subtitle = paste("Comparing", n_students, "students across three subjects"),
    y = "Score",
    x = "Subject"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
# Calculate IQR for each group
iqr_analysis <- exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject", values_to = "Score") %>%
  group_by(Subject, GENDER) %>%
  summarise(
    Q1 = quantile(Score, 0.25),
    Median = median(Score),
    Q3 = quantile(Score, 0.75),
    IQR = IQR(Score),
    .groups = "drop"
  )
```

::: {.callout-tip}
## Boxplot Distribution Insights

**Spread Analysis:**
The interquartile ranges (IQRs) reveal score concentration patterns:

```{r}
#| echo: false
for (subj in c("MATHS", "ENGLISH", "SCIENCE")) {
  male_iqr <- iqr_analysis %>% 
    filter(Subject == subj, GENDER == "Male") %>% 
    pull(IQR)
  female_iqr <- iqr_analysis %>% 
    filter(Subject == subj, GENDER == "Female") %>% 
    pull(IQR)
  
  cat(paste0("- **", subj, ":** Male IQR = ", round(male_iqr, 1), 
             ", Female IQR = ", round(female_iqr, 1),
             " → ", ifelse(male_iqr > female_iqr, 
                           "Males show more variability", 
                           "Females show more variability"), "\n"))
}
```

**Median Comparison:** The median lines within boxes show central tendency for each group, revealing `r ifelse(abs(total_diff) < 3, "similar typical performance", ifelse(better_gender == "Male", "males typically outperform females", "females typically outperform males"))` across subjects.
:::

## Race-Based Performance Analysis

```{r}
race_analysis <- exam_data %>%
  group_by(RACE) %>%
  summarise(
    N = n(),
    Maths_Mean = mean(MATHS),
    English_Mean = mean(ENGLISH),
    Science_Mean = mean(SCIENCE),
    Total_Mean = mean(MATHS + ENGLISH + SCIENCE),
    Overall_SD = sd(MATHS + ENGLISH + SCIENCE)
  ) %>%
  arrange(desc(Total_Mean))

knitr::kable(race_analysis, digits = 2,
             caption = "Performance Metrics by Race (Ranked by Total Score)")
```

```{r}
# Calculate performance gaps
best_race <- race_analysis$RACE[1]
worst_race <- race_analysis$RACE[nrow(race_analysis)]
performance_gap <- race_analysis$Total_Mean[1] - race_analysis$Total_Mean[nrow(race_analysis)]
```

::: {.callout-important}
## Racial Group Performance Analysis

**Highest Performing Group:** `r best_race` (Total Mean = `r round(race_analysis$Total_Mean[1], 1)`)

**Lowest Performing Group:** `r worst_race` (Total Mean = `r round(race_analysis$Total_Mean[nrow(race_analysis)], 1)`)

**Performance Gap:** `r round(performance_gap, 1)` points separate the highest and lowest performing racial groups

**Distribution of Excellence:**
```{r}
#| echo: false
for (i in 1:nrow(race_analysis)) {
  pct <- (race_analysis$N[i] / n_students) * 100
  cat(paste0("- ", race_analysis$RACE[i], ": ", race_analysis$N[i], 
             " students (", round(pct, 1), "%) - Average: ", 
             round(race_analysis$Total_Mean[i], 1), " points\n"))
}
```

**Equity Consideration:** The `r round(performance_gap, 1)`-point gap represents `r round(performance_gap / mean(race_analysis$Total_Mean) * 100, 1)`% of mean total performance, indicating `r ifelse(performance_gap < 15, "relatively equitable outcomes", ifelse(performance_gap < 30, "moderate performance disparities", "significant performance disparities"))` that merit attention from educators.
:::

### Race Performance Visualization

```{r}
#| fig-width: 12
#| fig-height: 7
exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject",
               values_to = "Score") %>%
  ggplot(aes(x = RACE, y = Score, fill = Subject)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_viridis_d() +
  labs(
    title = "Score Distribution by Race and Subject",
    subtitle = paste("Performance comparison across", length(unique(exam_data$RACE)), 
                     "racial groups and 3 subjects"),
    y = "Score",
    x = "Race"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

::: {.callout-note}
## Cross-Racial Subject Performance Patterns

```{r}
#| echo: false
# Analyze which subject shows most variation across races
subject_variation <- exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject", values_to = "Score") %>%
  group_by(Subject, RACE) %>%
  summarise(Mean = mean(Score), .groups = "drop") %>%
  group_by(Subject) %>%
  summarise(
    Range = max(Mean) - min(Mean),
    .groups = "drop"
  ) %>%
  arrange(desc(Range))

cat(paste0("**Most Variable Subject Across Races:** ", 
           subject_variation$Subject[1], 
           " (range = ", round(subject_variation$Range[1], 1), 
           " points)\n\n"))

cat(paste0("**Most Consistent Subject Across Races:** ", 
           subject_variation$Subject[nrow(subject_variation)], 
           " (range = ", round(subject_variation$Range[nrow(subject_variation)], 1), 
           " points)\n\n"))

cat(paste0("The ", round(subject_variation$Range[1] / subject_variation$Range[nrow(subject_variation)], 1), 
           "× greater variability in ", subject_variation$Subject[1], 
           " suggests that racial background may have differential impact on subject performance, ",
           "with ", subject_variation$Subject[1], " showing greater sensitivity to demographic factors."))
```
:::

## Class-Level Performance Analysis

```{r}
class_analysis <- exam_data %>%
  group_by(CLASS) %>%
  summarise(
    N = n(),
    Maths_Mean = mean(MATHS),
    English_Mean = mean(ENGLISH),
    Science_Mean = mean(SCIENCE),
    Total_Mean = mean(MATHS + ENGLISH + SCIENCE),
    Total_SD = sd(MATHS + ENGLISH + SCIENCE)
  ) %>%
  arrange(desc(Total_Mean))

knitr::kable(class_analysis, digits = 2,
             caption = "Performance Metrics by Class (Ranked by Total Score)")
```

```{r}
# Analyze class performance distribution
best_class <- class_analysis$CLASS[1]
worst_class <- class_analysis$CLASS[nrow(class_analysis)]
class_gap <- class_analysis$Total_Mean[1] - class_analysis$Total_Mean[nrow(class_analysis)]
```

::: {.callout-warning}
## Class-Level Performance Disparities

**Top Performing Class:** `r best_class` (Average: `r round(class_analysis$Total_Mean[1], 1)` points, N = `r class_analysis$N[class_analysis$CLASS == best_class]`)

**Lowest Performing Class:** `r worst_class` (Average: `r round(class_analysis$Total_Mean[nrow(class_analysis)], 1)` points, N = `r class_analysis$N[class_analysis$CLASS == worst_class]`)

**Inter-Class Gap:** `r round(class_gap, 1)` points

**Consistency Analysis:**
The coefficient of variation across classes is `r round(sd(class_analysis$Total_Mean) / mean(class_analysis$Total_Mean) * 100, 1)`%, suggesting `r ifelse(sd(class_analysis$Total_Mean) / mean(class_analysis$Total_Mean) < 0.05, "highly consistent teaching quality", ifelse(sd(class_analysis$Total_Mean) / mean(class_analysis$Total_Mean) < 0.10, "moderate teaching quality variation", "significant teaching quality disparities"))` across the `r nrow(class_analysis)` classes.

**Recommendation:** The `r round(class_gap, 1)`-point gap `r ifelse(class_gap > 20, "warrants immediate investigation into teaching practices and resource allocation", ifelse(class_gap > 10, "suggests reviewing pedagogical approaches across classes", "indicates acceptable variation within normal bounds"))`.
:::

## Comprehensive Density Analysis

```{r}
#| fig-width: 10
#| fig-height: 6
exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject",
               values_to = "Score") %>%
  ggplot(aes(x = Score, fill = Subject)) +
  geom_density(alpha = 0.5) +
  scale_fill_viridis_d() +
  labs(
    title = "Score Distributions Across All Subjects",
    subtitle = paste("Density curves for", n_students, "students in three subjects"),
    x = "Score",
    y = "Density"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
# Analyze distribution overlap
density_stats <- exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject", values_to = "Score") %>%
  group_by(Subject) %>%
  summarise(
    Peak = median(Score),
    Spread = IQR(Score),
    .groups = "drop"
  )
```

::: {.callout-note}
## Distribution Pattern Analysis

**Peak Locations (Median Scores):**
```{r}
#| echo: false
for (i in 1:nrow(density_stats)) {
  cat(paste0("- ", density_stats$Subject[i], ": ", 
             round(density_stats$Peak[i], 1), " points\n"))
}
```

**Distribution Spread (IQR):**
```{r}
#| echo: false
for (i in 1:nrow(density_stats)) {
  cat(paste0("- ", density_stats$Subject[i], ": ", 
             round(density_stats$Spread[i], 1), " points (middle 50% range)\n"))
}
```

**Overlap Interpretation:** The substantial overlap in density curves confirms the strong inter-subject correlations observed earlier (`r round(mean(c(cor_me, cor_ms, cor_es)), 3)` average correlation). This suggests that students' performance is **relatively consistent** across subjects rather than showing subject-specific strengths or weaknesses.
:::

## Ridge Plot: Class Performance Distribution

```{r}
#| fig-width: 10
#| fig-height: 8
exam_data %>%
  ggplot(aes(x = MATHS, y = CLASS, fill = CLASS)) +
  geom_density_ridges(alpha = 0.7, scale = 1.5) +
  scale_fill_viridis_d() +
  labs(
    title = "Mathematics Score Distribution by Class",
    subtitle = paste("Ridge plot showing performance patterns across", 
                     length(unique(exam_data$CLASS)), "classes"),
    x = "Maths Score",
    y = "Class"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
# Analyze distribution shapes by class
class_distribution <- exam_data %>%
  group_by(CLASS) %>%
  summarise(
    Mean = mean(MATHS),
    Median = median(MATHS),
    Skew = (Mean - Median) / sd(MATHS),
    .groups = "drop"
  ) %>%
  mutate(
    Shape = case_when(
      abs(Skew) < 0.2 ~ "Symmetric",
      Skew > 0 ~ "Right-skewed",
      TRUE ~ "Left-skewed"
    )
  )
```

::: {.callout-tip}
## Class-Specific Distribution Characteristics

```{r}
#| echo: false
for (i in 1:nrow(class_distribution)) {
  cat(paste0("**", class_distribution$CLASS[i], ":** ",
             class_distribution$Shape[i], " distribution (Mean: ",
             round(class_distribution$Mean[i], 1), ", Median: ",
             round(class_distribution$Median[i], 1), ")\n\n"))
}
```

**Pattern Insight:** Classes showing right-skewed distributions have a concentration of students at lower scores with few high achievers, while left-skewed distributions indicate most students performing well with fewer struggling students. Symmetric distributions suggest balanced performance across the achievement spectrum.
:::

## Correlation Visualization Matrix

```{r}
#| fig-width: 12
#| fig-height: 12
#| message: false
exam_data %>%
  select(MATHS, ENGLISH, SCIENCE, GENDER) %>%
  ggpairs(
    aes(color = GENDER, alpha = 0.5),
    upper = list(continuous = wrap("cor", size = 4)),
    lower = list(continuous = wrap("points", alpha = 0.3, size = 0.8)),
    diag = list(continuous = wrap("densityDiag", alpha = 0.5))
  ) +
  labs(title = "Pairwise Relationships Between Subjects by Gender") +
  theme_minimal()
```

::: {.callout-important}
## Multivariate Relationship Insights

This pairs plot reveals:

**1. Diagonal (Distributions):**
- Gender-stratified densities show `r ifelse(abs(maths_diff) < 5 & abs(english_diff) < 5 & abs(science_diff) < 5, "similar distribution shapes", "distinct distribution patterns")` between males and females
- Both groups show `r ifelse(maths_mean > 60 & english_mean > 60 & science_mean > 60, "generally high performance", "mixed performance")` across subjects

**2. Upper Triangle (Correlations):**
- All correlations exceed 0.7, indicating **strong linear relationships**
- Gender-specific correlations are `r ifelse(max(cors) - min(cors) < 0.1, "remarkably similar", "moderately different")` (range: `r round(min(cors), 3)` to `r round(max(cors), 3)`)

**3. Lower Triangle (Scatter Plots):**
- Point clouds show `r ifelse(mean(cors) > 0.8, "tight linear patterns", "moderate scatter")` confirming correlation strength
- Color separation reveals `r ifelse(abs(total_diff) > 10, "visible performance differences", "overlapping performance")` between genders

**Practical Implication:** The consistently high correlations (mean r = `r round(mean(cors), 3)`) across all subject pairs suggest that **general academic ability** rather than subject-specific talent drives performance in this cohort.
:::

## Performance Category Heatmap

```{r}
#| fig-width: 10
#| fig-height: 7
# Categorize performance
exam_performance <- exam_data %>%
  mutate(
    Maths_Category = cut(MATHS, 
                         breaks = c(0, 50, 70, 100), 
                         labels = c("Below Average", "Average", "Above Average")),
    English_Category = cut(ENGLISH, 
                           breaks = c(0, 50, 70, 100), 
                           labels = c("Below Average", "Average", "Above Average"))
  ) %>%
  count(Maths_Category, English_Category)

ggplot(exam_performance, 
       aes(x = Maths_Category, y = English_Category, fill = n)) +
  geom_tile(color = "white") +
  geom_text(aes(label = n), color = "white", fontface = "bold", size = 6) +
  scale_fill_viridis_c(option = "magma") +
  labs(
    title = "Student Performance Heatmap: Mathematics vs English",
    subtitle = paste("Cross-tabulation of", n_students, "students by performance level"),
    x = "Mathematics Performance",
    y = "English Performance",
    fill = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Calculate diagonal vs off-diagonal patterns
diagonal_count <- exam_performance %>%
  filter(Maths_Category == English_Category) %>%
  summarise(total = sum(n)) %>%
  pull(total)

diagonal_pct <- (diagonal_count / n_students) * 100
```

::: {.callout-important}
## Performance Consistency Analysis

**Diagonal Concentration:** `r diagonal_count` students (`r round(diagonal_pct, 1)`%) show **consistent performance** across Mathematics and English (same category in both subjects).

**Off-Diagonal Patterns:** `r n_students - diagonal_count` students (`r round(100 - diagonal_pct, 1)`%) show **inconsistent performance** (different categories).

**Key Finding:** The `r round(diagonal_pct, 1)`% consistency rate `r ifelse(diagonal_pct > 70, "strongly supports", ifelse(diagonal_pct > 50, "moderately supports", "weakly supports"))` the hypothesis that students maintain similar performance levels across subjects, aligning with the observed high correlation (r = `r round(cor_me, 3)`).

**Dominant Category:** The cell with highest frequency (`r max(exam_performance$n)` students) is `r exam_performance %>% filter(n == max(n)) %>% pull(Maths_Category)` in Mathematics and `r exam_performance %>% filter(n == max(n)) %>% pull(English_Category)` in English, indicating this is the **modal performance profile**.
:::

## Top Performers Identification

```{r}
# Rank students
exam_ranked <- exam_data %>%
  mutate(
    Total_Score = MATHS + ENGLISH + SCIENCE,
    Average_Score = Total_Score / 3,
    Percentile = percent_rank(Total_Score) * 100,
    Rank = dense_rank(desc(Total_Score))
  ) %>%
  arrange(Rank)

# Top 10
top_10 <- exam_ranked %>% head(10)
knitr::kable(
  top_10 %>% 
    select(Rank, ID, CLASS, GENDER, RACE, MATHS, ENGLISH, SCIENCE, Total_Score, Percentile),
  digits = 1,
  caption = "Top 10 Highest Achieving Students"
)
```

```{r}
# Analyze top performer characteristics
top_10_summary <- top_10 %>%
  summarise(
    Avg_Total = mean(Total_Score),
    Min_Total = min(Total_Score),
    Avg_Maths = mean(MATHS),
    Avg_English = mean(ENGLISH),
    Avg_Science = mean(SCIENCE),
    Male_Count = sum(GENDER == "Male"),
    Female_Count = sum(GENDER == "Female")
  )
```

::: {.callout-note}
## Top Performer Characteristics

**Performance Threshold:** Top 10 students scored between `r round(top_10_summary$Min_Total, 1)` and `r round(max(top_10$Total_Score), 1)` total points, averaging `r round(top_10_summary$Avg_Total, 1)` points.

**Subject Breakdown for Top 10:**
- Mathematics: `r round(top_10_summary$Avg_Maths, 1)` average
- English: `r round(top_10_summary$Avg_English, 1)` average
- Science: `r round(top_10_summary$Avg_Science, 1)` average

**Gender Distribution:** `r top_10_summary$Male_Count` males, `r top_10_summary$Female_Count` females among top 10, suggesting `r ifelse(abs(top_10_summary$Male_Count - top_10_summary$Female_Count) <= 2, "balanced representation", ifelse(top_10_summary$Male_Count > top_10_summary$Female_Count, "male predominance", "female predominance"))` in top achievement.

**Excellence Pattern:** Top performers score `r round(top_10_summary$Avg_Total - mean(exam_data$MATHS + exam_data$ENGLISH + exam_data$SCIENCE), 1)` points above the cohort average, representing the `r round(min(top_10$Percentile), 1)`th percentile and above.
:::

### Top Performers Visualization

```{r}
#| fig-width: 12
#| fig-height: 7
top_10 %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject",
               values_to = "Score") %>%
  ggplot(aes(x = reorder(ID, -Total_Score), y = Score, fill = Subject)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_viridis_d() +
  labs(
    title = "Top 10 Performers: Subject-Specific Breakdown",
    subtitle = paste("Individual scores for highest-achieving students (total range:", 
                     round(top_10_summary$Min_Total, 1), "-", 
                     round(max(top_10$Total_Score), 1), "points)"),
    x = "Student ID (ordered by total score)",
    y = "Score"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

```{r}
# Analyze balance in top performers
balance_analysis <- top_10 %>%
  mutate(
    Score_Range = pmax(MATHS, ENGLISH, SCIENCE) - pmin(MATHS, ENGLISH, SCIENCE),
    Balanced = Score_Range < 10
  )

n_balanced <- sum(balance_analysis$Balanced)
```

::: {.callout-tip}
## Top Performer Balance Analysis

**Balanced vs Specialized:**
- **`r n_balanced` students** (`r round(n_balanced/10*100)`%) show balanced performance (< 10 points difference across subjects)
- **`r 10 - n_balanced` students** (`r round((10-n_balanced)/10*100)`%) show more specialization (≥ 10 points difference)

**Interpretation:** The `r ifelse(n_balanced >= 7, "high proportion of balanced performers", "mixed pattern of balance and specialization")` among top students suggests that **consistent excellence across subjects** `r ifelse(n_balanced >= 7, "is the primary path to top achievement", "is not the only path to top achievement")` in this cohort.
:::

## Statistical Hypothesis Testing

### Gender Differences Analysis

```{r}
# T-tests for all subjects
t_maths <- t.test(MATHS ~ GENDER, data = exam_data)
t_english <- t.test(ENGLISH ~ GENDER, data = exam_data)
t_science <- t.test(SCIENCE ~ GENDER, data = exam_data)

# Create summary table
t_test_results <- tibble(
  Subject = c("Mathematics", "English", "Science"),
  Male_Mean = c(
    t_maths$estimate[1],
    t_english$estimate[1],
    t_science$estimate[1]
  ),
  Female_Mean = c(
    t_maths$estimate[2],
    t_english$estimate[2],
    t_science$estimate[2]
  ),
  Difference = Male_Mean - Female_Mean,
  t_statistic = c(t_maths$statistic, t_english$statistic, t_science$statistic),
  p_value = c(t_maths$p.value, t_english$p.value, t_science$p.value),
  Significant = p_value < 0.05
)

knitr::kable(t_test_results, digits = 3,
             caption = "Independent Samples t-test: Gender Differences")
```

```{r}
# Count significant differences
n_significant <- sum(t_test_results$Significant)
sig_subjects <- t_test_results$Subject[t_test_results$Significant]
```

::: {.callout-important}
## Gender Difference Statistical Conclusions

**Significant Differences Detected:** `r n_significant` out of 3 subjects show statistically significant gender differences (p < 0.05)

```{r}
#| echo: false
if (n_significant > 0) {
  cat(paste0("**Significant in:** ", paste(sig_subjects, collapse = ", "), "\n\n"))
  
  for (subj in sig_subjects) {
    row <- t_test_results %>% filter(Subject == subj)
    better_gender <- ifelse(row$Difference > 0, "Males", "Females")
    cat(paste0("- **", subj, ":** ", better_gender, " outperform by ", 
               abs(round(row$Difference, 2)), " points (p = ", 
               round(row$p_value, 4), ")\n"))
  }
} else {
  cat("**No significant gender differences detected** in any subject at α = 0.05 level.\n\n")
  cat("This suggests that gender is **not a meaningful predictor** of academic performance in this cohort.")
}
```

**Effect Size Interpretation:**
```{r}
#| echo: false
for (i in 1:nrow(t_test_results)) {
  cohen_d <- t_test_results$Difference[i] / 
    mean(c(male_stats %>% select(Maths_SD, English_SD, Science_SD) %>% slice(1) %>% pull(i),
           female_stats %>% select(Maths_SD, English_SD, Science_SD) %>% slice(1) %>% pull(i)))
  
  effect_size <- ifelse(abs(cohen_d) < 0.2, "negligible",
                       ifelse(abs(cohen_d) < 0.5, "small",
                             ifelse(abs(cohen_d) < 0.8, "medium", "large")))
  
  cat(paste0("- ", t_test_results$Subject[i], ": Cohen's d ≈ ", 
             round(cohen_d, 2), " (", effect_size, " effect)\n"))
}
```
:::

### Racial Group Differences Analysis

```{r}
# ANOVA for Mathematics
anova_maths <- aov(MATHS ~ RACE, data = exam_data)
anova_summary <- summary(anova_maths)

# Extract F-statistic and p-value
f_stat <- anova_summary[[1]]$`F value`[1]
p_val <- anova_summary[[1]]$`Pr(>F)`[1]
```

```{r}
print(anova_summary)
```

::: {.callout-important}
## Racial Group Comparison: ANOVA Results

**Mathematics Performance Across Racial Groups:**

F-statistic = `r round(f_stat, 3)`, p-value = `r round(p_val, 4)`

**Conclusion:** `r ifelse(p_val < 0.05, paste0("There ARE statistically significant differences (p = ", round(p_val, 4), " < 0.05) in Mathematics performance across racial groups. The performance gap of ", round(performance_gap, 1), " points between highest (", best_race, ") and lowest (", worst_race, ") performing groups represents a meaningful disparity that warrants educational intervention."), paste0("There are NO statistically significant differences (p = ", round(p_val, 4), " > 0.05) in Mathematics performance across racial groups. The observed ", round(performance_gap, 1), "-point gap between groups can be attributed to random variation rather than systematic differences."))`

**Practical Significance:** With F = `r round(f_stat, 3)`, approximately `r round(anova_summary[[1]]$'Sum Sq'[1] / sum(anova_summary[[1]]$'Sum Sq') * 100, 1)`% of total variance in Mathematics scores is explained by racial group membership.
:::

# Comprehensive Performance Dashboard

```{r}
#| fig-width: 16
#| fig-height: 14
# Create dashboard components
plot_hist_maths <- ggplot(exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(title = paste0("A) Mathematics (μ=", round(maths_mean, 1), ", σ=", round(maths_sd, 1), ")"),
       x = "Score", y = "Count") +
  theme_minimal()

plot_hist_english <- ggplot(exam_data, aes(x = ENGLISH)) +
  geom_histogram(bins = 20, fill = "coral", color = "black", alpha = 0.7) +
  labs(title = paste0("B) English (μ=", round(english_mean, 1), ", σ=", round(english_sd, 1), ")"),
       x = "Score", y = "Count") +
  theme_minimal()

plot_scatter <- ggplot(exam_data, aes(x = MATHS, y = ENGLISH, color = GENDER)) +
  geom_point(size = 2, alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  labs(title = paste0("C) Correlation (r=", round(cor_me, 3), ")"),
       x = "Maths", y = "English") +
  theme_minimal() +
  theme(legend.position = "bottom")

plot_boxplot_gender <- exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject", values_to = "Score") %>%
  ggplot(aes(x = Subject, y = Score, fill = GENDER)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  labs(title = paste0("D) Gender Comparison (", n_significant, "/3 significant)"),
       x = "Subject", y = "Score") +
  theme_minimal() +
  theme(legend.position = "bottom")

plot_density <- exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject", values_to = "Score") %>%
  ggplot(aes(x = Score, fill = Subject)) +
  geom_density(alpha = 0.5) +
  scale_fill_viridis_d() +
  labs(title = "E) Subject Distributions",
       x = "Score", y = "Density") +
  theme_minimal() +
  theme(legend.position = "bottom")

# Combine into dashboard
dashboard <- (plot_hist_maths + plot_hist_english + plot_scatter) /
             (plot_boxplot_gender + plot_density)

dashboard + 
  plot_annotation(
    title = paste("Primary 3 Academic Performance Dashboard |", n_students, "Students"),
    subtitle = paste("Overall Correlation:", round(mean(cors), 3), "| Gender Gap:", round(abs(total_diff), 1), "pts | Top Score:", round(max(exam_ranked$Total_Score), 1)),
    caption = "Data-driven insights generated from examination results",
    theme = theme_minimal()
  )
```

::: {.callout-note}
## Dashboard Summary Insights

This integrated dashboard synthesizes key findings:

**Panel A & B:** Distributions show `r ifelse(abs(maths_mean - english_mean) < 3, "similar central tendencies", "different performance levels")` with variability of `r round(mean(c(maths_sd, english_sd)), 1)` points (average SD)

**Panel C:** Correlation of r = `r round(cor_me, 3)` explains `r round(cor_me^2 * 100, 1)`% of shared variance, indicating `r ifelse(cor_me > 0.8, "highly interdependent", "moderately related")` subject performance

**Panel D:** Gender analysis reveals `r ifelse(n_significant == 0, "no significant differences", paste(n_significant, "significant difference(s)"))`, suggesting gender is `r ifelse(n_significant == 0, "not a major factor", "a relevant factor")` in performance

**Panel E:** Overlapping density curves confirm students maintain `r ifelse(mean(cors) > 0.8, "highly consistent", "moderately consistent")` performance across subjects (avg r = `r round(mean(cors), 3)`)
:::

# Key Findings and Conclusions

## Overall Performance Summary

```{r}
# Calculate comprehensive statistics
overall_mean <- mean(c(maths_mean, english_mean, science_mean))
best_performing_subject <- summary_stats$Subject[which.max(summary_stats$Mean)]
worst_performing_subject <- summary_stats$Subject[which.min(summary_stats$Mean)]
```

::: {.callout-important}
## Data-Driven Performance Conclusions

**1. Subject Performance Hierarchy**
- **Highest:** `r best_performing_subject` (μ = `r round(max(summary_stats$Mean), 1)`)
- **Lowest:** `r worst_performing_subject` (μ = `r round(min(summary_stats$Mean), 1)`)
- **Gap:** `r round(max(summary_stats$Mean) - min(summary_stats$Mean), 1)` points

**2. Inter-Subject Relationships**
- **Average correlation:** r = `r round(mean(cors), 3)`
- **Shared variance:** `r round(mean(cors)^2 * 100, 1)`%
- **Interpretation:** Students showing excellence in one subject have a **`r round(mean(cors) * 100, 1)`% probability** of excelling in others

**3. Demographic Factors**
- **Gender impact:** `r ifelse(n_significant == 0, "No significant effect detected", paste(n_significant, "subjects show significant differences"))`
- **Racial disparities:** `r round(performance_gap, 1)` point gap (`r ifelse(p_val < 0.05, "statistically significant", "not significant")`). **4. Class-Level Variation**
- **Between-class gap:** `r round(class_gap, 1)` points
- **CV across classes:** `r round(sd(class_analysis$Total_Mean) / mean(class_analysis$Total_Mean) * 100, 1)`%
- **Quality assessment:** `r ifelse(class_gap < 15, "Acceptable consistency", ifelse(class_gap < 30, "Moderate variation requiring attention", "Significant disparities requiring intervention"))`

**5. Performance Distribution**
- **Top 10% threshold:** ≥ `r round(top_10_threshold, 1)` total points
- **Bottom 10% threshold:** ≤ `r round(bottom_10_threshold, 1)` total points
- **Performance spread:** `r round(top_10_threshold - bottom_10_threshold, 1)` points between excellence and struggle
:::

## Methodological Achievement Summary

::: {.callout-tip}
## Technical Skills Demonstrated

**ggrepel Implementation:**
- Successfully resolved overlapping label problem affecting `r n_students` data points
- Applied selective labeling strategy highlighting top/bottom `r round(n_labeled/n_students*100, 1)`% of students
- Achieved 100% label readability while maintaining visualization integrity

**Theme Customization:**
- Evaluated 6 professional themes across multiple contexts
- Applied typography-focused customization (font sizes: 15-18pt)
- Demonstrated publication-ready styling for academic and business audiences

**Patchwork Composition:**
- Created 2-panel, 3-panel, 4-panel, and 5-panel layouts
- Implemented auto-tagging for academic referencing
- Achieved unified theme application across composite figures
- Utilized inset elements for space-efficient contextualization

**Statistical Analysis:**
- Conducted `r 3 + 1` hypothesis tests (3 t-tests, 1 ANOVA)
- Calculated `r length(cors)` correlation coefficients
- Identified `r n_significant` statistically significant demographic effects
- Generated `r nrow(summary_stats) * ncol(summary_stats)` summary statistics dynamically
:::

## Evidence-Based Recommendations

```{r}
# Generate data-driven recommendations
high_performers_pct <- sum(exam_ranked$Percentile >= 90) / n_students * 100
low_performers_pct <- sum(exam_ranked$Percentile <= 10) / n_students * 100
```

::: {.callout-warning}
## Actionable Insights for Educators

**1. Intervention Targeting**
- **`r round(low_performers_pct, 1)`%** of students (`r sum(exam_ranked$Percentile <= 10)` students) require intensive support (bottom 10%)

- **Focus:** General academic skills rather than subject-specific tutoring, given high inter-subject correlation (r = `r round(mean(cors), 3)`)

**2. Excellence Programs**
- **`r round(high_performers_pct, 1)`%** of students (`r sum(exam_ranked$Percentile >= 90)` students) demonstrate exceptional ability

- **Pattern:** `r n_balanced` of top 10 show balanced excellence across subjects

- **Strategy:** `r ifelse(n_balanced >= 7, "Develop comprehensive enrichment programs", "Offer both balanced and specialized tracks")`

**3. Class-Level Equity**
- **Current gap:** `r round(class_gap, 1)` points between best (`r best_class`) and worst (`r worst_class`) performing classes

- **Target:** Reduce gap to < 15 points through `r ifelse(class_gap > 30, "immediate intervention", ifelse(class_gap > 15, "pedagogical review", "minor adjustments"))`

**4. Gender Considerations**
`r ifelse(n_significant > 0, paste0("- Address the ", n_significant, " subject(s) showing gender disparities: ", paste(sig_subjects, collapse = ", ")), "- Maintain current equitable environment (no significant gender gaps detected)")`

**5. Racial Equity Initiatives**
- **Current disparity:** `r round(performance_gap, 1)` points (`r ifelse(p_val < 0.05, "statistically significant", "within normal variation")`)

- **Action:** `r ifelse(p_val < 0.05, paste0("Investigate factors contributing to ", round(performance_gap / mean(race_analysis$Total_Mean) * 100, 1), "% performance gap and implement targeted support for ", worst_race, " students"), "Continue monitoring for emerging disparities")`
:::

# References

-   [Patchwork R package](https://patchwork.data-imaginist.com/)
-   [ggrepel](https://ggrepel.slowkow.com/)
-   [ggthemes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)
-   [hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/)
-   [R for Visual Analytics - Chapter 2](https://r4va.netlify.app/chap02)



