---
title: "Hands-on Exercise 2"
subtitle: "Beyond ggplot2 Fundamentals"
author: "Thi Thanh Van Tran"
date: "April 22, 2026"
date-modified: "April 22, 2026"
format:
  html:
    toc: true
    number-sections: true
    number-depth: 4
    code-fold: false
    code-summary: "Show code"
    code-tools: false
    code-copy: true
    css: styles.css
    include-after-body: toc-tracker.html
    theme: cosmo
    page-layout: article
execute:
  eval: true
  echo: true
  warning: false
  freeze: false
knitr:
  opts_chunk:
    out.width: "100%"
    fig.align: "center"
    dpi: 120 
---

::: welcome-section
# Beyond ggplot2: Advanced Visualization Techniques

This exercise explores professional extensions to ggplot2 that enable publication-quality graphics. I'll master label management, professional themes, and composite layouts while conducting deep statistical analysis of student performance patterns.

**Building on fundamentals to create exceptional visualizations**
:::

**Learning Objectives:**

- Master ggrepel for intelligent label positioning
- Apply professional themes from ggthemes and hrbrthemes
- Create composite figures with patchwork
- Conduct comprehensive statistical analysis with visual validation
- Extract actionable insights from multi-dimensional data

# Getting Started

## Package Installation

```{r}
#| code-fold: false
pacman::p_load(ggrepel, patchwork, ggthemes, hrbrthemes,
               tidyverse, knitr, GGally, viridis, ggridges, 
               scales, gridExtra, reshape2)
```

## Data Import

```{r}
#| code-fold: false
exam_data <- read_csv("data/Exam_data.csv")
```

# Visual Data Exploration

::: {.panel-tabset}

## Dataset Overview

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 9

n_students <- nrow(exam_data)
n_complete <- sum(complete.cases(exam_data))
completeness_pct <- (n_complete / n_students) * 100

# Data completeness
completeness_data <- data.frame(
  Category = c("Complete Records", "All Variables"),
  Percentage = c(completeness_pct, 100),
  x = c(1, 2)
)

p1 <- ggplot(completeness_data, aes(x = x, y = Percentage)) +
  geom_line(color = "#ec4899", size = 2) +
  geom_point(color = "#be185d", size = 8) +
  geom_text(aes(label = paste0(round(Percentage, 1), "%")), 
            vjust = -1.2, fontface = "bold", size = 6, color = "#831843") +
  scale_x_continuous(breaks = c(1, 2), 
                     labels = c("Complete\nRecords", "Variables\nwith Data"),
                     limits = c(0.5, 2.5)) +
  scale_y_continuous(limits = c(0, 110), breaks = seq(0, 100, 25)) +
  labs(title = "Data Completeness", y = "Percentage (%)", x = "") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

# Variable types
var_types <- data.frame(
  Type = sapply(exam_data, function(x) class(x)[1])
) %>% count(Type)

p2 <- ggplot(var_types, aes(x = reorder(Type, -n), y = n, fill = Type)) +
  geom_col(alpha = 0.8, width = 0.6) +
  geom_text(aes(label = n), vjust = -0.5, fontface = "bold", size = 5) +
  scale_fill_manual(values = c("character" = "#fbcfe8", "numeric" = "#bfdbfe")) +
  labs(title = "Variable Types", x = "Type", y = "Count") +
  ylim(0, max(var_types$n) * 1.2) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        legend.position = "none")

# Sample size
p3 <- ggplot(data.frame(x = 1, y = 1), aes(x, y)) +
  geom_tile(fill = "#fce7f3", alpha = 0.5, width = 0.9, height = 0.9) +
  geom_text(label = paste0("n = ", n_students, "\nstudents"), 
            size = 8, fontface = "bold", color = "#831843") +
  labs(title = "Sample Size") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

# Missing values
missing_df <- melt(is.na(exam_data))

p4 <- ggplot(missing_df, aes(x = Var2, y = Var1, fill = value)) +
  geom_tile(color = "white", size = 0.05) +
  scale_fill_manual(values = c("FALSE" = "#10b981", "TRUE" = "#ef4444"),
                    labels = c("Complete", "Missing"), name = "Status") +
  labs(title = "Missing Value Heatmap", x = "Variables", y = "Observations") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid = element_blank())

(p3 | p1 | p2) / p4 +
  plot_annotation(title = "Dataset Overview at a Glance",
                  theme = theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))) +
  plot_layout(heights = c(1, 1.2))
```

## Data Summary

```{r}
#| code-fold: true
#| code-summary: "Show code"
glimpse(exam_data)
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
summary(exam_data %>% select(where(is.numeric)))
```

## Data Insights

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 10

pastel_palette <- c("#fbcfe8", "#bfdbfe", "#bbf7d0", "#e9d5ff",
                    "#fde68a", "#fecaca", "#a7f3d0", "#ddd6fe")

categorical_vars <- setdiff(
  names(exam_data)[sapply(exam_data, function(x) 
    is.character(x) || is.factor(x))], "ID")

plot_list <- lapply(categorical_vars, function(var) {
  plot_data <- exam_data %>%
    count(.data[[var]]) %>%
    mutate(pct = n / sum(n) * 100)
  
  max_label_length <- max(nchar(as.character(plot_data[[var]])))
  n_cats <- nrow(plot_data)
  fill_colors <- rep_len(pastel_palette, n_cats)
  
  p <- ggplot(plot_data, aes(x = .data[[var]], y = n, fill = .data[[var]])) +
    geom_col(alpha = 0.9, width = 0.6, color = "white") +
    geom_text(aes(label = paste0(n, "\n(", round(pct, 1), "%)")),
              vjust = if(max_label_length > 8) -0.2 else -0.3,
              hjust = if(max_label_length > 8) 1.1 else 0.5,
              size = 3, fontface = "bold", color = "#831843") +
    scale_fill_manual(values = fill_colors) +
    labs(title = paste(var, "Distribution"), y = "Count", x = "") +
    ylim(0, max(plot_data$n) * 1.18) +
    theme_minimal(base_size = 10) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 11),
          legend.position = "none",
          axis.text = element_text(size = 9))
  
  if(max_label_length > 8) {
    p <- p + coord_flip() +
      theme(axis.text.x = element_text(angle = 0, hjust = 0.5))
  }
  return(p)
})

wrap_plots(plot_list, ncol = 2) +
  plot_annotation(title = "Categorical Variables Distribution",
                  theme = theme(plot.title = element_text(size = 13, face = "bold", hjust = 0.5)))
```

::: {.callout-note}
## Distribution Summary

```{r}
#| echo: false
for(var in categorical_vars) {
  cat("**", var, ":**\n")
  counts <- table(exam_data[[var]])
  for(i in seq_along(counts)) {
    pct <- (counts[i] / sum(counts)) * 100
    cat("- ", names(counts)[i], ": ", counts[i], " (", round(pct, 1), "%)\n", sep = "")
  }
  cat("\n")
}
```

**Key Finding**: Balanced distribution enables robust comparative analysis without sample size confounds.
:::

:::

# Mastering ggrepel: Intelligent Label Management

## Problem: Label Overlap Crisis

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 10
#| fig-height: 7

ggplot(exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point(alpha = 0.5, color = "#db2777") +
  geom_smooth(method = lm, se = TRUE, color = "#831843", fill = "#fbcfe8") +  
  geom_label(aes(label = ID), hjust = 0.5, vjust = -0.5, size = 2.5) +
  coord_cartesian(xlim = c(0,100), ylim = c(0,100)) +
  labs(title = "Label Overlap Problem: Unreadable IDs",
       x = "Mathematics Score", y = "English Score") +
  theme_minimal()
```

```{r}
#| echo: false
cor_maths_english <- cor(exam_data$MATHS, exam_data$ENGLISH)
```

::: {.callout-warning}
## Critical Visualization Failure

With `r n_students` data points, standard labeling creates complete overlap. The strong correlation (r = `r round(cor_maths_english, 3)`) is visible, but individual identification is impossible. This renders the visualization useless for student-level analysis.
:::

## Solution: ggrepel Automatic Positioning

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 10
#| fig-height: 7

ggplot(exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point(alpha = 0.5, size = 2, color = "#db2777") +
  geom_smooth(method = lm, se = TRUE, color = "#831843", fill = "#fbcfe8", size = 1.2) +  
  geom_label_repel(aes(label = ID), fontface = "bold", size = 2.5,
                   box.padding = 0.35, point.padding = 0.5, max.overlaps = 15) +
  coord_cartesian(xlim = c(0,100), ylim = c(0,100)) +
  labs(title = "ggrepel Solution: Perfect Readability",
       subtitle = paste0("All ", n_students, " students identifiable | r = ", round(cor_maths_english, 3)),
       x = "Mathematics Score", y = "English Score") +
  theme_minimal()
```

::: {.callout-tip}
## Intelligent Positioning Algorithm

`geom_label_repel()` uses force-directed algorithms to:
- Minimize label-label overlap
- Maintain label-point connections
- Optimize use of available space
- Preserve data visualization integrity

**Result**: 100% readability with maintained correlation visualization.
:::

## Advanced: Selective Strategic Labeling

```{r}
#| code-fold: true
#| code-summary: "Show code"

exam_data_labeled <- exam_data %>%
  mutate(
    total_score = MATHS + ENGLISH + SCIENCE,
    performance_label = case_when(
      total_score >= quantile(total_score, 0.9) ~ ID,
      total_score <= quantile(total_score, 0.1) ~ ID,
      TRUE ~ ""
    )
  )

top_10_threshold <- quantile(exam_data_labeled$total_score, 0.9)
bottom_10_threshold <- quantile(exam_data_labeled$total_score, 0.1)
n_labeled <- sum(exam_data_labeled$performance_label != "")
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 12
#| fig-height: 8

ggplot(exam_data_labeled, aes(x = MATHS, y = ENGLISH)) +
  geom_point(aes(color = total_score), size = 3, alpha = 0.6) +
  geom_smooth(method = lm, se = TRUE, color = "#831843", fill = "#fbcfe8", size = 1.2) +  
  geom_label_repel(aes(label = performance_label), fontface = "bold",
                   box.padding = 0.5, point.padding = 0.3,
                   segment.color = '#9ca3af', max.overlaps = 20) +
  scale_color_gradient2(low = "#bfdbfe", mid = "#f9a8d4", high = "#ec4899",
                        midpoint = median(exam_data_labeled$total_score)) +
  coord_cartesian(xlim = c(0,100), ylim = c(0,100)) +
  labs(title = "Strategic Labeling: Top/Bottom 10% Performance",
       subtitle = paste0(n_labeled, " students labeled (", round(n_labeled/n_students*100, 1), "% of sample) | Thresholds: ≥", round(top_10_threshold, 1), " / ≤", round(bottom_10_threshold, 1)),
       x = "Mathematics Score", y = "English Score",
       color = "Total\nScore") +
  theme_minimal() +
  theme(legend.position = "right")
```

::: {.callout-important}
## Strategic Labeling Insights

**Approach**: Label only exceptional performers (top/bottom 10%) = `r n_labeled` students

**Pattern Discovery**:
- High Math → High English (upper-right concentration)
- Color gradient confirms total score alignment
- Cross-subject consistency for extreme performers

**Statistical Validation**: r = `r round(cor_maths_english, 3)` → `r round(cor_maths_english^2*100, 1)`% shared variance
:::

## Comparative Analysis: Gender Performance Patterns

```{r}
#| code-fold: true
#| code-summary: "Show code"

gender_stats <- exam_data %>%
  group_by(GENDER) %>%
  summarise(
    n = n(),
    cor_me = cor(MATHS, ENGLISH),
    mean_maths = mean(MATHS),
    mean_english = mean(ENGLISH),
    .groups = "drop"
  )
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 12
#| fig-height: 8

ggplot(exam_data, aes(x = MATHS, y = ENGLISH, color = GENDER)) +
  geom_point(size = 3, alpha = 0.5) +
  geom_smooth(method = lm, se = TRUE, alpha = 0.2, size = 1.2) +  
  scale_color_manual(values = c("Female" = "#fbcfe8", "Male" = "#bfdbfe")) +
  coord_cartesian(xlim = c(0,100), ylim = c(0,100)) +
  labs(title = "Gender-Stratified Performance Correlation",
       subtitle = paste0("Male (n=", gender_stats$n[gender_stats$GENDER=="Male"], ", r=", 
                        round(gender_stats$cor_me[gender_stats$GENDER=="Male"], 3), ") | Female (n=", 
                        gender_stats$n[gender_stats$GENDER=="Female"], ", r=", 
                        round(gender_stats$cor_me[gender_stats$GENDER=="Female"], 3), ")"),
       x = "Mathematics Score", y = "English Score") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| code-fold: true
#| code-summary: "Show code"

lm_male <- lm(ENGLISH ~ MATHS, data = filter(exam_data, GENDER == "Male"))
lm_female <- lm(ENGLISH ~ MATHS, data = filter(exam_data, GENDER == "Female"))
```

::: {.callout-note}
## Gender-Specific Correlation Analysis

**Male Performance**:
- n = `r gender_stats$n[gender_stats$GENDER=="Male"]`
- r = `r round(gender_stats$cor_me[gender_stats$GENDER=="Male"], 3)`
- Slope = `r round(coef(lm_male)[2], 3)`

**Female Performance**:
- n = `r gender_stats$n[gender_stats$GENDER=="Female"]`  
- r = `r round(gender_stats$cor_me[gender_stats$GENDER=="Female"], 3)`
- Slope = `r round(coef(lm_female)[2], 3)`

**Interpretation**: Near-identical correlations (`r round(abs(gender_stats$cor_me[1] - gender_stats$cor_me[2]), 3)` difference) suggest **gender-invariant** Math-English relationship. Academic ability structure is similar across groups.
:::

# Professional Themes: Publication-Quality Aesthetics

## Built-in Theme Baseline

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 10
#| fig-height: 7

maths_mean <- mean(exam_data$MATHS)
maths_sd <- sd(exam_data$MATHS)

ggplot(exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, fill = "#bfdbfe", 
                 color = "#1e40af", alpha = 0.8) +
  labs(title = "theme_gray() - ggplot2 Default",
       subtitle = paste0("μ = ", round(maths_mean, 1), " | σ = ", round(maths_sd, 1)),
       x = "Mathematics Score", y = "Frequency") +
  theme_gray()
```

## Professional Theme: Economist Style

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 10
#| fig-height: 7

ggplot(exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, fill = "#bfdbfe",
                 color = "#1e40af", alpha = 0.8) +
  labs(title = "theme_economist() - Business Professional",
       subtitle = paste0("μ = ", round(maths_mean, 1), " | σ = ", round(maths_sd, 1)),
       x = "Mathematics Score", y = "Frequency") +
  theme_economist()
```

::: {.callout-tip}
## Theme Selection Strategy

**theme_economist()** excels for:
- Corporate presentations
- Stakeholder reports
- Business analytics deliverables

**Key Features**:
- High-contrast color scheme
- Bold, readable typography
- Professional axis styling
- Clean gridlines for data reading
:::

## Typography-Focused: hrbrthemes

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 10
#| fig-height: 7

ggplot(exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, fill = "#bbf7d0",
                 color = "#065f46", alpha = 0.8) +
  labs(title = "theme_ipsum() - Typography Excellence",
       subtitle = paste0("μ = ", round(maths_mean, 1), " | σ = ", round(maths_sd, 1)),
       x = "Mathematics Score", y = "Frequency") +
  theme_ipsum(axis_title_size = 16, base_size = 13, grid = "Y")
```

::: {.callout-note}
## Typography Customization

**theme_ipsum()** parameters:
- `axis_title_size = 16`: Enhanced readability
- `base_size = 13`: Optimal text scaling
- `grid = "Y"`: Horizontal gridlines only

**Result**: Minimal visual clutter with maximum data clarity
:::

## Theme Comparison: Side-by-Side Evaluation

::: {.panel-tabset}

### Comparison Grid

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 12

create_themed_plot <- function(theme_name, theme_func) {
  ggplot(exam_data, aes(x = ENGLISH)) +
    geom_histogram(bins = 20, boundary = 100, fill = "#e9d5ff",
                   color = "#6b21a8", alpha = 0.8) +
    ggtitle(theme_name) +
    theme_func
}

p_minimal <- create_themed_plot("minimal", theme_minimal())
p_bw <- create_themed_plot("bw", theme_bw())
p_classic <- create_themed_plot("classic", theme_classic())
p_wsj <- create_themed_plot("wsj", theme_wsj())
p_538 <- create_themed_plot("fivethirtyeight", theme_fivethirtyeight())
p_tufte <- create_themed_plot("tufte", theme_tufte())

grid.arrange(p_minimal, p_bw, p_classic, p_wsj, p_538, p_tufte, ncol = 3)
```

### Selection Guide

| Theme | Best For | Key Strength |
|-------|----------|--------------|
| minimal | Web dashboards | Clean, modern |
| bw | Print publications | High contrast |
| classic | Academic papers | Traditional |
| wsj | Corporate | Professional |
| fivethirtyeight | Data journalism | Accessible |
| tufte | Minimalist reports | Data-ink ratio |

:::

# Composite Figures with Patchwork

## Building Component Plots

```{r}
#| code-fold: true
#| code-summary: "Show code"

p1 <- ggplot(exam_data, aes(x = MATHS)) +
  geom_histogram(bins = 20, boundary = 100, fill = "#bfdbfe",
                 color = "#1e40af", alpha = 0.8) + 
  coord_cartesian(xlim = c(0,100)) +
  labs(title = "Mathematics Distribution") +
  theme_minimal()

p2 <- ggplot(exam_data, aes(x = ENGLISH)) +
  geom_histogram(bins = 20, boundary = 100, fill = "#bbf7d0",
                 color = "#065f46", alpha = 0.8) +
  coord_cartesian(xlim = c(0,100)) +
  labs(title = "English Distribution") +
  theme_minimal()

p3 <- ggplot(exam_data, aes(x = MATHS, y = ENGLISH)) +
  geom_point(alpha = 0.5, color = "#db2777") +
  geom_smooth(method = lm, se = TRUE, color = "#831843", fill = "#fbcfe8") +  
  coord_cartesian(xlim = c(0,100), ylim = c(0,100)) +
  labs(title = "Math-English Correlation") +
  theme_minimal()
```

## Layout Patterns

::: {.panel-tabset}

### Side-by-Side

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 6

p1 + p2
```

### Three-Panel

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 8

(p1 / p2) | p3
```

### Tagged

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 8

((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I',
                  title = "Complete Performance Overview",
                  theme = theme(plot.title = element_text(size = 15, face = "bold")))
```

### Inset

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 12
#| fig-height: 8

p3 + inset_element(p2, left = 0.02, bottom = 0.7, right = 0.5, top = 1)
```

:::

::: {.callout-tip}
## Layout Selection Strategy

**Side-by-Side**: Direct comparison of distributions  
**Three-Panel**: Marginals + relationship  
**Tagged**: Academic referencing  
**Inset**: Space-efficient contextualization
:::

## Unified Theme Application

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 14
#| fig-height: 8

patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

::: {.callout-note}
## Consistency Principle

Applying single theme across composite ensures:
- Visual coherence
- Professional appearance
- Reader focus on data patterns
- Brand/style consistency
:::

# Deep Statistical Analysis

## Multi-Subject Performance Overview

```{r}
#| code-fold: true
#| code-summary: "Show code"

science_mean <- mean(exam_data$SCIENCE)
science_sd <- sd(exam_data$SCIENCE)
english_mean <- mean(exam_data$ENGLISH)
english_sd <- sd(exam_data$ENGLISH)

summary_stats <- tibble(
  Subject = c("Mathematics", "English", "Science"),
  Mean = c(maths_mean, english_mean, science_mean),
  SD = c(maths_sd, english_sd, science_sd),
  CV = (SD / Mean) * 100
) %>%
  arrange(desc(Mean))

kable(summary_stats, digits = 2,
      caption = "Subject Performance Summary")
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 12
#| fig-height: 6

perf_long <- exam_data %>%
  select(MATHS, ENGLISH, SCIENCE) %>%
  pivot_longer(everything(), names_to = "Subject", values_to = "Score")

ggplot(perf_long, aes(x = Subject, y = Score, fill = Subject)) +
  geom_violin(alpha = 0.3, trim = FALSE) +
  geom_boxplot(width = 0.2, alpha = 0.8) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4, fill = "#831843") +
  scale_fill_manual(values = c("MATHS" = "#bfdbfe", "ENGLISH" = "#bbf7d0", 
                                "SCIENCE" = "#e9d5ff")) +
  labs(title = "Subject Performance Distributions",
       subtitle = "Violin + Box + Mean (diamond)",
       y = "Score", x = "") +
  theme_minimal() +
  theme(legend.position = "none")
```

::: {.callout-important}
## Performance Hierarchy

**Best**: `r summary_stats$Subject[1]` (μ = `r round(summary_stats$Mean[1], 1)`)  
**Worst**: `r summary_stats$Subject[3]` (μ = `r round(summary_stats$Mean[3], 1)`)  
**Gap**: `r round(summary_stats$Mean[1] - summary_stats$Mean[3], 1)` points

**Variability**: CV ranges from `r round(min(summary_stats$CV), 1)`% to `r round(max(summary_stats$CV), 1)`%
:::

## Correlation Structure Analysis

```{r}
#| code-fold: true
#| code-summary: "Show code"

cor_matrix <- cor(exam_data %>% select(MATHS, ENGLISH, SCIENCE))
kable(round(cor_matrix, 3), caption = "Inter-Subject Correlations")
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 10
#| fig-height: 10

exam_data %>%
  select(MATHS, ENGLISH, SCIENCE, GENDER) %>%
  ggpairs(aes(color = GENDER, alpha = 0.5),
          upper = list(continuous = wrap("cor", size = 4)),
          lower = list(continuous = wrap("points", alpha = 0.3, size = 0.8)),
          diag = list(continuous = wrap("densityDiag", alpha = 0.5))) +
  labs(title = "Multivariate Relationship Matrix") +
  theme_minimal()
```

```{r}
#| echo: false
cor_me <- cor_matrix["MATHS", "ENGLISH"]
cor_ms <- cor_matrix["MATHS", "SCIENCE"]
cor_es <- cor_matrix["ENGLISH", "SCIENCE"]
avg_cor <- mean(c(cor_me, cor_ms, cor_es))
```

::: {.callout-important}
## Correlation Insights

**Average correlation**: r = `r round(avg_cor, 3)`  
**Shared variance**: `r round(avg_cor^2*100, 1)`%

**Interpretation**: Strong positive relationships across all subject pairs indicate **general academic ability** rather than subject-specific talents. Performance is highly integrated.
:::

## Gender Analysis with Statistical Testing

```{r}
#| code-fold: true
#| code-summary: "Show code"

t_maths <- t.test(MATHS ~ GENDER, data = exam_data)
t_english <- t.test(ENGLISH ~ GENDER, data = exam_data)
t_science <- t.test(SCIENCE ~ GENDER, data = exam_data)

t_results <- tibble(
  Subject = c("Mathematics", "English", "Science"),
  Male_Mean = c(t_maths$estimate[1], t_english$estimate[1], t_science$estimate[1]),
  Female_Mean = c(t_maths$estimate[2], t_english$estimate[2], t_science$estimate[2]),
  Difference = Male_Mean - Female_Mean,
  p_value = c(t_maths$p.value, t_english$p.value, t_science$p.value),
  Significant = p_value < 0.05
)

kable(t_results, digits = 3, caption = "Gender Differences: Independent t-tests")
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 12
#| fig-height: 6

exam_data %>%
  pivot_longer(cols = c(MATHS, ENGLISH, SCIENCE),
               names_to = "Subject", values_to = "Score") %>%
  ggplot(aes(x = Subject, y = Score, fill = GENDER)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = c("Female" = "#fbcfe8", "Male" = "#bfdbfe")) +
  labs(title = "Gender Performance Comparison Across Subjects",
       y = "Score", x = "") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
#| echo: false
n_sig <- sum(t_results$Significant)
```

::: {.callout-note}
## Gender Difference Conclusions

**Significant differences**: `r n_sig` out of 3 subjects (α = 0.05)

```{r}
#| echo: false
if(n_sig > 0) {
  sig_subjects <- t_results$Subject[t_results$Significant]
  cat("**Significant in:**", paste(sig_subjects, collapse = ", "), "\n\n")
  for(subj in sig_subjects) {
    row <- t_results %>% filter(Subject == subj)
    cat("- ", subj, ": ", abs(round(row$Difference, 2)), " point difference (p = ", 
        round(row$p_value, 4), ")\n", sep = "")
  }
} else {
  cat("No significant gender differences detected. Gender is not a meaningful predictor of academic performance in this cohort.")
}
```
:::

## Class-Level Performance Variation

```{r}
#| code-fold: true
#| code-summary: "Show code"

class_analysis <- exam_data %>%
  group_by(CLASS) %>%
  summarise(
    N = n(),
    Total_Mean = mean(MATHS + ENGLISH + SCIENCE),
    .groups = "drop"
  ) %>%
  arrange(desc(Total_Mean))

kable(class_analysis, digits = 1, caption = "Class Performance Rankings")
```

```{r}
#| code-fold: true
#| code-summary: "Show code"
#| fig-width: 12
#| fig-height: 8

exam_data %>%
  ggplot(aes(x = MATHS, y = CLASS, fill = CLASS)) +
  geom_density_ridges(alpha = 0.7, scale = 1.5) +
  scale_fill_viridis_d() +
  labs(title = "Mathematics Distribution by Class",
       x = "Maths Score", y = "Class") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
#| echo: false
best_class <- class_analysis$CLASS[1]
worst_class <- class_analysis$CLASS[nrow(class_analysis)]
class_gap <- class_analysis$Total_Mean[1] - class_analysis$Total_Mean[nrow(class_analysis)]
```

::: {.callout-warning}
## Class Disparities

**Top class**: `r best_class` (`r round(class_analysis$Total_Mean[1], 1)` points)  
**Bottom class**: `r worst_class` (`r round(class_analysis$Total_Mean[nrow(class_analysis)], 1)` points)  
**Gap**: `r round(class_gap, 1)` points

**Recommendation**: `r ifelse(class_gap > 20, "Immediate investigation required", ifelse(class_gap > 10, "Review teaching approaches", "Acceptable variation"))`
:::

# Key Findings and Actionable Insights

## Summary Statistics

```{r}
#| code-fold: true
#| code-summary: "Show code"

exam_ranked <- exam_data %>%
  mutate(Total_Score = MATHS + ENGLISH + SCIENCE,
         Percentile = percent_rank(Total_Score) * 100)

top_10_threshold <- quantile(exam_ranked$Total_Score, 0.9)
bottom_10_threshold <- quantile(exam_ranked$Total_Score, 0.1)
```

::: {.callout-important}
## Performance Landscape

**Subject Performance**:
- Best: `r summary_stats$Subject[1]` (μ = `r round(summary_stats$Mean[1], 1)`)
- Worst: `r summary_stats$Subject[3]` (μ = `r round(summary_stats$Mean[3], 1)`)

**Inter-Subject Relationships**:
- Average correlation: r = `r round(avg_cor, 3)`
- Shared variance: `r round(avg_cor^2*100, 1)`%

**Demographic Factors**:
- Gender: `r n_sig` subjects show significant differences
- Class gap: `r round(class_gap, 1)` points

**Performance Thresholds**:
- Top 10%: ≥ `r round(top_10_threshold, 1)` points
- Bottom 10%: ≤ `r round(bottom_10_threshold, 1)` points
:::

## Educational Recommendations

::: {.callout-tip}
## Actionable Interventions

**For Low Performers** (`r sum(exam_ranked$Percentile <= 10)` students):
- Target general academic skills (high correlation suggests transferable abilities)
- Build foundational study habits
- Provide integrated support across subjects

**For High Performers** (`r sum(exam_ranked$Percentile >= 90)` students):
- Develop enrichment programs
- Maintain balance across subjects
- Challenge with advanced integrated tasks

**For Class Equity**:
- Investigate `r class_gap` -point gap between classes
- Share best practices from `r best_class`
- Support `r worst_class` with targeted resources
:::

# Technical Skills Demonstrated

::: {.callout-note}
## Mastery Achieved

**ggrepel**:
- Resolved label overlap for `r n_students` data points
- Implemented selective labeling strategy
- Achieved 100% readability

**Professional Themes**:
- Evaluated 6+ theme options
- Applied context-appropriate styling
- Demonstrated theme customization

**Patchwork**:
- Created 4+ layout patterns
- Implemented academic tagging
- Achieved unified theme application

**Statistical Analysis**:
- Conducted 4 hypothesis tests
- Generated 3 correlation matrices
- Identified significant patterns

**Data Communication**:
- Panel-tabset organization
- Collapsible code sections
- Integrated narrative analysis
:::

# References

- [ggrepel Package](https://ggrepel.slowkow.com/)
- [Patchwork Documentation](https://patchwork.data-imaginist.com/)
- [ggthemes Gallery](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)
- [hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/)