---
title: "In-class Exercise 2"
subtitle: "Analysing Superstore Dataset"
author: "Thi Thanh Van Tran"
date: "April 18, 2026"
date-modified: "April 18, 2026"
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    number-sections: true
    number-depth: 3
    code-fold: true
    code-summary: "Show Code"
    code-tools: true
    code-copy: true
    theme: flatly
    highlight-style: github
    fig-align: center
execute:
  eval: true
  echo: true
  warning: false
  message: false
  freeze: false
  fig-width: 10
  fig-height: 5
---

```{css, echo=FALSE}
img, .plotly, iframe { max-width: 100% !important; height: auto !important; }
.figure { margin-bottom: 1rem; }
.callout { margin-top: 0.75rem; }
```

# Tableau Dashboard

```{=html}
<iframe
  src="https://public.tableau.com/views/in-class3b/Sheet2?:showVizHome=no&:embed=true&:toolbar=yes"
  style="width:100%; height:620px; border:none;"
  allowfullscreen>
</iframe>
```

---

# Getting Started

## Installing and Loading the Required Libraries

::: callout-note
Ensure the [pacman](https://cran.r-project.org/web/packages/pacman/) package is installed before running this document.
:::

```{r}
#| code-fold: false
pacman::p_load(
  readxl, tidyverse, lubridate, scales,
  ggplot2, patchwork, ggridges, ggrepel, ggthemes,
  viridis, RColorBrewer, treemapify,
  plotly, ggiraph, DT,
  gt, kableExtra
)
```

## Importing Data

```{r}
#| code-fold: false
superstore <- read_excel("data/Superstore.xls") %>%
  mutate(
    Year      = year(`Order Date`),
    Month     = month(`Order Date`, label = TRUE, abbr = TRUE),
    YearMonth = floor_date(`Order Date`, "month"),
    Quarter   = paste0("Q", quarter(`Order Date`)),
    ShipDays  = as.numeric(difftime(`Ship Date`, `Order Date`, units = "days"))
  )
```

---

# Data Overview

::: {.panel-tabset}

## Variable Structure

```{r}
#| fig-height: 5
original_cols <- superstore

var_types <- sapply(original_cols, function(x) class(x)[1]) %>%
  enframe(name = "Variable", value = "Type") %>%
  mutate(Type_Clean = case_when(
    Type %in% c("numeric","integer","double") ~ "Numeric",
    Type %in% c("character","factor")         ~ "Character",
    str_detect(Type, "POSIX|Date")            ~ "Date / Time",
    TRUE                                      ~ "Other"
  )) %>%
  arrange(Type_Clean, Variable)

type_counts <- var_types %>% count(Type_Clean) %>% mutate(pct = n / sum(n))

p_types <- ggplot(type_counts,
                  aes(x = reorder(Type_Clean, n), y = n, fill = Type_Clean)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = paste0(n, " vars")), hjust = -0.15, fontface = "bold", size = 4.2) +
  scale_fill_viridis_d(option = "D", begin = 0.2, end = 0.75) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.35))) +
  coord_flip() +
  labs(title = "Variable Types",
       subtitle = sprintf("Total: %d variables", ncol(original_cols)),
       x = NULL, y = "Count") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))

p_varlist <- ggplot(var_types,
                    aes(y = reorder(Variable, desc(Variable)), x = 1, fill = Type_Clean)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = Variable), x = 1, size = 2.8, color = "white", fontface = "bold") +
  scale_fill_viridis_d(option = "D", begin = 0.2, end = 0.75, name = "Type") +
  scale_x_continuous(expand = c(0, 0)) +
  labs(title = "All Variables", x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(axis.text = element_blank(), axis.ticks = element_blank(),
        panel.grid = element_blank(),
        plot.title = element_text(face = "bold"), legend.position = "right")

p_types + p_varlist + plot_layout(widths = c(1, 1.6))
```

> The dataset has **`r ncol(original_cols)`** variables — `r sum(var_types$Type_Clean == "Numeric")` numeric, `r sum(var_types$Type_Clean == "Character")` character, and `r sum(var_types$Type_Clean == "Date / Time")` date/time. Five derived columns added for analysis.

## Raw Data Preview

```{r}
DT::datatable(
  head(superstore %>% select(-Year,-Month,-YearMonth,-Quarter,-ShipDays), 200),
  options = list(pageLength = 8, scrollX = TRUE, dom = "ltipr",
                 columnDefs = list(list(className = "dt-center", targets = "_all"))),
  filter  = "top",
  caption = htmltools::tags$caption(
    style = "caption-side:top; text-align:left; font-weight:bold;",
    "Superstore Dataset — first 200 rows"
  )
)
```

## Key Distributions

::: {.panel-tabset}

### Sales & Profit

```{r}
#| fig-height: 4
prof_mean   <- mean(superstore$Profit)
prof_median <- median(superstore$Profit)

p_sales <- ggplot(superstore, aes(x = Sales)) +
  geom_histogram(aes(y = after_stat(density)), bins = 50,
                 fill = "#3498db", alpha = 0.75, color = "white") +
  geom_density(color = "#1a252f", linewidth = 1.2) +
  scale_x_log10(labels = dollar_format()) +
  labs(title = "Sales Distribution (log scale)",
       subtitle = "Right-skewed — bulk of orders: $10–$1,000",
       x = "Sales (log₁₀)", y = "Density") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_profit <- ggplot(superstore, aes(x = Profit)) +
  geom_histogram(aes(y = after_stat(density)), bins = 70,
                 fill = "#2ecc71", alpha = 0.65, color = "white") +
  geom_density(color = "#1e8449", linewidth = 1.2) +
  geom_vline(xintercept = 0,           linetype = "dashed", color = "#e74c3c", linewidth = 1.0) +
  geom_vline(xintercept = prof_mean,   linetype = "dotted", color = "#2980b9", linewidth = 0.9) +
  geom_vline(xintercept = prof_median, linetype = "dotted", color = "#8e44ad", linewidth = 0.9) +
  scale_x_continuous(labels = dollar_format()) +
  labs(title = "Profit Distribution",
       subtitle = "Red = break-even · Blue = mean · Purple = median",
       x = "Profit / Transaction", y = "Density") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_sales | p_profit
```

### Discount & Quantity

```{r}
#| fig-height: 4
p_disc <- ggplot(superstore, aes(x = Discount)) +
  geom_histogram(aes(y = after_stat(density)), bins = 30,
                 fill = "#e67e22", alpha = 0.75, color = "white") +
  geom_density(color = "#a04000", linewidth = 1.2) +
  scale_x_continuous(labels = percent_format()) +
  labs(title = "Discount Distribution",
       subtitle = "Spikes at 0%, 20%, 30%… confirm structured discount tiers",
       x = "Discount Rate", y = "Density") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_qty <- ggplot(superstore, aes(x = Quantity)) +
  geom_bar(fill = "#9b59b6", alpha = 0.85, color = "white") +
  scale_y_continuous(labels = comma_format()) +
  labs(title = "Quantity Distribution",
       subtitle = "Most orders are 2–4 units",
       x = "Units per Order", y = "Count") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_disc | p_qty
```

### Profit by Category

```{r}
#| fig-height: 4
ggplot(superstore, aes(x = Profit, y = Category, fill = Category)) +
  geom_density_ridges(alpha = 0.75, scale = 1.1,
                      quantile_lines = TRUE, quantiles = c(0.25, 0.5, 0.75)) +
  geom_vline(xintercept = 0, color = "#e74c3c", linetype = "dashed", linewidth = 0.9) +
  scale_fill_brewer(palette = "Set2") +
  scale_x_continuous(labels = dollar_format(), limits = c(-300, 600)) +
  labs(title = "Profit by Category — Ridge Plot",
       subtitle = "Lines = Q1/Median/Q3 · Red = break-even · Furniture has the most losses",
       x = "Profit / Transaction", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")
```

:::

**Key observations:** Sales is right-skewed ($10–$1,000 dominates). Profit centres near zero with many losses. Discounts cluster at structured tiers. Furniture shows the most loss-making transactions.

:::

---

# Profitability Analysis

## By Category

::: {.panel-tabset}

### Distribution & Totals

```{r}
#| fig-height: 5
category_metrics <- superstore %>%
  group_by(Category) %>%
  summarise(
    Total_Profit   = sum(Profit),
    Total_Sales    = sum(Sales),
    Avg_Profit     = mean(Profit),
    Profit_Margin  = Total_Profit / Total_Sales * 100,
    Count          = n(),
    Pct_Profitable = mean(Profit > 0) * 100,
    .groups        = "drop"
  ) %>%
  arrange(desc(Total_Profit))

p_cat_box <- ggplot(superstore,
                    aes(x = reorder(Category, Profit, FUN = median),
                        y = Profit, fill = Category)) +
  geom_boxplot(alpha = 0.75, outlier.alpha = 0.2, outlier.size = 0.8) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4,
               fill = "white", color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#e74c3c", linewidth = 0.8) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = dollar_format()) +
  coord_flip() +
  labs(title = "Profit Distribution by Category",
       subtitle = "◆ = mean · Box = IQR · Red dashed = break-even",
       x = NULL, y = "Profit / Transaction") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")

p_cat_total <- ggplot(category_metrics,
                      aes(x = reorder(Category, Total_Profit),
                          y = Total_Profit, fill = Category)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = dollar(Total_Profit, accuracy = 1)),
            hjust = -0.1, size = 3.8, fontface = "bold") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = dollar_format(),
                     expand = expansion(mult = c(0, 0.3))) +
  coord_flip() +
  labs(title = "Total Profit by Category", x = NULL, y = "Total Profit") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_cat_box | p_cat_total
```

### Margin vs % Profitable

```{r}
#| fig-height: 4
ggplot(category_metrics,
       aes(x = Profit_Margin, y = Pct_Profitable,
           color = Category, size = Total_Sales, label = Category)) +
  geom_point(alpha = 0.85) +
  geom_text_repel(size = 4.5, fontface = "bold", show.legend = FALSE, box.padding = 0.6) +
  scale_color_brewer(palette = "Set2") +
  scale_size_continuous(labels = dollar_format(), name = "Total Sales", range = c(10, 20)) +
  scale_x_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%"), limits = c(50, 100)) +
  labs(title = "Profit Margin vs % Profitable Transactions",
       subtitle = "Bubble size ∝ Total Sales · Upper right = best performers",
       x = "Overall Profit Margin (%)", y = "% Transactions with Profit > 0") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")
```

:::

::: callout-tip
**Findings:**
```{r}
#| echo: false
best_cat    <- category_metrics$Category[1]
best_total  <- dollar(category_metrics$Total_Profit[1])
best_margin <- paste0(round(category_metrics$Profit_Margin[1], 1), "%")
```
- **`r best_cat`** leads with **`r best_total`** total profit and a **`r best_margin`** margin.
- **Furniture** has the lowest share of profitable transactions — systemic pricing problem.
:::

## By Sub-Category

::: {.panel-tabset}

### Pareto Chart — Who drives 80% of Profit?

```{r}
#| fig-height: 5

# Build Pareto for profitable sub-categories only (ranked by contribution)
subcat_pareto <- superstore %>%
  group_by(`Sub-Category`, Category) %>%
  summarise(Total_Profit = sum(Profit), .groups = "drop") %>%
  arrange(desc(Total_Profit)) %>%
  mutate(
    Profitable    = Total_Profit > 0,
    cumsum_profit = cumsum(pmax(Total_Profit, 0)),
    total_pos     = sum(pmax(Total_Profit, 0)),
    cum_pct       = cumsum_profit / total_pos * 100,
    rank          = row_number(),
    bar_color     = ifelse(Total_Profit > 0, "#2980b9", "#e74c3c")
  )

# Threshold where cumulative hits 80%
threshold_rank <- min(subcat_pareto$rank[subcat_pareto$cum_pct >= 80], na.rm = TRUE)

ggplot(subcat_pareto, aes(x = reorder(`Sub-Category`, -rank))) +
  # Bars
  geom_col(aes(y = Total_Profit / 1000, fill = Profitable),
           width = 0.75, alpha = 0.9) +
  # Cumulative % line (right axis, scaled to bar height)
  geom_line(aes(y = cum_pct / 100 * max(Total_Profit / 1000, na.rm = TRUE),
                group = 1),
            color = "#e67e22", linewidth = 1.6, na.rm = TRUE) +
  geom_point(aes(y = cum_pct / 100 * max(Total_Profit / 1000, na.rm = TRUE)),
             color = "#e67e22", size = 2.5, na.rm = TRUE) +
  # 80% reference line
  geom_hline(yintercept = 0.8 * max(subcat_pareto$Total_Profit / 1000, na.rm = TRUE),
             linetype = "dashed", color = "#e67e22", linewidth = 0.8, alpha = 0.7) +
  annotate("text",
           x    = nrow(subcat_pareto) - 0.3,
           y    = 0.82 * max(subcat_pareto$Total_Profit / 1000, na.rm = TRUE),
           label = "80% threshold",
           color = "#e67e22", size = 3.2, fontface = "bold", hjust = 1) +
  # Colour scale for bars
  scale_fill_manual(values = c("TRUE" = "#2980b9", "FALSE" = "#e74c3c"),
                    labels  = c("FALSE" = "Loss", "TRUE" = "Profit"), name = NULL) +
  # Right Y axis label (cumulative %)
  scale_y_continuous(
    name     = "Total Profit (K$)",
    labels   = function(x) paste0("$", round(x, 0), "K"),
    sec.axis = sec_axis(
      ~ . / max(subcat_pareto$Total_Profit / 1000, na.rm = TRUE) * 100,
      name   = "Cumulative % of Total Profit",
      labels = function(x) paste0(round(x, 0), "%")
    )
  ) +
  labs(title = "Pareto Chart: Sub-Category Profit Contribution",
       subtitle = "Blue bars = profit · Red bars = loss · Orange line = cumulative %",
       x = "Sub-Category",
       caption = "Source: Superstore Dataset") +
  theme_minimal(base_size = 11) +
  theme(
    plot.title   = element_text(face = "bold", size = 13),
    axis.text.x  = element_text(angle = 35, hjust = 1, size = 9),
    axis.title.y.right = element_text(color = "#e67e22"),
    legend.position = "bottom"
  )
```

### Lollipop — Total Profit Ranking

```{r}
#| fig-height: 6
subcategory_metrics <- superstore %>%
  group_by(`Sub-Category`, Category) %>%
  summarise(Total_Profit = sum(Profit), Total_Sales = sum(Sales),
            Count = n(), .groups = "drop") %>%
  mutate(Profitable = Total_Profit > 0) %>%
  arrange(Total_Profit)

ggplot(subcategory_metrics,
       aes(x = reorder(`Sub-Category`, Total_Profit), y = Total_Profit,
           color = Profitable)) +
  geom_segment(aes(xend = reorder(`Sub-Category`, Total_Profit), yend = 0),
               linewidth = 1.2, alpha = 0.7) +
  geom_point(size = 4.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.6) +
  geom_text(aes(label = dollar(Total_Profit, accuracy = 1),
                hjust = ifelse(Total_Profit >= 0, -0.15, 1.15)),
            size = 2.8, fontface = "bold") +
  scale_color_manual(values = c("TRUE" = "#27ae60", "FALSE" = "#e74c3c"),
                     labels  = c("FALSE" = "Loss", "TRUE" = "Profit"), name = NULL) +
  scale_y_continuous(labels = dollar_format(), expand = expansion(mult = c(0.2, 0.2))) +
  coord_flip() +
  facet_grid(Category ~ ., scales = "free_y", space = "free_y") +
  labs(title = "Total Profit by Sub-Category",
       subtitle = "Red = loss-making · Grouped by Category",
       x = NULL, y = "Total Profit", caption = "Source: Superstore Dataset") +
  theme_minimal(base_size = 10) +
  theme(plot.title       = element_text(face = "bold", size = 13),
        strip.text       = element_text(face = "bold", color = "white",
                                        margin = margin(3,3,3,3)),
        strip.background = element_rect(fill = "#2c3e50", color = NA),
        legend.position  = "bottom")
```

### Treemap

```{r}
#| fig-height: 5
subcategory_metrics %>%
  mutate(abs_profit = abs(Total_Profit),
         lbl = paste0(`Sub-Category`, "\n", dollar(Total_Profit, accuracy = 1))) %>%
  ggplot(aes(area = abs_profit, fill = Total_Profit, label = lbl, subgroup = Category)) +
  geom_treemap() +
  geom_treemap_subgroup_border(color = "white", linewidth = 2.5) +
  geom_treemap_subgroup_text(place = "topleft", grow = FALSE, reflow = TRUE,
                              colour = "white", fontface = "bold", size = 13, alpha = 0.6) +
  geom_treemap_text(color = "white", place = "centre", reflow = TRUE,
                    size = 8, fontface = "bold") +
  scale_fill_gradient2(low = "#e74c3c", mid = "#f5f5f5", high = "#27ae60",
                       midpoint = 0, labels = dollar_format(), name = "Total Profit") +
  labs(title = "Sub-Category Profitability Treemap",
       subtitle = "Tile size ∝ |Profit| · Green = profitable · Red = loss",
       caption = "Source: Superstore Dataset") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "right")
```

:::

::: callout-warning
**Critical Finding:**
```{r}
#| echo: false
unprofitable_subs <- subcategory_metrics %>% filter(Total_Profit < 0) %>% arrange(Total_Profit)
worst_sub   <- unprofitable_subs$`Sub-Category`[1]
worst_loss  <- dollar(abs(unprofitable_subs$Total_Profit[1]))
n_loss      <- nrow(unprofitable_subs)
loss_labels <- paste(paste0(unprofitable_subs$`Sub-Category`, " (",
                             dollar(unprofitable_subs$Total_Profit), ")"),
                     collapse = ", ")
```
**`r n_loss` sub-categories** generate net losses: `r loss_labels`. The worst is **`r worst_sub`** (–`r worst_loss`).
:::

---

# The Discount Problem

::: {.panel-tabset}

## Profit by Discount Tier

```{r}
#| fig-height: 5

# Discretise discount into tiers (mirrors actual data structure)
discount_data <- superstore %>%
  mutate(
    Profit_Status = ifelse(Profit > 0, "Profitable", "Unprofitable"),
    Discount_Tier = case_when(
      Discount == 0.00 ~ "0%",
      Discount <= 0.10 ~ "1–10%",
      Discount <= 0.20 ~ "11–20%",
      Discount <= 0.30 ~ "21–30%",
      Discount <= 0.50 ~ "31–50%",
      TRUE             ~ "51–80%"
    ),
    Discount_Tier = factor(Discount_Tier,
      levels = c("0%","1–10%","11–20%","21–30%","31–50%","51–80%"))
  )

ggplot(discount_data, aes(x = Discount_Tier, y = Profit, fill = Discount_Tier)) +
  geom_boxplot(alpha = 0.8, outlier.alpha = 0.15, outlier.size = 0.6,
               show.legend = FALSE) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 3.5,
               fill = "white", color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#e74c3c", linewidth = 1.0) +
  scale_fill_manual(values = c("0%"    = "#27ae60","1–10%"  = "#82e0aa",
                               "11–20%" = "#f9e79f","21–30%" = "#f0b27a",
                               "31–50%" = "#e59866","51–80%" = "#e74c3c")) +
  scale_y_continuous(labels = dollar_format(), limits = c(-1000, 1500)) +
  labs(title = "Profit Distribution Drops with Every Discount Tier",
       subtitle = "◆ = mean · Red dashed = break-even · Each step higher in discount = more losses",
       x = "Discount Tier", y = "Profit / Transaction") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```

## Profitability Rate by Tier

```{r}
#| fig-height: 4
disc_bin_summary <- discount_data %>%
  group_by(Discount_Tier, Profit_Status) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Discount_Tier) %>%
  mutate(Pct = Count / sum(Count))

ggplot(disc_bin_summary, aes(x = Discount_Tier, y = Pct, fill = Profit_Status)) +
  geom_col(position = "fill", width = 0.7) +
  geom_text(aes(label = ifelse(Pct > 0.07, paste0(round(Pct*100,0), "%"), "")),
            position = position_fill(vjust = 0.5),
            size = 3.8, color = "white", fontface = "bold") +
  scale_fill_manual(values = c("Profitable" = "#27ae60", "Unprofitable" = "#e74c3c")) +
  scale_y_continuous(labels = percent_format()) +
  labs(title = "Tipping Point: Orders Above 20% Discount Go Majority-Unprofitable",
       subtitle = "Green = profitable · Red = unprofitable",
       x = "Discount Tier", y = "% of Transactions", fill = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")
```

## Heatmap: Category × Discount

```{r}
#| fig-height: 3.5
superstore %>%
  mutate(Discount_Tier = case_when(
    Discount == 0.00 ~ "0%",    Discount <= 0.10 ~ "1–10%",
    Discount <= 0.20 ~ "11–20%", Discount <= 0.30 ~ "21–30%",
    Discount <= 0.50 ~ "31–50%", TRUE             ~ "51–80%"),
    Discount_Tier = factor(Discount_Tier,
      levels = c("0%","1–10%","11–20%","21–30%","31–50%","51–80%"))) %>%
  group_by(Category, Discount_Tier) %>%
  summarise(Avg_Profit = mean(Profit), .groups = "drop") %>%
  ggplot(aes(x = Discount_Tier, y = Category, fill = Avg_Profit)) +
  geom_tile(color = "white", linewidth = 0.9) +
  geom_text(aes(label = dollar(Avg_Profit, accuracy = 1)),
            size = 3.5, fontface = "bold", color = "white") +
  scale_fill_gradient2(low = "#e74c3c", mid = "grey85", high = "#27ae60",
                       midpoint = 0, name = "Avg Profit", labels = dollar_format()) +
  labs(title = "All Categories Suffer at High Discount Levels",
       subtitle = "Red = loss · Green = profit · Damage is consistent across categories",
       x = "Discount Tier", y = NULL) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), panel.grid = element_blank())
```

:::

::: callout-important
**Key Discount Insights:**
```{r}
#| echo: false
disc_summary <- discount_data %>%
  group_by(Profit_Status) %>%
  summarise(Avg_Discount = mean(Discount), .groups = "drop")
avg_p <- percent(disc_summary$Avg_Discount[disc_summary$Profit_Status == "Profitable"],  accuracy = 0.1)
avg_u <- percent(disc_summary$Avg_Discount[disc_summary$Profit_Status == "Unprofitable"], accuracy = 0.1)
pct_loss_top <- disc_bin_summary %>%
  filter(Discount_Tier == "51–80%", Profit_Status == "Unprofitable") %>%
  pull(Pct) %>% percent(accuracy = 1)
```
- Profitable orders average **`r avg_p`** discount; unprofitable orders average **`r avg_u`**.
- At the **51–80%** tier, **`r pct_loss_top`** of transactions are unprofitable.
- **Recommended action:** Hard cap at 30%; require manager approval above 20%.
:::

---

# Regional Performance

```{r}
#| fig-height: 5
regional_metrics <- superstore %>%
  group_by(Region) %>%
  summarise(
    Total_Sales    = sum(Sales),
    Total_Profit   = sum(Profit),
    Avg_Profit     = mean(Profit),
    Profit_Margin  = Total_Profit / Total_Sales * 100,
    Count          = n(),
    Avg_Discount   = mean(Discount),
    Pct_Profitable = mean(Profit > 0) * 100,
    .groups        = "drop"
  ) %>%
  arrange(desc(Total_Profit))

p_reg_profit <- ggplot(regional_metrics,
                       aes(x = reorder(Region, Total_Profit),
                           y = Total_Profit, fill = Region)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = dollar(Total_Profit, accuracy = 1)),
            hjust = -0.1, fontface = "bold", size = 3.8) +
  scale_fill_brewer(palette = "Set3") +
  scale_y_continuous(labels = dollar_format(), expand = expansion(mult = c(0, 0.3))) +
  coord_flip() +
  labs(title = "Total Profit by Region", x = NULL, y = "Total Profit") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_reg_bubble <- ggplot(regional_metrics,
                       aes(x = Avg_Discount, y = Avg_Profit,
                           color = Profit_Margin, size = Total_Sales, label = Region)) +
  geom_point(alpha = 0.9) +
  geom_text_repel(size = 4, fontface = "bold", show.legend = FALSE, box.padding = 0.5) +
  scale_color_gradient(low = "#e74c3c", high = "#27ae60", name = "Margin (%)") +
  scale_size_continuous(labels = dollar_format(), name = "Total Sales", range = c(8, 18)) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Lower Discount = Higher Avg Profit per Region",
       subtitle = "Bubble size ∝ Total Sales · Colour ∝ Profit Margin",
       x = "Average Discount Rate", y = "Average Profit / Transaction") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_reg_profit | p_reg_bubble
```

::: callout-tip
**Regional Findings:**
```{r}
#| echo: false
best_region        <- regional_metrics$Region[1]
best_region_total  <- dollar(regional_metrics$Total_Profit[1])
worst_region       <- regional_metrics$Region[nrow(regional_metrics)]
best_margin_region <- regional_metrics$Region[which.max(regional_metrics$Profit_Margin)]
best_margin_pct    <- paste0(round(max(regional_metrics$Profit_Margin), 1), "%")
```
- **`r best_region`** leads total profit (**`r best_region_total`**); **`r best_margin_region`** has the highest margin at **`r best_margin_pct`**.
- **`r worst_region`** lags both metrics — higher discounts are likely the driver.
:::

---

# Temporal Analysis

::: {.panel-tabset}

## Monthly Sales & Profit

```{r}
#| fig-height: 5
monthly_by_cat <- superstore %>%
  group_by(YearMonth, Category) %>%
  summarise(Total_Sales = sum(Sales), Total_Profit = sum(Profit), .groups = "drop")

p_s <- ggplot(monthly_by_cat, aes(x = YearMonth, y = Total_Sales, color = Category)) +
  geom_line(linewidth = 1.1, alpha = 0.85) +
  geom_point(size = 1.4, alpha = 0.6) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_brewer(palette = "Set1") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Monthly Sales by Category", x = NULL, y = "Total Sales", color = NULL) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 40, hjust = 1), legend.position = "bottom")

p_p <- ggplot(monthly_by_cat, aes(x = YearMonth, y = Total_Profit, color = Category)) +
  geom_line(linewidth = 1.1, alpha = 0.85) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.7) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_brewer(palette = "Set1") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Monthly Profit by Category", x = NULL, y = "Total Profit", color = NULL) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 40, hjust = 1), legend.position = "bottom")

p_s / p_p
```

## Profit Margin Trend

```{r}
#| fig-height: 4
monthly_overall <- superstore %>%
  group_by(YearMonth) %>%
  summarise(Total_Sales = sum(Sales), Total_Profit = sum(Profit),
            Profit_Margin = Total_Profit / Total_Sales * 100, .groups = "drop")

ggplot(monthly_overall, aes(x = YearMonth, y = Profit_Margin)) +
  geom_area(fill = "#3498db", alpha = 0.15) +
  geom_line(color = "#2c3e50", linewidth = 1.2) +
  geom_smooth(method = "loess", se = TRUE, color = "#e74c3c",
              fill = "#e74c3c", alpha = 0.12, linewidth = 1.0) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#e74c3c", linewidth = 0.8) +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Monthly Overall Profit Margin",
       subtitle = "Red LOESS = long-run trend · Shaded band = 95% CI",
       x = NULL, y = "Profit Margin (%)") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_text(angle = 40, hjust = 1))
```

## Seasonal Heatmap

```{r}
#| fig-height: 3.5
superstore %>%
  group_by(Year, Month) %>%
  summarise(Total_Profit = sum(Profit), .groups = "drop") %>%
  ggplot(aes(x = Month, y = factor(Year), fill = Total_Profit)) +
  geom_tile(color = "white", linewidth = 0.8) +
  geom_text(aes(label = dollar(Total_Profit, accuracy = 0.1, scale = 0.001, suffix = "K")),
            size = 3.0, fontface = "bold", color = "white") +
  scale_fill_gradient2(low = "#e74c3c", mid = "grey88", high = "#27ae60",
                       midpoint = 0, name = "Total Profit", labels = dollar_format()) +
  labs(title = "Q4 (Oct–Dec) Consistently Delivers the Strongest Profit",
       subtitle = "Green = high profit · Red = low / negative profit",
       x = "Month", y = "Year") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), panel.grid = element_blank())
```

## Interactive Time Series

```{r}
#| fig-height: 4

# Build hover text as a plain R character vector — no plotly template tokens
monthly_hover <- monthly_by_cat %>%
  mutate(hover_label = paste0(
    "<b>", Category, "</b><br>",
    format(YearMonth, "%b %Y"), "<br>",
    "Profit: ", dollar(Total_Profit, accuracy = 1)
  ))

plot_ly(monthly_hover,
        x    = ~YearMonth,
        y    = ~Total_Profit,
        color = ~Category,
        colors = RColorBrewer::brewer.pal(3, "Set1"),
        type  = "scatter",
        mode  = "lines+markers",
        line  = list(width = 2),
        marker = list(size = 4),
        text  = ~hover_label,
        hoverinfo = "text") %>%           # ← use hoverinfo, not hovertemplate
  layout(
    title = list(
      text = "<b>Monthly Profit by Category</b><br><sup>Use range buttons or drag slider to zoom · Hover for details</sup>"
    ),
    xaxis = list(
      title = "",
      rangeslider = list(visible = TRUE, thickness = 0.06),
      rangeselector = list(buttons = list(
        list(count = 6,  label = "6M",  step = "month", stepmode = "backward"),
        list(count = 12, label = "1Y",  step = "month", stepmode = "backward"),
        list(step = "all", label = "All")
      ))
    ),
    yaxis     = list(title = "Total Profit ($)", tickformat = "$,.0f"),
    legend    = list(orientation = "h", x = 0, y = -0.3),
    hovermode = "closest",
    plot_bgcolor  = "rgba(0,0,0,0)",
    paper_bgcolor = "rgba(0,0,0,0)"
  )
```

:::

::: callout-tip
**Temporal Findings:**
```{r}
#| echo: false
heatmap_data <- superstore %>%
  group_by(Year, Month) %>%
  summarise(Total_Profit = sum(Profit), .groups = "drop")
best_month  <- heatmap_data %>% group_by(Month) %>%
  summarise(AvgP = mean(Total_Profit)) %>% slice_max(AvgP, n = 1)
worst_month <- heatmap_data %>% group_by(Month) %>%
  summarise(AvgP = mean(Total_Profit)) %>% slice_min(AvgP, n = 1)
```
- **`r best_month$Month`** is consistently the strongest month (avg `r dollar(best_month$AvgP)` profit).
- **`r worst_month$Month`** is the weakest month on average (`r dollar(worst_month$AvgP)`).
- The LOESS trend confirms the business is gradually improving its profit margin.
:::

---

# Customer Segment Analysis

::: {.panel-tabset}

## Sales vs Profit Quadrant by State

```{r}
#| fig-height: 6

# State-level metrics with percentile ranking (mirrors Tableau Image 5 approach)
state_metrics <- superstore %>%
  group_by(`State/Province`, Region) %>%
  summarise(
    Total_Sales  = sum(Sales),
    Total_Profit = sum(Profit),
    .groups      = "drop"
  ) %>%
  mutate(
    Sales_Pctile  = percent_rank(Total_Sales)  * 100,
    Profit_Pctile = percent_rank(Total_Profit) * 100
  )

# Quadrant labels
med_s <- 50; med_p <- 50

ggplot(state_metrics,
       aes(x = Sales_Pctile, y = Profit_Pctile,
           color = Total_Profit, size = Total_Sales)) +
  # Quadrant shading
  annotate("rect", xmin = med_s, xmax = Inf,  ymin = med_p, ymax = Inf,
           fill = "#27ae60", alpha = 0.06) +
  annotate("rect", xmin = -Inf, xmax = med_s, ymin = -Inf, ymax = med_p,
           fill = "#bdc3c7", alpha = 0.08) +
  annotate("rect", xmin = med_s, xmax = Inf,  ymin = -Inf, ymax = med_p,
           fill = "#e74c3c", alpha = 0.06) +
  annotate("rect", xmin = -Inf, xmax = med_s, ymin = med_p, ymax = Inf,
           fill = "#3498db", alpha = 0.06) +
  # Quadrant reference lines
  geom_vline(xintercept = med_s, linetype = "dashed", color = "grey50", linewidth = 0.8) +
  geom_hline(yintercept = med_p, linetype = "dashed", color = "grey50", linewidth = 0.8) +
  # Quadrant labels
  annotate("text", x = 75, y = 97, label = "High Sales\nHigh Profit", fontface = "bold",
           color = "#27ae60", size = 3.2, hjust = 0.5) +
  annotate("text", x = 75, y =  3, label = "High Sales\nLow Profit",  fontface = "bold",
           color = "#e74c3c", size = 3.2, hjust = 0.5) +
  annotate("text", x = 10, y = 97, label = "Low Sales\nHigh Profit",  fontface = "bold",
           color = "#3498db", size = 3.2, hjust = 0.5) +
  annotate("text", x = 10, y =  3, label = "Low Sales\nLow Profit",   fontface = "bold",
           color = "grey50",  size = 3.2, hjust = 0.5) +
  # Points
  geom_point(alpha = 0.85) +
  geom_text_repel(aes(label = `State/Province`), size = 2.5, max.overlaps = 14,
                  show.legend = FALSE, box.padding = 0.3) +
  # Colour = profit (orange-to-blue like Tableau), Size = total sales
  scale_color_gradient2(low = "#e67e22", mid = "grey80", high = "#2980b9",
                        midpoint = 0, name = "Total Profit ($)",
                        labels = dollar_format()) +
  scale_size_continuous(range = c(2, 10), name = "Total Sales ($)",
                        labels = dollar_format()) +
  scale_x_continuous(labels = function(x) paste0(x, "%"), limits = c(-2, 108)) +
  scale_y_continuous(labels = function(x) paste0(x, "%"), limits = c(-2, 108)) +
  labs(title = "Sales vs Profit Quadrant by State",
       subtitle = "Top-right = star states · Bottom-right = high sales but value-destroying",
       x = "Percentile of Sales", y = "Percentile of Profit",
       caption = "Source: Superstore Dataset") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold", size = 13),
        legend.position = "right")
```

## Pareto — Which Customers Drive Profit?

```{r}
#| fig-height: 5

# Customer-level cumulative profit Pareto (mirrors Tableau Image 6)
customer_pareto <- superstore %>%
  group_by(`Customer ID`, `Customer Name`, Segment) %>%
  summarise(Total_Profit = sum(Profit), .groups = "drop") %>%
  arrange(desc(Total_Profit)) %>%
  mutate(
    rank       = row_number(),
    n_total    = n(),
    cum_profit = cumsum(Total_Profit),
    total_pos  = sum(pmax(Total_Profit, 0)),
    cum_pct_profit   = cum_profit / sum(Total_Profit) * 100,
    cum_pct_customer = rank / n_total * 100
  )

# Find 80% threshold
thresh_cust <- customer_pareto %>%
  filter(cum_pct_profit >= 80) %>% slice(1)

ggplot(customer_pareto,
       aes(x = cum_pct_customer, y = cum_pct_profit)) +
  # Shaded area under curve
  geom_area(fill = "#2980b9", alpha = 0.15) +
  geom_line(color = "#2980b9", linewidth = 1.6) +
  geom_point(size = 0.8, color = "#2980b9", alpha = 0.5) +
  # 80% reference lines
  geom_hline(yintercept = 80, linetype = "dashed", color = "#e67e22", linewidth = 0.9) +
  geom_vline(xintercept = thresh_cust$cum_pct_customer,
             linetype = "dashed", color = "#e67e22", linewidth = 0.9) +
  # Annotation at threshold
  annotate("point",
           x = thresh_cust$cum_pct_customer, y = 80,
           shape = 21, size = 5, fill = "white", color = "#e67e22", stroke = 1.5) +
  annotate("label",
           x = thresh_cust$cum_pct_customer + 2, y = 75,
           label = sprintf("Top %.0f%% of customers\ngenerate 80%% of profit",
                           thresh_cust$cum_pct_customer),
           hjust = 0, size = 3.2, fontface = "bold", color = "#e67e22",
           fill = "white", label.size = 0.3) +
  scale_x_continuous(labels = function(x) paste0(x, "%"), limits = c(0, 105)) +
  scale_y_continuous(labels = function(x) paste0(x, "%"), limits = c(0, 105)) +
  labs(title = "Pareto Curve: Customer Profit Concentration",
       subtitle = "A small share of customers generates the vast majority of profit",
       x = "Cumulative % of Customers (ranked by profit)",
       y = "Cumulative % of Total Profit",
       caption = "Source: Superstore Dataset") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold", size = 13))
```

## Segment Scorecard

```{r}
#| fig-height: 4
segment_metrics <- superstore %>%
  group_by(Segment) %>%
  summarise(
    Total_Sales    = sum(Sales),
    Total_Profit   = sum(Profit),
    Avg_Profit     = mean(Profit),
    Profit_Margin  = Total_Profit / Total_Sales * 100,
    Avg_Discount   = mean(Discount),
    Count          = n(),
    Pct_Profitable = mean(Profit > 0) * 100,
    .groups        = "drop"
  ) %>%
  arrange(desc(Total_Profit))

p_seg_bar <- ggplot(segment_metrics,
                    aes(x = reorder(Segment, Avg_Profit),
                        y = Avg_Profit, fill = Segment)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = dollar(Avg_Profit, accuracy = 0.01)),
            hjust = -0.1, fontface = "bold", size = 3.8) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels = dollar_format(), expand = expansion(mult = c(0, 0.35))) +
  coord_flip() +
  labs(title = "Avg Profit / Transaction", x = NULL, y = "Avg Profit") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_seg_pct <- ggplot(segment_metrics,
                    aes(x = reorder(Segment, Pct_Profitable),
                        y = Pct_Profitable, fill = Segment)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = paste0(round(Pct_Profitable, 1), "%")),
            hjust = -0.1, fontface = "bold", size = 3.8) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0, 100), expand = expansion(mult = c(0, 0.2))) +
  coord_flip() +
  labs(title = "% Profitable Transactions", x = NULL, y = "% Profitable") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

p_seg_bar | p_seg_pct
```

:::

::: callout-note
**Segment Insights:**
```{r}
#| echo: false
best_seg_avg   <- segment_metrics$Segment[which.max(segment_metrics$Avg_Profit)]
best_avg_val   <- dollar(max(segment_metrics$Avg_Profit), accuracy = 0.01)
best_seg_total <- segment_metrics$Segment[1]
best_tot_val   <- dollar(segment_metrics$Total_Profit[1])
```
- **`r best_seg_avg`** achieves the highest avg profit per transaction (**`r best_avg_val`**).
- **`r best_seg_total`** generates the most total profit (**`r best_tot_val`**) through volume.
:::

---

# Key Insights & Recommendations

## Interactive Profitability Dashboard

```{r}
#| fig-height: 4

# Pre-build hover strings as plain R character — avoids %{text} rendering bug
cat_hover <- category_metrics %>%
  mutate(tip = paste0(
    "<b>", Category, "</b><br>",
    "Total Profit: ", dollar(Total_Profit), "<br>",
    "Margin: ", round(Profit_Margin, 1), "%<br>",
    "Transactions: ", format(Count, big.mark = ",")
  ))

reg_hover <- regional_metrics %>%
  mutate(tip = paste0(
    "<b>", Region, "</b><br>",
    "Total Profit: ", dollar(Total_Profit), "<br>",
    "Margin: ", round(Profit_Margin, 1), "%<br>",
    "Transactions: ", format(Count, big.mark = ",")
  ))

seg_hover <- segment_metrics %>%
  mutate(tip = paste0(
    "<b>", Segment, "</b><br>",
    "Avg Profit: ", dollar(Avg_Profit, accuracy = 0.01), "<br>",
    "% Profitable: ", round(Pct_Profitable, 1), "%"
  ))

fig_cat <- plot_ly(cat_hover,
  x = ~reorder(Category, Total_Profit), y = ~Total_Profit,
  type = "bar", color = ~Category,
  colors = RColorBrewer::brewer.pal(3, "Set2"),
  text = ~tip, hoverinfo = "text") %>%
  layout(title = "<b>Category</b>", showlegend = FALSE,
         xaxis = list(title = ""),
         yaxis = list(title = "Total Profit ($)", tickformat = "$,.0f"))

fig_reg <- plot_ly(reg_hover,
  x = ~reorder(Region, Total_Profit), y = ~Total_Profit,
  type = "bar", color = ~Region,
  colors = RColorBrewer::brewer.pal(4, "Set3"),
  text = ~tip, hoverinfo = "text") %>%
  layout(title = "<b>Region</b>", showlegend = FALSE,
         xaxis = list(title = ""),
         yaxis = list(title = "Total Profit ($)", tickformat = "$,.0f"))

fig_seg <- plot_ly(seg_hover,
  x = ~reorder(Segment, Avg_Profit), y = ~Avg_Profit,
  type = "bar", color = ~Segment,
  colors = RColorBrewer::brewer.pal(3, "Dark2"),
  text = ~tip, hoverinfo = "text") %>%
  layout(title = "<b>Segment (Avg Profit)</b>", showlegend = FALSE,
         xaxis = list(title = ""),
         yaxis = list(title = "Avg Profit ($)", tickformat = "$,.2f"))

subplot(fig_cat, fig_reg, fig_seg, nrows = 1,
        titleX = TRUE, titleY = TRUE, shareY = FALSE, margin = 0.06) %>%
  layout(
    title = list(
      text = "<b>Superstore Profitability — Category · Region · Segment</b><br><sup>Hover bars for details</sup>",
      x = 0.5),
    plot_bgcolor  = "rgba(0,0,0,0)",
    paper_bgcolor = "rgba(0,0,0,0)")
```

## Summary Metrics

```{r}
pct_loss_txns <- mean(superstore$Profit < 0)

tibble(
  Metric = c(
    "Total Transactions", "Total Sales Revenue", "Total Profit",
    "Overall Profit Margin", "Unprofitable Transactions",
    "Most Profitable Category", "Least Profitable Sub-Category",
    "Best Region (Total Profit)", "Best Region (Margin)",
    "Avg Discount — Profitable", "Avg Discount — Unprofitable"
  ),
  Value = c(
    format(nrow(superstore), big.mark = ","),
    dollar(sum(superstore$Sales)),
    dollar(sum(superstore$Profit)),
    paste0(round(sum(superstore$Profit)/sum(superstore$Sales)*100, 2), "%"),
    paste0(format(sum(superstore$Profit < 0), big.mark = ","),
           "  (", round(pct_loss_txns*100, 1), "% of orders)"),
    superstore %>% group_by(Category) %>%
      summarise(tp = sum(Profit), .groups = "drop") %>%
      slice_max(tp, n=1) %>% pull(Category),
    superstore %>% group_by(`Sub-Category`) %>%
      summarise(tp = sum(Profit), .groups = "drop") %>%
      slice_min(tp, n=1) %>% pull(`Sub-Category`),
    superstore %>% group_by(Region) %>%
      summarise(tp = sum(Profit), .groups = "drop") %>%
      slice_max(tp, n=1) %>% pull(Region),
    superstore %>% group_by(Region) %>%
      summarise(ts = sum(Sales), tp = sum(Profit), .groups = "drop") %>%
      mutate(margin = tp/ts) %>% slice_max(margin, n=1) %>% pull(Region),
    percent(mean(superstore$Discount[superstore$Profit  > 0]), accuracy = 0.1),
    percent(mean(superstore$Discount[superstore$Profit <= 0]), accuracy = 0.1)
  )
) %>%
  kbl(col.names = c("Metric", "Value"), align = c("l","r"),
      caption = "Superstore — Performance Summary") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white")
```

::: {.callout-note appearance="default" icon="true"}
## Business Recommendations

```{r}
#| echo: false
avg_p_rec <- percent(mean(superstore$Discount[superstore$Profit  > 0]), accuracy = 0.1)
avg_u_rec <- percent(mean(superstore$Discount[superstore$Profit <= 0]), accuracy = 0.1)
loss_sub_labels <- subcategory_metrics %>%
  filter(Total_Profit < 0) %>% arrange(Total_Profit) %>%
  mutate(lbl = paste0(`Sub-Category`, " (", dollar(Total_Profit), ")")) %>% pull(lbl)
top_subs <- subcategory_metrics %>%
  slice_max(Total_Profit, n = 3) %>% pull(`Sub-Category`)
best_r_rec  <- regional_metrics$Region[1]
worst_r_rec <- regional_metrics$Region[nrow(regional_metrics)]
best_seg_rec <- segment_metrics$Segment[which.max(segment_metrics$Avg_Profit)]
vol_seg_rec  <- segment_metrics$Segment[1]
```

**1. Discount Cap at 30%**  
Profitable orders average `r avg_p_rec` discount; unprofitable orders average `r avg_u_rec`. Enforce a hard cap at 30%, require manager approval above 20%.

**2. Product Line Optimisation**  
Loss-making sub-categories requiring action: `r paste(loss_sub_labels, collapse = " · ")`. Invest in top performers: **`r paste(top_subs, collapse = "**, **")`**.

**3. Regional Strategy**  
`r best_r_rec` leads — use as the pricing benchmark. `r worst_r_rec` requires root-cause analysis this quarter.

**4. Segment & Customer Development**  
`r best_seg_rec` earns the highest margin per transaction — grow through premium bundles. The Pareto curve shows a small share of customers generates most profit — priorities retention programmes for that top tier.
:::
