---
title: "In-class Exercise 2"
subtitle: "Analysing Superstore Dataset"
author: "Thi Thanh Van Tran"
date: "April 18, 2026"
date-modified: "last-modified"
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    number-sections: true
    number-depth: 3
    code-fold: true
    code-summary: "Show Code"
    code-tools: true
    code-copy: true
    theme: flatly
    highlight-style: github
execute:
  eval: true
  echo: true
  warning: false
  message: false
  freeze: false
  fig-width: 10
  fig-height: 6
  opts_chunk:
    out.width: "100%"
    fig.align: "center"
    dpi: 120 
---

# Tableau Dashboard

```{=html}
<div style="width: 100%; height: 400px;">
  <iframe
    src="https://public.tableau.com/views/in-class_Ex02_17771009333620/Dashboard1?:showVizHome=no&:embed=true"
    style="width: 100%; height: 100%; border: none;"
    loading="lazy"
    allowfullscreen>
  </iframe>
</div>
```

# Getting Started

## Installing and Loading the Required Libraries

::: callout-note
**Note:** Ensure that the [pacman](https://cran.r-project.org/web/packages/pacman/) package has already been installed before running this document.
:::

```{r}
#| code-fold: false

pacman::p_load(
  # Data import & wrangling
  readxl, tidyverse, lubridate,

  # Static visualisation
  ggplot2, scales, patchwork,
  ggridges, ggrepel, ggthemes,
  viridis, RColorBrewer, treemapify,

  # Interactive visualisation
  plotly, ggiraph, DT,

  # Tables
  gt, kableExtra,

  # Missing data
  naniar,

  # Animation
  gganimate, gifski
)
```

## Importing Data

The dataset is loaded from an Excel file using `read_excel()` from the **readxl** package. Four derived time variables are added for temporal analysis.

```{r}
#| code-fold: false
superstore <- read_excel("data/Superstore.xls")
# Add derived time variables for temporal analysis
superstore <- superstore %>%
  mutate(
    Year      = year(`Order Date`),
    Month     = month(`Order Date`, label = TRUE, abbr = TRUE),
    YearMonth = floor_date(`Order Date`, "month"),
    Quarter   = paste0("Q", quarter(`Order Date`)),
    ShipDays  = as.numeric(difftime(`Ship Date`, `Order Date`, units = "days"))
  )
```

---

# Data Overview

::: {.panel-tabset}

##  Variable Structure

```{r}
#| fig-height: 7
original_cols <- superstore   # that's it, no pipe

# Then your sapply + enframe code works
var_types <- sapply(original_cols, function(x) class(x)[1]) %>% 
  enframe(name = "Variable", value = "Type") %>% 
  mutate(
    Type_Clean = case_when(
      Type %in% c("numeric", "integer", "double") ~ "Numeric",
      Type %in% c("character", "factor") ~ "Character",
      str_detect(Type, "POSIXct|POSIXt|Date") ~ "Date / Time",
      TRUE ~ "Other"
    )
  ) %>% 
  arrange(Type_Clean, Variable)

type_counts <- var_types %>%
  count(Type_Clean) %>%
  mutate(pct = n / sum(n))

# Left: donut-style bar by type
p_types <- ggplot(type_counts,
                  aes(x = reorder(Type_Clean, n), y = n, fill = Type_Clean)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = paste0(n, " vars")),
            hjust = -0.15, fontface = "bold", size = 4.5) +
  scale_fill_viridis_d(option = "D", begin = 0.2, end = 0.75) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.30))) +
  coord_flip() +
  labs(
    title    = "Variable Types",
    subtitle = sprintf("Total: %d original variables", ncol(original_cols)),
    x = NULL, y = "Count"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", size = 14))

# Right: coloured tile list of all variables
p_varlist <- ggplot(var_types,
                    aes(y = reorder(Variable, desc(Variable)),
                        x = 1, fill = Type_Clean)) +
  geom_tile(color = "white", linewidth = 0.6) +
  geom_text(aes(label = Variable), x = 1, size = 3.0, color = "white",
            fontface = "bold") +
  scale_fill_viridis_d(option = "D", begin = 0.2, end = 0.75, name = "Type") +
  scale_x_continuous(expand = c(0, 0)) +
  labs(title = "All Variables", x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text  = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", size = 14),
    legend.position = "right"
  )

p_types + p_varlist + plot_layout(widths = c(1, 1.8))
```

**Observation:** The dataset has `r ncol(original_cols)` variables — `r sum(var_types$Type_Clean == "Numeric")` numeric, `r sum(var_types$Type_Clean == "Character")` character, and `r sum(var_types$Type_Clean == "Date / Time")` date/time fields. Five additional derived columns (Year, Month, YearMonth, Quarter, ShipDays) have been added for analysis.

Below is an interactive, scrollable preview of the raw data:

```{r}
DT::datatable(
  head(superstore %>% select(-Year, -Month, -YearMonth, -Quarter, -ShipDays), 200),
  options = list(
    pageLength = 10,
    scrollX    = TRUE,
    dom        = "ltipr",
    columnDefs = list(list(className = "dt-center", targets = "_all"))
  ),
  filter  = "top",
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Superstore Dataset — first 200 rows (filterable & sortable)"
  )
)
```

## Missing Values

```{r}
#| fig-height: 6

miss_df <- original_cols %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing") %>%
  mutate(
    Pct_Missing = Missing / nrow(superstore) * 100,
    Status      = ifelse(Missing == 0, "Complete", "Has Missing")
  ) %>%
  arrange(desc(Missing))

n_complete <- sum(miss_df$Missing == 0)
n_missing  <- sum(miss_df$Missing > 0)

p_miss <- ggplot(miss_df,
                 aes(x = reorder(Variable, -Pct_Missing),
                     y = Pct_Missing, fill = Status)) +
  geom_col(width = 0.7, alpha = 0.9) +
  geom_text(
    aes(label = ifelse(Missing > 0,
                       sprintf("%.1f%%  (n=%d)", Pct_Missing, Missing),
                       "✓ Complete")),
    hjust = -0.05, size = 3.4, fontface = "bold"
  ) +
  scale_fill_manual(values = c("Complete" = "#27ae60", "Has Missing ✗" = "#e74c3c")) +
  scale_y_continuous(labels  = percent_format(scale = 1),
                     expand  = expansion(mult = c(0, 0.3))) +
  coord_flip() +
  labs(
    title    = "Missing Values by Variable",
    subtitle = sprintf("%d of %d variables are fully complete  |  %d variable(s) have missing data",
                       n_complete, nrow(miss_df), n_missing),
    x = NULL, y = "% Missing", fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 14),
    legend.position = "bottom"
  )

print(p_miss)
```

::: callout-tip
**Observation:** `r if(n_missing == 0) "The dataset is perfectly complete — zero missing values detected across all 21 original variables." else paste0("⚠️ ", n_missing, " variable(s) contain missing values that need attention before modelling.")`
:::

## Key Distributions

```{r}
#| fig-height: 12

# --- Sales (log10 scale) ---
p_sales <- ggplot(superstore, aes(x = Sales)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 50, fill = "#3498db", alpha = 0.75, color = "white") +
  geom_density(color = "#1a252f", linewidth = 1.2) +
  scale_x_continuous(labels = dollar_format(), trans = "log10") +
  labs(title = "Sales Distribution",
       subtitle = "Log₁₀ scale — reveals right-skewed structure",
       x = "Sales (log₁₀)", y = "Density") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

# --- Profit ---
prof_mean   <- mean(superstore$Profit)
prof_median <- median(superstore$Profit)

p_profit <- ggplot(superstore, aes(x = Profit)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 70, fill = "#2ecc71", alpha = 0.65, color = "white") +
  geom_density(color = "#1e8449", linewidth = 1.2) +
  geom_vline(xintercept = 0,           linetype = "dashed", color = "#e74c3c", linewidth = 1.0) +
  geom_vline(xintercept = prof_mean,   linetype = "dotted", color = "#2980b9", linewidth = 0.9) +
  geom_vline(xintercept = prof_median, linetype = "dotted", color = "#8e44ad", linewidth = 0.9) +
  annotate("text", x = prof_mean   + 5, y = Inf, vjust = 1.5, hjust = 0,
           label = paste0("Mean: ", dollar(prof_mean, accuracy = 0.01)),
           color = "#2980b9", size = 3.2, fontface = "bold") +
  annotate("text", x = prof_median + 5, y = Inf, vjust = 3.5, hjust = 0,
           label = paste0("Median: ", dollar(prof_median, accuracy = 0.01)),
           color = "#8e44ad", size = 3.2, fontface = "bold") +
  scale_x_continuous(labels = dollar_format()) +
  labs(title = "Profit Distribution",
       subtitle = "Red dashed = break-even | Blue = mean | Purple = median",
       x = "Profit per Transaction", y = "Density") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

# --- Discount ---
p_discount <- ggplot(superstore, aes(x = Discount)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, fill = "#e67e22", alpha = 0.75, color = "white") +
  geom_density(color = "#a04000", linewidth = 1.2) +
  scale_x_continuous(labels = percent_format()) +
  labs(title = "Discount Distribution",
       subtitle = "Clustered values reveal structured discount tiers",
       x = "Discount Rate", y = "Density") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

# --- Quantity ---
p_qty <- ggplot(superstore, aes(x = Quantity)) +
  geom_bar(fill = "#9b59b6", alpha = 0.85, color = "white") +
  scale_y_continuous(labels = comma_format()) +
  labs(title = "Quantity Distribution",
       subtitle = "Most orders are 1–4 units",
       x = "Units per Order", y = "Count") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

# --- Shipping Days ---
p_ship <- ggplot(superstore, aes(x = ShipDays)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 15, fill = "#1abc9c", alpha = 0.75, color = "white") +
  geom_density(color = "#0e6655", linewidth = 1.2) +
  labs(title = "Shipping Duration Distribution",
       subtitle = "Days from order to ship",
       x = "Days to Ship", y = "Density") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))

# --- Ridge: Profit by Category ---
p_ridge <- ggplot(superstore, aes(x = Profit, y = Category, fill = Category)) +
  geom_density_ridges(
    alpha = 0.75, scale = 1.3,
    quantile_lines = TRUE, quantiles = c(0.25, 0.5, 0.75),
    jittered_points = FALSE
  ) +
  geom_vline(xintercept = 0, color = "#e74c3c", linetype = "dashed", linewidth = 0.9) +
  scale_fill_brewer(palette = "Set2") +
  scale_x_continuous(labels = dollar_format(), limits = c(-300, 600)) +
  labs(
    title    = "Profit Distribution by Category (Ridge Plot)",
    subtitle = "Lines show Q1 / Median / Q3  |  Red dashed = break-even",
    x = "Profit per Transaction", y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none")

(p_sales | p_profit) /
(p_discount | p_qty | p_ship) /
p_ridge
```

**Key Observations:**

- **Sales** is strongly right-skewed — a handful of high-value orders dominate total revenue; the log scale reveals the bulk of transactions sit in the \$10–\$1,000 range.
- **Profit** centres near zero with a heavy left tail — many transactions generate losses. Mean (`r dollar(prof_mean, accuracy = 0.01)`) exceeds median (`r dollar(prof_median, accuracy = 0.01)`), pulled right by outliers.
- **Discount** is not uniformly distributed — distinct spikes at 0 %, 20 %, 30 %, 40 %, 50 % confirm a structured tiered discount policy.
- **Quantity** peaks at 2–3 units; very few orders exceed 10 units.
- **Shipping** is concentrated around 3–5 days with a small cluster of 0-day same-day shipments.
- **Ridge by Category**: Technology shows the widest profitable tail; Furniture has the most loss-making transactions.

## Data Summary

```{r}
#| code-fold: false

# Structure of the dataset
glimpse(superstore %>% select(-Year, -Month, -YearMonth, -Quarter, -ShipDays))
```

```{r}
# Numerical summary of key financial & operational variables
superstore %>%
  select(Sales, Quantity, Discount, Profit, ShipDays) %>%
  summary()
```

:::

---

# Profitability Analysis

## By Category

```{r}
#| fig-height: 9

category_metrics <- superstore %>%
  group_by(Category) %>%
  summarise(
    Total_Profit   = sum(Profit),
    Total_Sales    = sum(Sales),
    Avg_Profit     = mean(Profit),
    Median_Profit  = median(Profit),
    Profit_Margin  = Total_Profit / Total_Sales * 100,
    Count          = n(),
    Pct_Profitable = mean(Profit > 0) * 100,
    .groups        = "drop"
  ) %>%
  arrange(desc(Total_Profit))

# Panel A: Boxplot profit distribution
p_cat_box <- ggplot(superstore,
                    aes(x = reorder(Category, Profit, FUN = median),
                        y = Profit, fill = Category)) +
  geom_boxplot(alpha = 0.75, outlier.alpha = 0.25, outlier.size = 1,
               outlier.color = "grey60") +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4.5,
               fill = "white", color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#e74c3c", linewidth = 0.8) +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels = dollar_format()) +
  coord_flip() +
  labs(
    title    = "Profit Distribution by Category",
    subtitle = "Diamond = mean  |  Box = IQR  |  Whiskers = 1.5×IQR  |  Red dashed = break-even",
    x = NULL, y = "Profit per Transaction"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")

# Panel B: Total profit bar
p_cat_total <- ggplot(category_metrics,
                      aes(x = reorder(Category, Total_Profit),
                          y = Total_Profit, fill = Category)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = dollar(Total_Profit, accuracy = 1)),
            hjust = -0.1, size = 4.2, fontface = "bold") +
  scale_fill_brewer(palette = "Set2") +
  scale_y_continuous(labels  = dollar_format(),
                     expand  = expansion(mult = c(0, 0.28))) +
  coord_flip() +
  labs(title = "Total Profit by Category", x = NULL, y = "Total Profit") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))

# Panel C: Margin vs % Profitable bubble
p_cat_bubble <- ggplot(category_metrics,
                       aes(x = Profit_Margin, y = Pct_Profitable,
                           color = Category, size = Total_Sales,
                           label = Category)) +
  geom_point(alpha = 0.85) +
  geom_text_repel(size = 4, fontface = "bold", show.legend = FALSE,
                  box.padding = 0.5) +
  scale_color_brewer(palette = "Set2") +
  scale_size_continuous(labels = dollar_format(), name = "Total Sales",
                        range = c(8, 18)) +
  scale_x_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%"),
                     limits = c(50, 100)) +
  labs(
    title    = "Profit Margin vs % Profitable Transactions",
    subtitle = "Bubble size ∝ Total Sales",
    x = "Overall Profit Margin (%)",
    y = "% of Transactions with Profit > 0"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "none")

(p_cat_box | p_cat_total) / p_cat_bubble
```

::: callout-tip
**Findings:**

```{r}
#| echo: false
best_cat   <- category_metrics$Category[1]
best_total <- dollar(category_metrics$Total_Profit[1])
best_margin <- paste0(round(category_metrics$Profit_Margin[1], 1), "%")
```

- **`r best_cat`** leads with **`r best_total`** in total profit and a **`r best_margin`** profit margin, despite having fewer transactions than Office Supplies.
- The bubble chart reveals that higher margin does not always mean higher total profit — transaction volume matters.
- **Furniture** shows the lowest % of profitable transactions, indicating a systemic pricing or discounting problem in that category.
:::

## By Sub-Category

::: {.panel-tabset}

### Lollipop Chart

```{r}
#| fig-height: 9

subcategory_metrics <- superstore %>%
  group_by(`Sub-Category`, Category) %>%
  summarise(
    Total_Profit = sum(Profit),
    Total_Sales  = sum(Sales),
    Avg_Profit   = mean(Profit),
    Count        = n(),
    .groups      = "drop"
  ) %>%
  arrange(Total_Profit)

# Colour flag
subcategory_metrics <- subcategory_metrics %>%
  mutate(Profitable = Total_Profit > 0)

ggplot(subcategory_metrics,
       aes(x = reorder(`Sub-Category`, Total_Profit),
           y = Total_Profit, color = Profitable)) +
  geom_segment(aes(xend = reorder(`Sub-Category`, Total_Profit), yend = 0),
               linewidth = 1.3, alpha = 0.7) +
  geom_point(size = 5.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.6) +
  geom_text(
    aes(label = dollar(Total_Profit, accuracy = 1),
        hjust = ifelse(Total_Profit >= 0, -0.18, 1.18)),
    size = 3.0, fontface = "bold"
  ) +
  scale_color_manual(values = c("TRUE" = "#27ae60", "FALSE" = "#e74c3c"),
                     labels  = c("FALSE" = "Loss-making", "TRUE" = "Profitable"),
                     name    = "Status") +
  scale_y_continuous(labels = dollar_format(),
                     expand = expansion(mult = c(0.2, 0.2))) +
  coord_flip() +
  facet_grid(Category ~ ., scales = "free_y", space = "free_y") +
  labs(
    title    = "Total Profit by Sub-Category",
    subtitle = "Grouped by main Category  |  Red = loss-making sub-categories",
    x = NULL, y = "Total Profit",
    caption  = "Source: Superstore Dataset"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 15),
    strip.text      = element_text(face = "bold", size = 11,
                                   color = "white",
                                   margin = margin(4, 4, 4, 4)),
    strip.background = element_rect(fill = "#2c3e50", color = NA),
    legend.position  = "bottom"
  )
```

### Treemap

```{r}
#| fig-height: 7

subcategory_tm <- subcategory_metrics %>%
  mutate(
    abs_profit = abs(Total_Profit),
    tile_label = paste0(`Sub-Category`, "\n",
                        dollar(Total_Profit, accuracy = 1))
  )

ggplot(subcategory_tm,
       aes(area      = abs_profit,
           fill      = Total_Profit,
           label     = tile_label,
           subgroup  = Category)) +
  geom_treemap() +
  geom_treemap_subgroup_border(color = "white", linewidth = 3) +
  geom_treemap_subgroup_text(
    place = "topleft", grow = FALSE, reflow = TRUE,
    colour = "white", fontface = "bold", size = 14, alpha = 0.65
  ) +
  geom_treemap_text(
    color = "white", place = "centre",
    reflow = TRUE, size = 9, fontface = "bold"
  ) +
  scale_fill_gradient2(
    low      = "#e74c3c",
    mid      = "#f5f5f5",
    high     = "#27ae60",
    midpoint = 0,
    labels   = dollar_format(),
    name     = "Total Profit"
  ) +
  labs(
    title    = "Treemap: Sub-Category Profitability",
    subtitle = "Tile size ∝ |Profit|   |   Colour: Green = Profitable · Red = Loss",
    caption  = "Source: Superstore Dataset"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 15),
    legend.position = "right"
  )
```

:::

::: callout-warning
**Critical Finding:**

```{r}
#| echo: false
unprofitable_subs <- subcategory_metrics %>%
  filter(Total_Profit < 0) %>%
  arrange(Total_Profit)

worst_sub   <- unprofitable_subs$`Sub-Category`[1]
worst_loss  <- dollar(abs(unprofitable_subs$Total_Profit[1]))
n_loss_subs <- nrow(unprofitable_subs)
loss_labels <- paste(
  paste0(unprofitable_subs$`Sub-Category`, " (", dollar(unprofitable_subs$Total_Profit), ")"),
  collapse = ", "
)
```

**`r n_loss_subs` sub-categories** are generating net losses: `r loss_labels`. The worst is **`r worst_sub`** at **–`r worst_loss`**. These require immediate strategic review — consider repricing, capping discounts, or discontinuation.
:::

---

# The Discount Problem

```{r}
#| fig-height: 11

discount_data <- superstore %>%
  mutate(
    Profit_Status = ifelse(Profit > 0, "Profitable", "Unprofitable"),
    Discount_Bin  = cut(
      Discount,
      breaks = c(-0.001, 0, 0.10, 0.20, 0.30, 0.50, 1.0),
      labels = c("0 %", "1–10 %", "11–20 %", "21–30 %", "31–50 %", "51–80 %")
    )
  )

disc_summary <- discount_data %>%
  group_by(Profit_Status) %>%
  summarise(
    Avg_Discount = mean(Discount),
    Count        = n(),
    Total_Profit = sum(Profit),
    .groups      = "drop"
  )

# --- Panel A: Scatter with LM trend ---
p_scatter <- ggplot(discount_data,
                    aes(x = Discount, y = Profit, color = Profit_Status)) +
  geom_point(alpha = 0.2, size = 1.0) +
  geom_smooth(aes(group = 1), method = "lm", se = TRUE,
              color = "#2c3e50", fill = "grey75", linewidth = 1.5) +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "#e74c3c", linewidth = 0.8) +
  scale_color_manual(values = c("Profitable" = "#27ae60",
                                "Unprofitable" = "#e74c3c")) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title    = "Discount vs Profit — Linear Trend",
    subtitle = "Negative correlation is clear across all orders",
    x = "Discount Rate", y = "Profit", color = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom")

# --- Panel B: Violin of discount by status ---
p_violin <- ggplot(discount_data,
                   aes(x = Profit_Status, y = Discount, fill = Profit_Status)) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_jitter(width = 0.12, alpha = 0.15, size = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 5,
               fill = "white", color = "black") +
  scale_fill_manual(values = c("Profitable"   = "#27ae60",
                               "Unprofitable" = "#e74c3c")) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title    = "Discount by Profit Status",
    subtitle = "Diamond = mean  |  Lines = Q1/Q2/Q3",
    x = NULL, y = "Discount Rate"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none")

# --- Panel C: Stacked bar ---
disc_bin_summary <- discount_data %>%
  group_by(Discount_Bin, Profit_Status) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Discount_Bin) %>%
  mutate(Pct = Count / sum(Count))

p_stacked <- ggplot(disc_bin_summary,
                    aes(x = Discount_Bin, y = Pct, fill = Profit_Status)) +
  geom_col(position = "fill", width = 0.7) +
  geom_text(
    aes(label = ifelse(Pct > 0.06, paste0(round(Pct * 100, 0), "%"), "")),
    position = position_fill(vjust = 0.5),
    size = 3.8, color = "white", fontface = "bold"
  ) +
  scale_fill_manual(values = c("Profitable"   = "#27ae60",
                               "Unprofitable" = "#e74c3c")) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title    = "Profitability Rate by Discount Bracket",
    subtitle = "Tipping point is between 20 – 30 % discount",
    x = "Discount Range", y = "% of Transactions", fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "bottom")

# --- Panel D: Heatmap Category × Discount ---
disc_cat_heat <- superstore %>%
  mutate(
    Discount_Bin = cut(
      Discount,
      breaks = c(-0.001, 0, 0.10, 0.20, 0.30, 0.50, 1.0),
      labels = c("0 %", "1–10 %", "11–20 %", "21–30 %", "31–50 %", "51–80 %")
    )
  ) %>%
  group_by(Category, Discount_Bin) %>%
  summarise(Avg_Profit = mean(Profit), Count = n(), .groups = "drop")

p_heat <- ggplot(disc_cat_heat,
                 aes(x = Discount_Bin, y = Category, fill = Avg_Profit)) +
  geom_tile(color = "white", linewidth = 0.9) +
  geom_text(
    aes(label = dollar(Avg_Profit, accuracy = 1)),
    size = 3.8, fontface = "bold", color = "white"
  ) +
  scale_fill_gradient2(
    low = "#e74c3c", mid = "grey88", high = "#27ae60",
    midpoint = 0, name = "Avg Profit",
    labels = dollar_format()
  ) +
  labs(
    title    = "Avg Profit Heat Map: Category × Discount Level",
    subtitle = "Red = loss  |  Green = profit",
    x = "Discount Range", y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

(p_scatter | p_violin) /
(p_stacked | p_heat)
```

::: callout-important
**Key Insights:**

```{r}
#| echo: false
avg_p <- percent(
  disc_summary$Avg_Discount[disc_summary$Profit_Status == "Profitable"],
  accuracy = 0.1
)
avg_u <- percent(
  disc_summary$Avg_Discount[disc_summary$Profit_Status == "Unprofitable"],
  accuracy = 0.1
)
pct_loss_top <- disc_bin_summary %>%
  filter(Discount_Bin == "51–80 %", Profit_Status == "Unprofitable") %>%
  pull(Pct) %>%
  percent(accuracy = 1)
```

- Profitable transactions average just **`r avg_p`** discount; unprofitable orders average **`r avg_u`** — a **`r round(as.numeric(sub("%","", avg_u)) - as.numeric(sub("%","", avg_p)), 0)` percentage point** gap.
- At the **51–80 %** discount tier, **`r pct_loss_top`** of transactions are unprofitable.
- The heatmap confirms damage is cross-category — Furniture is worst, but Technology and Office Supplies also suffer at high discount levels.
:::

---

# Regional Performance

```{r}
#| fig-height: 9

regional_metrics <- superstore %>%
  group_by(Region) %>%
  summarise(
    Total_Sales    = sum(Sales),
    Total_Profit   = sum(Profit),
    Avg_Profit     = mean(Profit),
    Profit_Margin  = Total_Profit / Total_Sales * 100,
    Count          = n(),
    Avg_Discount   = mean(Discount),
    Pct_Profitable = mean(Profit > 0) * 100,
    .groups        = "drop"
  ) %>%
  arrange(desc(Total_Profit))

# Panel A: Total Profit bar
p_reg_profit <- ggplot(regional_metrics,
                       aes(x = reorder(Region, Total_Profit),
                           y = Total_Profit, fill = Region)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = dollar(Total_Profit, accuracy = 1)),
            hjust = -0.1, fontface = "bold", size = 4.2) +
  scale_fill_brewer(palette = "Set3") +
  scale_y_continuous(labels = dollar_format(),
                     expand = expansion(mult = c(0, 0.28))) +
  coord_flip() +
  labs(title = "Total Profit by Region", x = NULL, y = "Total Profit") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))

# Panel B: Profit Margin %
p_reg_margin <- ggplot(regional_metrics,
                       aes(x = reorder(Region, Profit_Margin),
                           y = Profit_Margin, fill = Region)) +
  geom_col(width = 0.55, show.legend = FALSE) +
  geom_text(aes(label = paste0(round(Profit_Margin, 1), "%")),
            hjust = -0.1, fontface = "bold", size = 4.2) +
  scale_fill_brewer(palette = "Set3") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.28))) +
  coord_flip() +
  labs(title = "Profit Margin by Region (%)", x = NULL, y = "Profit Margin (%)") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))

# Panel C: Boxplot of transaction-level profit
p_reg_box <- ggplot(superstore,
                    aes(x = reorder(Region, Profit, FUN = median),
                        y = Profit, fill = Region)) +
  geom_boxplot(alpha = 0.75, outlier.alpha = 0.2, outlier.size = 1) +
  stat_summary(fun = mean, geom = "point", shape = 23, size = 4.5,
               fill = "white", color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "#e74c3c", linewidth = 0.8) +
  scale_fill_brewer(palette = "Set3") +
  scale_y_continuous(labels = dollar_format()) +
  coord_flip() +
  labs(title    = "Profit Distribution by Region",
       subtitle = "Diamond = mean  |  Red dashed = break-even",
       x = NULL, y = "Profit per Transaction") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none")

# Panel D: Bubble — Avg Discount vs Avg Profit
p_reg_bubble <- ggplot(regional_metrics,
                       aes(x = Avg_Discount, y = Avg_Profit,
                           color = Profit_Margin, size = Total_Sales,
                           label = Region)) +
  geom_point(alpha = 0.9) +
  geom_text_repel(size = 4.5, fontface = "bold", show.legend = FALSE,
                  box.padding = 0.5) +
  scale_color_gradient(low = "#e74c3c", high = "#27ae60",
                       name = "Profit Margin (%)") +
  scale_size_continuous(labels = dollar_format(), name = "Total Sales",
                        range = c(8, 18)) +
  scale_x_continuous(labels = percent_format()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title    = "Avg Discount vs Avg Profit per Region",
    subtitle = "Bubble size ∝ Total Sales  |  Colour ∝ Profit Margin",
    x = "Average Discount Rate", y = "Average Profit / Transaction"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))

(p_reg_profit | p_reg_margin) /
(p_reg_box    | p_reg_bubble)
```

::: callout-tip
**Regional Findings:**

```{r}
#| echo: false
best_region         <- regional_metrics$Region[1]
best_region_total   <- dollar(regional_metrics$Total_Profit[1])
worst_region        <- regional_metrics$Region[nrow(regional_metrics)]
best_margin_region  <- regional_metrics$Region[which.max(regional_metrics$Profit_Margin)]
best_margin_pct     <- paste0(round(max(regional_metrics$Profit_Margin), 1), "%")
```

- **`r best_region`** leads total profit with **`r best_region_total`**; **`r best_margin_region`** achieves the highest profit margin at **`r best_margin_pct`**.
- **`r worst_region`** lags on both metrics — likely driven by a higher average discount rate or unfavourable product mix.
- The bubble chart confirms the discount–profit relationship holds at the regional level.
:::

---

# Temporal Analysis

## Sales & Profit Trends

```{r}
#| fig-height: 11

monthly_by_cat <- superstore %>%
  group_by(YearMonth, Category) %>%
  summarise(
    Total_Sales  = sum(Sales),
    Total_Profit = sum(Profit),
    .groups      = "drop"
  )

monthly_overall <- superstore %>%
  group_by(YearMonth) %>%
  summarise(
    Total_Sales   = sum(Sales),
    Total_Profit  = sum(Profit),
    Profit_Margin = Total_Profit / Total_Sales * 100,
    .groups       = "drop"
  )

# Sales trend by category
p_sales_trend <- ggplot(monthly_by_cat,
                        aes(x = YearMonth, y = Total_Sales, color = Category)) +
  geom_line(linewidth = 1.2, alpha = 0.85) +
  geom_point(size = 1.6, alpha = 0.6) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_brewer(palette = "Set1") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Monthly Sales Trend by Category",
       x = NULL, y = "Total Sales", color = "Category") +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

# Profit trend by category
p_profit_trend <- ggplot(monthly_by_cat,
                         aes(x = YearMonth, y = Total_Profit, color = Category)) +
  geom_line(linewidth = 1.2, alpha = 0.85) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.8) +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_brewer(palette = "Set1") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(title = "Monthly Profit Trend by Category",
       x = NULL, y = "Total Profit", color = "Category") +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

# Profit margin area + LOESS
p_margin_trend <- ggplot(monthly_overall,
                          aes(x = YearMonth, y = Profit_Margin)) +
  geom_area(fill = "#3498db", alpha = 0.18) +
  geom_line(color = "#2c3e50", linewidth = 1.2) +
  geom_smooth(method = "loess", se = TRUE, color = "#e74c3c",
              fill = "#e74c3c", alpha = 0.15, linewidth = 1.1) +
  geom_hline(yintercept = 0, linetype = "dashed",
             color = "#e74c3c", linewidth = 0.8) +
  scale_y_continuous(labels = function(x) paste0(round(x, 1), "%")) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b %Y") +
  labs(
    title    = "Monthly Overall Profit Margin",
    subtitle = "Red LOESS line = long-run trend  |  Shaded band = 95% CI",
    x = NULL, y = "Profit Margin (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title  = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

p_sales_trend / p_profit_trend / p_margin_trend
```

## Seasonal Patterns

```{r}
#| fig-height: 8

heatmap_data <- superstore %>%
  group_by(Year, Month) %>%
  summarise(
    Total_Profit = sum(Profit),
    Total_Sales  = sum(Sales),
    .groups      = "drop"
  )

# Month × Year profit heatmap
p_heatmap <- ggplot(heatmap_data,
                    aes(x = Month, y = factor(Year), fill = Total_Profit)) +
  geom_tile(color = "white", linewidth = 0.8) +
  geom_text(
    aes(label = dollar(Total_Profit, accuracy = 0.1, scale = 0.001, suffix = "K")),
    size = 3.2, fontface = "bold", color = "white"
  ) +
  scale_fill_gradient2(
    low = "#e74c3c", mid = "grey88", high = "#27ae60",
    midpoint = 0, name = "Total Profit",
    labels = dollar_format()
  ) +
  labs(
    title    = "Profit Heatmap — Year × Month",
    subtitle = "Green = high profit  |  Red = low or negative profit",
    x = "Month", y = "Year"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid = element_blank()
  )

# Avg monthly sales ridge across years
monthly_avg <- superstore %>%
  group_by(Month, Category) %>%
  summarise(Avg_Sales = mean(Sales), .groups = "drop")

p_season <- ggplot(monthly_avg,
                   aes(x = Month, y = Avg_Sales, fill = Category)) +
  geom_col(position = "dodge", width = 0.85, alpha = 0.9) +
  scale_fill_brewer(palette = "Set1") +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title    = "Average Monthly Sales by Category",
    subtitle = "Seasonal peaks are most visible in Q4 (Oct–Dec)",
    x = "Month", y = "Avg Sales per Month",
    fill = "Category"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold"),
    legend.position = "bottom"
  )

p_season / p_heatmap
```

::: callout-tip
**Temporal Findings:**

```{r}
#| echo: false
best_month <- heatmap_data %>%
  group_by(Month) %>%
  summarise(AvgP = mean(Total_Profit)) %>%
  slice_max(AvgP, n = 1)
worst_month <- heatmap_data %>%
  group_by(Month) %>%
  summarise(AvgP = mean(Total_Profit)) %>%
  slice_min(AvgP, n = 1)
```

- **`r best_month$Month`** is consistently the strongest month, averaging `r dollar(best_month$AvgP)` in profit — likely driven by end-of-year purchasing.
- **`r worst_month$Month`** is the weakest, averaging `r dollar(worst_month$AvgP)`.
- The LOESS trend in the margin chart reveals the long-run health trajectory of the business.
:::

## Interactive Time Series (Plotly)

```{r}
#| fig-height: 5

p_anim <- plot_ly(
  monthly_by_cat,
  x      = ~YearMonth,
  y      = ~Total_Profit,
  color  = ~Category,
  colors = RColorBrewer::brewer.pal(3, "Set1"),
  type   = "scatter",
  mode   = "lines+markers",
  line   = list(width = 2),
  marker = list(size = 5),
  text   = ~paste0(
    "<b>", Category, "</b><br>",
    format(YearMonth, "%b %Y"), "<br>",
    "Profit: ", dollar(Total_Profit, accuracy = 1)
  ),
  hovertemplate = "%{text}<extra></extra>"
) %>%
  layout(
    title = list(
      text = "<b>Monthly Profit by Category (Interactive)</b><br><sup>Drag slider to zoom — hover for details</sup>"
    ),
    xaxis = list(
      title       = "",
      rangeslider = list(visible = TRUE),
      rangeselector = list(
        buttons = list(
          list(count = 6,  label = "6M",  step = "month", stepmode = "backward"),
          list(count = 12, label = "1Y",  step = "month", stepmode = "backward"),
          list(count = 24, label = "2Y",  step = "month", stepmode = "backward"),
          list(step = "all", label = "All")
        )
      )
    ),
    yaxis       = list(title = "Total Profit ($)", tickformat = "$,.0f"),
    legend      = list(orientation = "h", x = 0, y = -0.25),
    hovermode   = "x unified",
    plot_bgcolor  = "rgba(0,0,0,0)",
    paper_bgcolor = "rgba(0,0,0,0)"
  )

p_anim
```

---

# Customer Segment Analysis

```{r}
#| fig-height: 10

segment_metrics <- superstore %>%
  group_by(Segment) %>%
  summarise(
    Total_Sales    = sum(Sales),
    Total_Profit   = sum(Profit),
    Avg_Profit     = mean(Profit),
    Profit_Margin  = Total_Profit / Total_Sales * 100,
    Avg_Discount   = mean(Discount),
    Count          = n(),
    Pct_Profitable = mean(Profit > 0) * 100,
    .groups        = "drop"
  ) %>%
  arrange(desc(Total_Profit))

# Panel A: Sales vs Profit scatter, faceted
p_seg_scatter <- ggplot(superstore,
                        aes(x = Sales, y = Profit, color = Segment)) +
  geom_point(alpha = 0.25, size = 1.2) +
  geom_smooth(method = "lm", se = TRUE, linewidth = 1.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  facet_wrap(~Segment, ncol = 3) +
  scale_x_continuous(labels = dollar_format(), trans = "log10") +
  scale_y_continuous(labels = dollar_format()) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title    = "Sales vs Profit by Customer Segment",
    subtitle = "Log₁₀ scale on x-axis  |  Shaded band = 95% confidence interval",
    x = "Sales (log₁₀)", y = "Profit"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"),
        legend.position = "none")

# Panel B: Avg profit bar
p_seg_bar <- ggplot(segment_metrics,
                    aes(x = reorder(Segment, Avg_Profit),
                        y = Avg_Profit, fill = Segment)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  geom_text(aes(label = dollar(Avg_Profit, accuracy = 0.01)),
            hjust = -0.1, fontface = "bold", size = 4.2) +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(labels  = dollar_format(),
                     expand  = expansion(mult = c(0, 0.35))) +
  coord_flip() +
  labs(title    = "Avg Profit / Transaction by Segment",
       subtitle = "Which segment earns most per order?",
       x = NULL, y = "Avg Profit") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))

# Panel C: Radar chart (normalised)
seg_radar <- segment_metrics %>%
  mutate(
    Norm_Sales    = rescale(Total_Sales),
    Norm_Profit   = rescale(Total_Profit),
    Norm_Margin   = rescale(Profit_Margin),
    Norm_LowDisc  = 1 - rescale(Avg_Discount),
    Norm_PctProf  = rescale(Pct_Profitable)
  ) %>%
  select(Segment, starts_with("Norm")) %>%
  pivot_longer(-Segment, names_to = "Dimension", values_to = "Score") %>%
  mutate(Dimension = recode(Dimension,
    "Norm_Sales"   = "Total Sales",
    "Norm_Profit"  = "Total Profit",
    "Norm_Margin"  = "Profit Margin",
    "Norm_LowDisc" = "Low Discount\n(inverted)",
    "Norm_PctProf" = "% Profitable"
  ))

p_radar <- ggplot(seg_radar,
                  aes(x = Dimension, y = Score,
                      color = Segment, group = Segment)) +
  geom_polygon(aes(fill = Segment), alpha = 0.12, show.legend = FALSE) +
  geom_line(linewidth = 1.3) +
  geom_point(size = 3.5) +
  coord_polar() +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  scale_y_continuous(limits = c(0, 1.1)) +
  labs(
    title    = "Segment Performance Radar",
    subtitle = "Normalised 0–1  |  Outer edge = better performance",
    x = NULL, y = NULL, color = "Segment"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold"),
    axis.text.y     = element_blank(),
    axis.ticks.y    = element_blank(),
    legend.position = "bottom"
  )

p_seg_scatter /
(p_seg_bar | p_radar)
```

::: callout-note
**Segment Insights:**

```{r}
#| echo: false
best_seg_avg   <- segment_metrics$Segment[which.max(segment_metrics$Avg_Profit)]
best_avg_val   <- dollar(max(segment_metrics$Avg_Profit), accuracy = 0.01)
best_seg_total <- segment_metrics$Segment[1]
best_tot_val   <- dollar(segment_metrics$Total_Profit[1])
```

- **`r best_seg_avg`** achieves the highest average profit per transaction (**`r best_avg_val`**) — best for margin-focused strategy.
- **`r best_seg_total`** generates the largest total profit (**`r best_tot_val`**) through transaction volume.
- The radar chart reveals no single segment dominates all five dimensions — each has distinct strengths and weaknesses.
:::

---

# Key Insights & Recommendations

## Performance Summary

```{r}
# Compute all metrics dynamically — no hardcoding
# Compute all metrics dynamically — no hardcoding
pct_loss_txns <- mean(superstore$Profit < 0)

summary_metrics <- tibble(
  Metric = c(
    "Total Transactions",
    "Total Sales Revenue",
    "Total Profit",
    "Overall Profit Margin",
    "Unprofitable Transactions",
    "Total Loss (Unprofitable Orders)",
    "Most Profitable Category",
    "Least Profitable Sub-Category",
    "Best Region (Total Profit)",
    "Best Region (Profit Margin)",
    "Avg Discount — Profitable Orders",
    "Avg Discount — Unprofitable Orders"
  ),
  Value = c(
    format(nrow(superstore), big.mark = ","),
    dollar(sum(superstore$Sales)),
    dollar(sum(superstore$Profit)),
    paste0(round(sum(superstore$Profit) / sum(superstore$Sales) * 100, 2), " %"),
    paste0(format(sum(superstore$Profit < 0), big.mark = ","),
           "  (", round(pct_loss_txns * 100, 1), " % of orders)"),
    dollar(sum(superstore$Profit[superstore$Profit < 0])),
    superstore %>% group_by(Category) %>%
      summarise(tp = sum(Profit), .groups = "drop") %>%
      slice_max(tp, n = 1) %>% pull(Category),
    superstore %>% group_by(`Sub-Category`) %>%
      summarise(tp = sum(Profit), .groups = "drop") %>%
      slice_min(tp, n = 1) %>% pull(`Sub-Category`),
    superstore %>% group_by(Region) %>%
      summarise(tp = sum(Profit), .groups = "drop") %>%
      slice_max(tp, n = 1) %>% pull(Region),
    superstore %>% group_by(Region) %>%
      summarise(ts = sum(Sales), tp = sum(Profit), .groups = "drop") %>%
      mutate(margin = tp / ts) %>%
      slice_max(margin, n = 1) %>% pull(Region),
    percent(mean(superstore$Discount[superstore$Profit > 0]), accuracy = 0.1),
    percent(mean(superstore$Discount[superstore$Profit <= 0]), accuracy = 0.1)
  )
)
```

## Interactive Profitability Dashboard

```{r}
#| fig-height: 5

fig_cat <- plot_ly(
  category_metrics,
  x    = ~reorder(Category, Total_Profit),
  y    = ~Total_Profit,
  type = "bar",
  color = ~Category,
  colors = RColorBrewer::brewer.pal(3, "Set2"),
  text = ~paste0(
    "<b>", Category, "</b><br>",
    "Total Profit: ", dollar(Total_Profit), "<br>",
    "Profit Margin: ", round(Profit_Margin, 1), "%<br>",
    "Transactions: ", format(Count, big.mark = ",")
  ),
  hovertemplate = "%{text}<extra></extra>"
) %>%
  layout(
    title   = "<b>Category</b>",
    xaxis   = list(title = ""),
    yaxis   = list(title = "Total Profit ($)", tickformat = "$,.0f"),
    showlegend = FALSE
  )

fig_reg <- plot_ly(
  regional_metrics,
  x    = ~reorder(Region, Total_Profit),
  y    = ~Total_Profit,
  type = "bar",
  color = ~Region,
  colors = RColorBrewer::brewer.pal(4, "Set3"),
  text = ~paste0(
    "<b>", Region, "</b><br>",
    "Total Profit: ", dollar(Total_Profit), "<br>",
    "Profit Margin: ", round(Profit_Margin, 1), "%<br>",
    "Transactions: ", format(Count, big.mark = ",")
  ),
  hovertemplate = "%{text}<extra></extra>"
) %>%
  layout(
    title   = "<b>Region</b>",
    xaxis   = list(title = ""),
    yaxis   = list(title = "Total Profit ($)", tickformat = "$,.0f"),
    showlegend = FALSE
  )

fig_seg <- plot_ly(
  segment_metrics,
  x    = ~reorder(Segment, Avg_Profit),
  y    = ~Avg_Profit,
  type = "bar",
  color = ~Segment,
  colors = RColorBrewer::brewer.pal(3, "Dark2"),
  text = ~paste0(
    "<b>", Segment, "</b><br>",
    "Avg Profit: ", dollar(Avg_Profit, accuracy = 0.01), "<br>",
    "Total Profit: ", dollar(Total_Profit), "<br>",
    "% Profitable: ", round(Pct_Profitable, 1), "%"
  ),
  hovertemplate = "%{text}<extra></extra>"
) %>%
  layout(
    title   = "<b>Customer Segment</b>",
    xaxis   = list(title = ""),
    yaxis   = list(title = "Avg Profit ($)", tickformat = "$,.2f"),
    showlegend = FALSE
  )

subplot(fig_cat, fig_reg, fig_seg, nrows = 1,
        titleX = TRUE, titleY = TRUE, shareY = FALSE,
        margin = 0.06) %>%
  layout(
    title = list(
      text = "<b>Superstore Profitability Dashboard</b><br><sup>Hover over bars for details</sup>",
      x    = 0.5
    ),
    plot_bgcolor  = "rgba(0,0,0,0)",
    paper_bgcolor = "rgba(0,0,0,0)"
  )
```

::: {.callout-note appearance="default" icon="true"}
## Business Recommendations

**1. Discount Strategy Reform**

```{r}
#| echo: false
avg_p_rec <- percent(mean(superstore$Discount[superstore$Profit > 0]),  accuracy = 0.1)
avg_u_rec <- percent(mean(superstore$Discount[superstore$Profit <= 0]), accuracy = 0.1)
```

- Implement a **hard discount cap at 30 %** — data shows that losses spike sharply beyond this threshold, especially in Furniture.
- Profitable orders average `r avg_p_rec` discount; unprofitable average `r avg_u_rec` — closing this gap is the single highest-leverage action available.
- Require management sign-off for any discount exceeding 20 %.

**2. Product Line Optimisation**

```{r}
#| echo: false
loss_sub_labels <- subcategory_metrics %>%
  filter(Total_Profit < 0) %>%
  arrange(Total_Profit) %>%
  mutate(lbl = paste0(`Sub-Category`, " (", dollar(Total_Profit), ")")) %>%
  pull(lbl)
top_subs <- subcategory_metrics %>%
  slice_max(Total_Profit, n = 3) %>%
  pull(`Sub-Category`)
```

- Loss-making sub-categories requiring action: `r paste(loss_sub_labels, collapse = " · ")`.
- Consider repricing or discontinuing **Tables** (worst performer).
- Double down on: **`r paste(top_subs, collapse = "**, **")`** — the top-3 most profitable sub-categories.

**3. Regional Strategy**

```{r}
#| echo: false
best_r_rec  <- regional_metrics$Region[1]
worst_r_rec <- regional_metrics$Region[nrow(regional_metrics)]
```

- **`r best_r_rec`** region: proven performance — maintain investment and use as the benchmark.
- **`r worst_r_rec`** region: diagnose root causes — is it discount abuse, product mix, or market conditions? Implement corrective actions in the next quarter.

**4. Segment Development**

```{r}
#| echo: false
best_seg_rec <- segment_metrics$Segment[which.max(segment_metrics$Avg_Profit)]
vol_seg_rec  <- segment_metrics$Segment[1]
```

- **`r best_seg_rec`**: highest profit per transaction — grow this segment through targeted outreach and premium product bundles.
- **`r vol_seg_rec`**: high volume, lower margin — focus campaigns on reducing discount frequency rather than acquiring more customers.
- **Corporate**: offer value-added services, SLAs, and dedicated account management instead of price discounts.
:::

---


